---
title: "Potential Sample Frames"
author: "Mike McManus"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Libraries

```{r libraries}
library(tidyverse)
library(sf)
library(mapview)
library(janitor)
library(leaflet)
library(leafpop) # for popups in 
library(leafsync) # sync mapview maps
library(leafem) # add labels to map
library(patchwork)
library(nhdplusTools)
library(tmap)
library(forcats)
library(gifski) #for tmap animation
library(plotly)
library(webshot) # save mapview object
library(nngeo)
library(micromap)
library(tidyverse)
library(spsurvey)

```

## Log

2026-08-31 sorted bbl_8 by SEGMENT_NA and see Alligator Bayou and Alligator Bayou_GeneratingPlant, but both have Major_Seg assignment of St. Andrew Bay. I believe the Alligator Bayou by the generating plant should be assigned to North Bay as Major_Seg.

2026-08-18 Ryann Rossi loaded Bays_Bayous_Lakes_V8.shp to GIS data. It contains updates to Alligator Bayou. Need to run that spatial data to make the 3 sample frames.

2026-08-11 made sf_bbl1, which is a sample frame of 80 polygons containing bays, bayous, creeks, lagoons, and lakes. saved as a shapefile in output_sampleframes_surveys folder.

2026-08-10 made sf_baysegmt1, which contains only 5 bay segments, including Deer Point Lake. No bayous, creeks or lagoons are included. Stratification for the survey design would be by SEGMENT_NA.

2026-08-06 made sf_all1, all one survey frame. Checked bay topologies.

## Read Bays, Bayous and Lakes V7

Using bbl_8 from Ryan Rossi.

Using bbl_7 from Rose Horn.

Using bbl_5 that Rose Horn sent, which has adjustments to distinguish polygons that had the same name. Still wonder why have Shape_Leng, Shape_Le_1 and Shape_Le_2.

Rose Horn took 2 polygons for Alligator Bayou, which borders the U.S. Coast Guard Station on St. Andrew Bay, and splits them into separate sampling areas. Also, in version the new update reflects the removal of small edge portions of St. Andrew Bay that were along the edges of East Bay and Grand Lagoon. Looks like I should pick-up Upper and Lower Burnt Mill Creek Stations.

Need to see if I have 8 polygon names for: East Bay, Grand Lagoon, North Bay, St. Andrew Bay, West Bay, Crooked and Burnt Mill Creeks, and Deer Point Lake as I want to use those 6 polygons in linked micromap. Perhaps by using SEGMENT_NA and Major_Seg I can get the polygons I want.

```{r bbl_v8}

bbl_8 <- st_read("data_gis/Bays_Bayous_Lakes_V8.shp")

bbl_8 <- bbl_8 %>% st_transform(3087)

names(bbl_8)
summary(bbl_8)
mapview(bbl_8)


distinct(bbl_8, SEGMENT_NA) %>% dplyr::arrange(., SEGMENT_NA)
# now with bbl_8 have 194 distinct segment names and 194 obs, I believe, because Harrison Bayou was split in to Harrison Bayou and Little Johnson Bayou. with bbl_5 have 193 distinct segment names and 193 obs. Previously in bbl_4 had 190 distinct SEGMENT_NA, but 191 obs suggesting one polygon is sharing a name

distinct(bbl_8, Major_Seg) %>% dplyr::arrange(., Major_Seg)
# returns 8 Major_Seg

st_is_valid(bbl_8, reason = TRUE)
# 189 is SEGMENT_NA == White Flank Pond

# I can't see ring self-intersection.
dplyr::filter(bbl_8, SEGMENT_NA == "White Flank Pond") %>% 
  mapview(.)

bbl_8 <- st_make_valid(bbl_8)

st_is_valid(bbl_8, reason = T)
```

## Filter Bays and Check Topology
For North Bay had to add SEGMENT_NA == "Alligator Bayou_GeneratingPlant". 
```{r filter_bays_bayous}
# filter west bay
bayous_wb1 <- dplyr::filter(bbl_8, SEGMENT_NA == "Burnt Mill Creek" | SEGMENT_NA == "Crooked Creek" | SEGMENT_NA == "Warren Bayou" | SEGMENT_NA == "Gulf ICWW between Choctawhatchee Bay and St. Andrew Bay" | SEGMENT_NA == "West Bay" | SEGMENT_NA == "Johnson Bayou_WestBay" | SEGMENT_NA == "Basin Bayou" | SEGMENT_NA == "Botheration Bayou" | SEGMENT_NA == "Doyle Bayou" | SEGMENT_NA == "Harrison Bayou" | SEGMENT_NA == "Little Johnson Bayou" |SEGMENT_NA == "Ward Creek")
# returns 12 obs matching 12 segment names

bayous_wb2 <- bayous_wb1 |> 
  mutate(
  SEGMENT_NA = SEGMENT_NA |>
    replace_values(
      "Gulf ICWW between Choctawhatchee Bay and St. Andrew Bay" ~ "West Bay ICWW"
    )
  )  

st_is_valid(bayous_wb2, reason = T)
# identified Johnson Bayou_WestBay as Ring Self-Intersection
# with bbl_8 no longer get Ring Self_Intersection as get all valid geometry
# bayous_wb2 <- st_make_valid(bayous_wb2)
# st_is_valid(bayous_wb2, reason = T)
mapview(bayous_wb2)


# filter for North Bay and include Robinson Bay, which is near North Bay and St. Andrew Bay border
bayous_nb1 <- dplyr::filter(bbl_8, Major_Seg == "North Bay" | SEGMENT_NA == "Robinson Bayou" | SEGMENT_NA == "Alligator Bayou_GeneratingPlant") %>% 
  filter_out(., SEGMENT_NA == "Deer Point Lake")
# using bbl_8 returns 17 obs/polygon

st_is_valid(bayous_nb1, reason = T)
mapview(bayous_nb1)

# filter for deer point lake
bayous_dpl1 <- dplyr::filter(bbl_8, SEGMENT_NA == "Deer Point Lake")

st_is_valid(bayous_dpl1, reason = T)
mapview(bayous_dpl1)

# filter for east bay 
bayous_eb1 <- dplyr::filter(bbl_8, Major_Seg == "East Bay")
# returns 42 polygons

bayous_eb2 <- bayous_eb1 |>
  mutate(
  SEGMENT_NA = SEGMENT_NA |>
    replace_values(
      "Gulf ICWW between St. Andrew Bay and St. Joseph Bay" ~ "East Bay ICWW"
    )
  )

st_is_valid(bayous_eb2, reason = T)
mapview(bayous_eb2)

# filter for st andrew bay
# dropped Crooked Island Sound polygon
bayous_sab1 <- dplyr::filter(bbl_8, SEGMENT_NA == "Alligator Bayou" |  SEGMENT_NA == "Grand Lagoon" |SEGMENT_NA == "Hog Island Sound" | SEGMENT_NA == "Johnson Bayou"| SEGMENT_NA == "Lake Huntington" | SEGMENT_NA == "Massalina Bayou" | SEGMENT_NA == "Posten Bayou" | SEGMENT_NA == "Pretty Bayou" | SEGMENT_NA == "St. Andrew Bay")
# returns 9 obs

st_is_valid(bayous_sab1, reason = T)

mapview(bayous_sab1)

```

## Mapview Bays and Bayous

For West Bay, from Burnt Mill Creek to West ICWW I think use downstream of the powerlines as the cutoff for that part of the sample frame.

```{r mapview_bays_bayous}

mapview(bayous_dpl1, col.regions = "purple") + mapview(bayous_eb2, col.regions = "red") + mapview(bayous_nb1, col.regions = "orange") + mapview(bayous_sab1, col.regions = "green") + mapview(bayous_wb2, col.regions = "blue")

```

## Areas

Not sure about these areas as topology of polygons has not been checked with `st_is_valid`.

```{r area_bays}

st_area(bayous_dpl1)

st_area(bayous_eb2)

sum(st_area(bayous_eb2))

dplyr::filter(bayous_eb2, SEGMENT_NA == "East Bay") %>% 
  st_area(.)

st_area(bayous_nb1)

sum(st_area(bayous_nb1))

dplyr::filter(bayous_nb1, SEGMENT_NA == "North Bay") %>% 
  st_area(.)

dplyr::filter(bayous_nb1, SEGMENT_NA == "Deer Point Lake") %>% 
  st_area(.)

st_area(bayous_sab1)

sum(st_area(bayous_sab1))

dplyr::filter(bayous_sab1, SEGMENT_NA == "St. Andrew Bay") %>% 
  st_area(.)

st_area(bayous_wb2)

sum(st_area(bayous_wb2))

dplyr::filter(bayous_wb2, SEGMENT_NA == "West Bay") %>% 
  st_area(.)

```

## All One Survey Frame Geoprocessing

I think the steps for combining polygons are:

1.  Check topology of each polygon with `st_is_valid(x, reason = TRUE)` If need, use 'st_make_valid`function.`

2.  `For each polygon do`st_combine`. I believe this unions any internal polygon, and the object now becomes a list of 1.`

3.  `Use`st_union`on 2 polygons to make 1 large sfc_multipolyon object.`

4.  `Use 'st_sf` to make 1 large sfc_multipolygon into an sf object.

```{r all_one_survey_geoprocessing}

st_is_valid(bayous_wb2, reason = T)
bayous_wb3 <- st_make_valid(bayous_wb2)
st_is_valid(bayous_wb3, reason = T)

st_is_valid(bayous_nb1, reason = T)

comb_wb1 <- st_combine(bayous_wb3)
st_is_valid(comb_wb1, reason = T)
comb_wb1 <- st_make_valid(comb_wb1)
st_is_valid(comb_wb1, reason = T)

comb_nb1 <- st_combine(bayous_nb1)
st_is_valid(comb_nb1, reason = T)
comb_nb1 <- st_make_valid(comb_nb1)
st_is_valid(comb_nb1)


wbnb1 <- st_union(comb_wb1,comb_nb1)
mapview(wbnb1)

# note change from large sfc_multipolygon to sf object
wbnb1 <- st_sf(wbnb1)
class(wbnb1)
mapview(wbnb1)

st_is_valid(bayous_dpl1, reason = T)
# st_combine makes object a list of 1
bayous_dpl1 <- st_combine(bayous_dpl1)

# st_union makes object into sfc_multipolygon
wbnbdpl1 <- st_union(bayous_dpl1, wbnb1)

# st_sf makes object into sf object
wbnbdpl1 <-st_sf(wbnbdpl1)
class(wbnbdpl1)

mapview(wbnbdpl1)

st_is_valid(bayous_sab1, reason = T)

# st_combine makes object a list of 1
bayous_sab1 <- st_combine(bayous_sab1)

st_is_valid(bayous_sab1, reason = T)
bayous_sab1 <- st_make_valid(bayous_sab1)
st_is_valid(bayous_sab1)

st_is_valid(wbnbdpl1, reason = T)

# st_union makes object into sfc_multipolygon
wbnbdplsab1 <- st_union(bayous_sab1, wbnbdpl1)

# st_sf makes object into sf object
wbnbdplsab1 <-st_sf(wbnbdplsab1)
class(wbnbdplsab1)

mapview(wbnbdplsab1)

st_is_valid(bayous_eb2, reason = T)

bayous_eb2 <- st_combine(bayous_eb2)

st_is_valid(bayous_eb2, reason = T)
bayous_eb2 <- st_make_valid(bayous_eb2)
st_is_valid(bayous_eb2, reason = T)


wbnbdplsabeb1 <- st_union(wbnbdplsab1, bayous_eb2)

wbnbdplsabeb1 <- st_sf(wbnbdplsabeb1)
class(wbnbdplsabeb1)

mapview(wbnbdplsabeb1)

st_area(wbnbdplsabeb1)

st_is_valid(wbnbdplsabeb1, reason = T)

sf_all1 <- wbnbdplsabeb1

# Writing a single shapefile to the working directory
st_write(sf_all1,
dsn = "output_sampleframes_surveys/sf_all1.shp",
driver = "ESRI Shapefile",
append = FALSE)

```

## Bay-Segments Geoprocesssing

This sample frame has only the main bay segments, including Deer Point Lake. Stratification for the survey design will be done by SEGMENT_NA. This sample frame does not contain any bayous, creeks, or lagoons.

```{r bay_segments_geoprocessing}

bay_segmts1 <- dplyr::filter(bbl_7, SEGMENT_NA == "West Bay" | SEGMENT_NA == "North Bay" | SEGMENT_NA == "Deer Point Lake" | SEGMENT_NA == "St. Andrew Bay" | SEGMENT_NA == "East Bay")

st_is_valid(bay_segmts1)

mapview(bay_segmts1)

class(bay_segmts1)

st_area(bay_segmts1)

bay_segmts1$area_m2 <- st_area(bay_segmts1)

sum(st_area(bay_segmts1))

# sample frame
sf_baysegmt1 <- bay_segmts1

mapview(sf_baysegmt1)

# Writing a single shapefile to the working directory
st_write(sf_baysegmt1,
dsn = "output_sampleframes_surveys/sf_baysegmt1.shp",
driver = "ESRI Shapefile",
append = FALSE)

```

## Bays, Bayous, Creeks, and Lakes Geoprocessing

I think in my inital coding I forgot to include bayous from North Bay, and on 08/11/2026 I corrected that.

Filter for bay segments, bayous, creeks, and lakes to make a bays and bayous sample frame. Create a stratification variable, waterbody type (wbody_type). The levels in waterbody type include 5 bay segments, this inlcudes Deer Point Lake, and for each bay segment that has bayous and other waterbody types, there is a correspond bayou indicated by "\_b". For example, there is an eb, for the East Bay segment, and then eb_b, which occurs 41 times, for the bayous, creek, etc. in East Bay.

```{r bays_bayous_geoprocessing}


bbl_frame1 <- dplyr::filter(bbl_7, SEGMENT_NA == "Burnt Mill Creek" | SEGMENT_NA == "Crooked Creek" | SEGMENT_NA == "Warren Bayou" | SEGMENT_NA == "Gulf ICWW between Choctawhatchee Bay and St. Andrew Bay" | SEGMENT_NA == "West Bay" | SEGMENT_NA == "Johnson Bayou_WestBay" | SEGMENT_NA == "Basin Bayou" | SEGMENT_NA == "Botheration Bayou" | SEGMENT_NA == "Doyle Bayou" | SEGMENT_NA == "Harrison Bayou" | SEGMENT_NA == "Little Johnson Bayou" |SEGMENT_NA == "Ward Creek" | Major_Seg == "North Bay" | SEGMENT_NA == "Robinson Bayou"| SEGMENT_NA == "Deer Point Lake" | Major_Seg == "East Bay" | SEGMENT_NA == "Alligator Bayou" | SEGMENT_NA == "Grand Lagoon" |SEGMENT_NA == "Hog Island Sound" | SEGMENT_NA == "Johnson Bayou"| SEGMENT_NA == "Lake Huntington" | SEGMENT_NA == "Massalina Bayou" | SEGMENT_NA == "Posten Bayou" | SEGMENT_NA == "Pretty Bayou" | SEGMENT_NA == "St. Andrew Bay" | SEGMENT_NA == "Gulf ICWW between St. Andrew Bay and St. Joseph Bay")
# returns 80 obs

mapview(bbl_frame1)

bbl_frame1 %>% 
  group_by(Major_Seg, SEGMENT_NA) %>% 
  summarize(., count = dplyr::n()) %>% 
  print(n = Inf)


bbl_frame2 <- bbl_frame1 |> 
  mutate(
  SEGMENT_NA = SEGMENT_NA |>
    replace_values(
      "Gulf ICWW between Choctawhatchee Bay and St. Andrew Bay" ~ "West Bay ICWW",
       "Gulf ICWW between St. Andrew Bay and St. Joseph Bay" ~ "East Bay ICWW"
    )
  )  

st_is_valid(bbl_frame2)

# creating a stratifcation variable of waterbody type, wbody_type
bbl_frame3 <- bbl_frame2 |>
  mutate(
    wbody_type = case_when(
      (Major_Seg == "East Bay" & SEGMENT_NA == "East Bay" ~ "eb"),
      (Major_Seg == "East Bay" & SEGMENT_NA != "East Bay" ~ "eb_b"),
      (Major_Seg == "North Bay" & SEGMENT_NA == "North Bay" ~ "nb"),
      (Major_Seg == "North Bay" & SEGMENT_NA == "Deer Point Lake" ~ "dpl"),
      (Major_Seg == "North Bay" & SEGMENT_NA != "Deer Point Lake" ~ "nb_b"),
      (Major_Seg == "St. Andrew Bay" & SEGMENT_NA == "St. Andrew Bay" ~ "sab"),
      (Major_Seg == "St. Andrew Bay" & SEGMENT_NA != "St. Andrew Bay" ~ "sab_b"),
      (Major_Seg == "West Bay" & SEGMENT_NA == "West Bay" ~ "wb"),
      (Major_Seg == "West Bay" & SEGMENT_NA != "North Bay" ~ "wb_b"),
      SEGMENT_NA == "West Bay ICWW" ~ "wb_b",
      SEGMENT_NA == "Alligator Bayou" ~ "sab_b"
       )
  )

st_area(bbl_frame3)

bbl_frame3$area_m2 <- st_area(bbl_frame3)

bbl_frame3 %>% 
  group_by(wbody_type) %>% 
  summarize(., count = dplyr::n()) %>% 
  print(n = Inf)

sf_bbl1 <- bbl_frame3

# Writing a single shapefile to the working directory
st_write(sf_bbl1,
dsn = "output_sampleframes_surveys/sf_bbl1.shp",
driver = "ESRI Shapefile",
append = FALSE)

```

## Polygon Edits for BBL_7

In West Bay, the Johnson Bayou WestBay should be corrected to have Maj_Seg be West Bay instead of St. Andrew Bay. Should Deer Point Lake be edited to label polygons for Williams Bayou, Bayou George, Bear Creek and Cedar Creek? This would create a segment fore Deer Point Lake and corresponding bayous analogous to rest of the bay segments. In St. Andrew Bay, why does Alligator Bayou have Maj_Seg assignment of Lake and not St. Andrew Bay?
