---
output: html_document
editor_options: 
  chunk_output_type: console
---
## Libraries

```{r libraries}
library(tidyverse)
library(sf)
library(mapview)
library(janitor)
library(leaflet)
library(leafpop) # for popups in 
library(leafsync) # sync mapview maps
library(patchwork)
library(nhdplusTools)
library(tmap)
library(forcats)
library(gifski) #for tmap animation
```

## Log

2025-12-27 read in win_baywatch_sabnut.csv and using code and data frame names from script_01_esda.qmd

## Reading in Data and Formatting Date Variables

One csv file, win_baywatch_sabnut.csv, uploaded by Ryann Rossi, SASJBEP, on December 23, 2025. Ryan pulled the data from WIN and did some data cleaning. Note activity_start_date comes in as character.

Read in Bays_Bayous_Lakes_V1 shapefile that Rose uploaded to repository.

```{r data}
np_ds1 <- read_csv("data/win_baywatch_sabnut.csv")
names(np_ds1)
np_ds1 <- clean_names(np_ds1)
# spec(np_ds1)
# names(np_ds1)


# See Long and Lat in decimal degrees and date as character
View(np_ds1) 
str(np_ds1)
glimpse(np_ds1)
class(np_ds1$date)

#zdates <- distinct(np_ds1, activity_start_date)
#ym_ds1 <- lubridate::mdy(zdates$activity_start_date)
#floor_date(ym_ds1, "month")
#saveRDS(zdates, "zdates")

# with activity start date still character have to assign factors
# have to create activity start date as factor for tmap animation
np_ds1$date_factor <- np_ds1$date
np_ds1$yrmm_factor <- np_ds1$date

# convert to date
np_ds1$date <- lubridate::mdy(np_ds1$date)
class(np_ds1$date)
np_ds1$yrmm <- floor_date(np_ds1$date, "month")
head(np_ds1$yrmm)
class(np_ds1$yrmm)

# have to create activity start date as factor for tmap animation
np_ds1 <- arrange(np_ds1, date)
np_ds1$date_factor <- as_factor(np_ds1$date_factor)
head(np_ds1$date_factor)
levels(np_ds1$date_factor)


np_ds1$yrmm_factor <- as_factor(as.character(np_ds1$yrmm))
head(np_ds1$yrmm_factor)
levels(np_ds1$yrmm_factor)

# filter(dg_sf1, (!is.na(log_n2o))

np_ds1 <- filter(np_ds1, (!is.na(long)))
# drops to 10,369 obs and summary(mp_ds2$lat) returns no NA

```

## Create sf object
Have NA for some lat and long entries so remove these
```{r create_sf}

# create sf object
coords <- data.frame(longitude  = np_ds1$long, latitude = np_ds1$lat)
head(coords)

# See Marc Weber workshop above about Coordinate Reference Systems (CRS)

np_sf1 <- st_as_sf(np_ds1, coords = c("long", "lat"), crs =4269) %>% st_transform((5070))

class(np_sf1)
# read in shapefile
nnc_1 <- st_read("data_gis/NNC_Bay_Segments.shp")
class(nnc_1)
summary(nnc_1)
# see that espg code is 3087
# asking do the two sf objects have the same CRS
st_crs(np_sf1) == st_crs(nnc_1)

np_sf1 <- np_sf1 %>% st_transform(3087)
st_crs(np_sf1) == st_crs(nnc_1)

mapview(np_sf1) + mapview(nnc_1)

# mapview(hab_sf1, zcol = "Depth", cex = "Depth", alpha.regions = .8, legend = TRUE, popup = popupTable(hab_sf1, zcol = c("Depth", "Site"))) + mapview(bigbend_seagrass)

bbl_1 <- st_read("data_gis/Bays_Bayous_Lakes_V1.shp")
class(bbl_1)
summary(bbl_1)

bbl_1 <- bbl_1 %>% st_transform(3087)

mapview(bbl_1)

mapview(np_sf1) + mapview(bbl_1)

```

## Data Structure and Summaries

Six distinct organizations collected monitoring data with Florida Lakewatch being a synonym for St. Andrew Baywatch. An NA organization is listed and missing is NWFWMD. Need to compare Ryann and Rose's queries. Get 1020 dates spanning 1998-07-10 to 2025-04-23. Have 130 distinct monitoring location IDs, including NA. Have 210 distinct monitoring location names. There are 7894 distinct coordinates/locations.

But, I am wondering if DEP assigns coordinates that are not the actual locations of the monitoring station, but instead are always to same place (the centroid) of the spatial feature. Also, are there any analyte code remarks about questionable values?

```{r data_structure}
names(np_ds1)
distinct(np_ds1, organization_name)

distinct(np_ds1, date) %>% arrange(., date) %>% 
  print(n = Inf)

np_ds1 %>% 
  group_by(organization_name) %>% 
  summarise(count = n_distinct(monitoring_location_name))

distinct(np_ds1, monitoring_location_id) %>% arrange(., monitoring_location_id) %>% 
  print(n = Inf)

distinct(np_ds1, monitoring_location_name) %>% arrange(., monitoring_location_name) %>% 
  print(n = Inf)

distinct(np_ds1, chl) %>% arrange(., chl) %>% print(n = Inf)
distinct(np_ds1, tn) %>% arrange(., tn) %>% print(n = Inf)
distinct(np_ds1, tp) %>% arrange(., tp) %>% print(n = Inf)

# distinct monitoring stations
np_stations <- unique.data.frame(np_sf1) # returns 10,367 obs/stations & geometries

np_stations %>% 
  group_by(organization_name) %>% 
  summarise(count = n_distinct(monitoring_location_name))


```

Interesting that total phosphorus has monitoring names that include format of DEP's status survey of zone 1 (Z1) for NWFMD region followed by resource type large lake (LL), but that's not the case for total nitrogen.

```{r summaries}

tn_ds1 <- filter(np_ds1, dep_analyte_name == "Nitrogen- Total")



summarize(tn_ds1, min = fivenum(dep_result_value_number)[1],
              q1 = fivenum(dep_result_value_number)[2],
              median = fivenum(dep_result_value_number)[3],
              q3 = fivenum(dep_result_value_number)[4],
              max = fivenum(dep_result_value_number)[5],
              count = dplyr::n(),
              mean = mean(dep_result_value_number))
  
tn_sum_monitoring_name <- tn_ds1 %>% 
  group_by(monitoring_location_name) %>% 
  summarize(min = fivenum(dep_result_value_number)[1],
              q1 = fivenum(dep_result_value_number)[2],
              median = fivenum(dep_result_value_number)[3],
              q3 = fivenum(dep_result_value_number)[4],
              max = fivenum(dep_result_value_number)[5],
              count = dplyr::n(),
              mean = mean(dep_result_value_number))


tp_ds1 <- filter(np_ds1, dep_analyte_name == "Phosphorus- Total")

summarize(tp_ds1, min = fivenum(dep_result_value_number)[1],
              q1 = fivenum(dep_result_value_number)[2],
              median = fivenum(dep_result_value_number)[3],
              q3 = fivenum(dep_result_value_number)[4],
              max = fivenum(dep_result_value_number)[5],
              count = dplyr::n(),
              mean = mean(dep_result_value_number))
  
tp_sum_monitoring_name <- tp_ds1 %>% 
  group_by(monitoring_location_name) %>% 
  summarize(min = fivenum(dep_result_value_number)[1],
              q1 = fivenum(dep_result_value_number)[2],
              median = fivenum(dep_result_value_number)[3],
              q3 = fivenum(dep_result_value_number)[4],
              max = fivenum(dep_result_value_number)[5],
              count = dplyr::n(),
              mean = mean(dep_result_value_number))

    iris %>%
          group_by(Species) %>%
          summarize(min = fivenum(Petal.Width)[1],
                    low = fivenum(Petal.Width)[2],
                    median = fivenum(Petal.Width)[3],
                    upper = fivenum(Petal.Width)[4],
                    max = fivenum(Petal.Width)[5],
                    count = dplyr::n())

```

## Spatial Intersect of Sub-bays

Does spatial intersection result in, for example, of upper Crooked Creek being excluded?

```{r st_intersect}
# four sub_bays
subbay_1 <- dplyr::filter(bbl_1, SEGMENT_NA == "West Bay" | SEGMENT_NA == "North Bay" | SEGMENT_NA == "St. Andrew Bay" | SEGMENT_NA == "East Bay")

mapview(subbay_1)
sub_np1 <- st_intersection(np_sf1, subbay_1)
head(sub_np1)


mapview(sub_np1) + mapview(bbl_1)
mapview(sub_np1) + mapview(subbay_1)

```
