---
output: html_document
editor_options: 
  chunk_output_type: console
---
## Libraries

```{r libraries}
library(tidyverse)
library(sf)
library(mapview)
library(janitor)
library(leaflet)
library(leafpop) # for popups in 
library(leafsync) # sync mapview maps
library(leafem) # add labels to map
library(patchwork)
library(nhdplusTools)
library(tmap)
library(forcats)
library(gifski) #for tmap animation
library(plotly)
library(webshot) # save mapview object
```

## Log

2026-03-05 try merging np_ds1 and bw_id1 by long/lat as data frames then make into sf object seeing as st_join, st_intersection, etc. are not working

2026-02-03 got synced mapview maps of old and recent Lakewatch stations

2026-02-01 looked at FDEP WATERSHED ASSESSMENT SECTION data

2026-01-31 Rose Horn uploaded Bays_Bayous_Lakes_V2.zip to data_gis

2026-01-12 exploring stations and organizations that have sta_nr assigned as NA

2026-01-22 sta_nr assignments seem to indicate Baywatch stations, but also have 1491 obs with NA assigned for sta_nr

2025-12-27 read in win_baywatch_sabnut.csv and using code and data frame names from script_01_esda.qmd

## Reading in Data and Formatting Date Variables

One csv file, win_baywatch_sabnut.csv, uploaded by Ryann Rossi, SASJBEP, on December 23, 2025. Ryan pulled the data from WIN and did some data cleaning. Note activity_start_date comes in as character.

Read in Bays_Bayous_Lakes_V1 shapefile that Rose uploaded to repository.

```{r data}
np_ds1 <- read_csv("data/win_baywatch_sabnut.csv")
names(np_ds1)
np_ds1 <- clean_names(np_ds1)
# spec(np_ds1)
# names(np_ds1)


# See Long and Lat in decimal degrees and date as character
View(np_ds1) 
str(np_ds1)
glimpse(np_ds1)
class(np_ds1$date)

#zdates <- distinct(np_ds1, activity_start_date)
#ym_ds1 <- lubridate::mdy(zdates$activity_start_date)
#floor_date(ym_ds1, "month")
#saveRDS(zdates, "zdates")

# with activity start date still character have to assign factors
# have to create activity start date as factor for tmap animation
np_ds1$date_factor <- np_ds1$date
np_ds1$yrmm_factor <- np_ds1$date

# convert to date
np_ds1$date <- lubridate::mdy(np_ds1$date)
class(np_ds1$date)
np_ds1$yrmm <- floor_date(np_ds1$date, "month")
head(np_ds1$yrmm)
class(np_ds1$yrmm)

# have to create activity start date as factor for tmap animation
np_ds1 <- arrange(np_ds1, date)
np_ds1$date_factor <- as_factor(np_ds1$date_factor)
head(np_ds1$date_factor)
levels(np_ds1$date_factor)


np_ds1$yrmm_factor <- as_factor(as.character(np_ds1$yrmm))
head(np_ds1$yrmm_factor)
levels(np_ds1$yrmm_factor)

# filter(dg_sf1, (!is.na(log_n2o))

np_ds1 <- filter(np_ds1, (!is.na(long)))
# drops to 10,369 obs and summary(mp_ds2$lat) returns no NA

```

## Explore Duplicates and NAs
```{r dupes_na}
# z1 <- get_dupes(np_ds1)
# 
# z2 <- filter(np_ds1, (is.na(tn) & !is.na(tp)))
# 
# z3 <- filter(np_ds1, is.na(tn))
```


## Create sf object
Have NA for some lat and long entries so remove these
```{r create_sf}

# create sf object
coords <- data.frame(longitude  = np_ds1$long, latitude = np_ds1$lat)
head(coords)

# See Marc Weber workshop above about Coordinate Reference Systems (CRS)

np_sf1 <- st_as_sf(np_ds1, coords = c("long", "lat"), crs =4269) %>% st_transform((5070))

class(np_sf1)
# read in shapefile
nnc_1 <- st_read("data_gis/NNC_Bay_Segments.shp")
class(nnc_1)
summary(nnc_1)
# see that espg code is 3087
# asking do the two sf objects have the same CRS
st_crs(np_sf1) == st_crs(nnc_1)

np_sf1 <- np_sf1 %>% st_transform(3087)
st_crs(np_sf1) == st_crs(nnc_1)

mapview(np_sf1) + mapview(nnc_1)

# mapview(hab_sf1, zcol = "Depth", cex = "Depth", alpha.regions = .8, legend = TRUE, popup = popupTable(hab_sf1, zcol = c("Depth", "Site"))) + mapview(bigbend_seagrass)

bbl_1 <- st_read("data_gis/Bays_Bayous_Lakes_V1.shp")
class(bbl_1)
summary(bbl_1)

bbl_1 <- bbl_1 %>% st_transform(3087)

mapview(bbl_1)

mapview(np_sf1) + mapview(bbl_1)

```

## Read Bays, Bayous and Lakes V2
Rose Horn took 2 polygons for Alligator Bayou, which borders the U.S. Coast Guard Station on St. Andrew Bay, and splits them into separate sampling areas. Will stick with using version 1 for now.
```{r bbl_v2}
# 
# bbl_2 <- st_read("data_gis/Bays_Bayous_Lakes_V2/Bays_Bayous_Lakes_V2.shp")
# 
# mapview(bbl_2)



```


## Data Structure and Summaries

Six distinct organizations collected monitoring data with Florida Lakewatch being a synonym for St. Andrew Baywatch. An NA organization is listed and missing is NWFWMD. Need to compare Ryann and Rose's queries. Get 1020 dates spanning 1998-07-10 to 2025-04-23. Have 130 distinct monitoring location IDs, including NA. Have 210 distinct monitoring location names. There are 7894 distinct coordinates/locations.

But, I am wondering if DEP assigns coordinates that are not the actual locations of the monitoring station, but instead are always to same place (the centroid) of the spatial feature. Also, are there any analyte code remarks about questionable values?

```{r data_structure}
names(np_ds1)
distinct(np_ds1, organization_name)

distinct(np_ds1, sampling_agency_name) # shows ST. ANDREW BAY RESOURCE MANAGEMENT ASSOCIATION INC.

distinct(np_ds1, date) %>% arrange(., date) %>% 
  print(n = Inf)

np_ds1 %>% 
  group_by(organization_name) %>% 
  summarise(count = n_distinct(monitoring_location_name))

distinct(np_ds1, monitoring_location_id) %>% arrange(., monitoring_location_id) %>% 
  print(n = Inf)

distinct(np_ds1, monitoring_location_name) %>% arrange(., monitoring_location_name) %>% 
  print(n = Inf)

distinct(np_ds1, chl) %>% arrange(., chl) %>% print(n = Inf)
distinct(np_ds1, tn) %>% arrange(., tn) %>% print(n = Inf)
distinct(np_ds1, tp) %>% arrange(., tp) %>% print(n = Inf)

# distinct monitoring stations
np_stations <- unique.data.frame(np_sf1) # returns 10,367 obs/stations & geometries

np_stations %>% 
  group_by(organization_name) %>% 
  summarise(count = n_distinct(monitoring_location_name))

# distinct bay
distinct(np_ds1, bay) %>% print(n = Inf)

np_ds1 %>% 
  group_by(bay) %>% 
  summarise(count = n_distinct(monitoring_location_name)) %>% 
  print(n = Inf)

```

Interesting that total phosphorus has monitoring names that include format of DEP's status survey of zone 1 (Z1) for NWFMD region followed by resource type large lake (LL), but that's not the case for total nitrogen.

```{r summaries}
# tn
dplyr::select(np_ds1, tn) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn)))

dplyr::select(np_ds1, tn, bay) %>% 
group_by(., bay) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)

# tp
dplyr::select(np_ds1, tp) %>% 
summarize(., min = fivenum(tp)[1],
              q1 = fivenum(tp)[2],
              median = fivenum(tp)[3],
              q3 = fivenum(tp)[4],
              max = fivenum(tp)[5],
              count = dplyr::n(),
              mean = mean(tp, na.rm = TRUE),
              na_count = sum(is.na(tp)))

dplyr::select(np_ds1, tp, bay) %>% 
group_by(., bay) %>% 
summarize(., min = fivenum(tp)[1],
              q1 = fivenum(tp)[2],
              median = fivenum(tp)[3],
              q3 = fivenum(tp)[4],
              max = fivenum(tp)[5],
              count = dplyr::n(),
              mean = mean(tp, na.rm = TRUE),
              na_count = sum(is.na(tp))) %>% 
print(n = Inf)

# chl
dplyr::select(np_ds1, chl) %>% 
summarize(., min = fivenum(chl)[1],
              q1 = fivenum(chl)[2],
              median = fivenum(chl)[3],
              q3 = fivenum(chl)[4],
              max = fivenum(chl)[5],
              count = dplyr::n(),
              mean = mean(chl, na.rm = TRUE),
              na_count = sum(is.na(chl)))

dplyr::select(np_ds1, chl, bay) %>% 
group_by(., bay) %>% 
summarize(., min = fivenum(chl)[1],
              q1 = fivenum(chl)[2],
              median = fivenum(chl)[3],
              q3 = fivenum(chl)[4],
              max = fivenum(chl)[5],
              count = dplyr::n(),
              mean = mean(chl, na.rm = TRUE),
              na_count = sum(is.na(chl))) %>% 
print(n = Inf)
  
```

## Spatial Intersect of Sub-bays

Does spatial intersection result in, for example, of upper Crooked Creek being excluded? Yes, it does. Do Rose and Ryann needed to include, possibly, NHDPlus polygons of Crooked Creek and Burnt Mill Creek so as not to "lose" stations in those locations?
Spatial intersection with sub-bay drops obs to 5737 because excluding coastal dune lakes and areas outside of St. Andrew Bay.
```{r st_intersect}
# four sub_bays
subbay_1 <- dplyr::filter(bbl_1, SEGMENT_NA == "West Bay" | SEGMENT_NA == "North Bay" | SEGMENT_NA == "St. Andrew Bay" | SEGMENT_NA == "East Bay")

mapview(subbay_1)
sub_np1 <- st_intersection(np_sf1, subbay_1)
head(sub_np1)


mapview(sub_np1) + mapview(bbl_1)
mapview(sub_np1) + mapview(subbay_1)

```

## Sub-bay Data Structure
Note sector does not always match SEGMENT_NA
Group by sta_nr return 4246 obs out of 5737 as having a sta_nr assignment so I believe those are Baywatch stations. Have 1491 obs with NA station were they collected by other organizations? Why does creating a unique simple features data frame drop observations down to 5736?
```{r subbay_data_struct}
names(sub_np1)

distinct(sub_np1, organization_id)
distinct(sub_np1, organization_name) # 4 organizations plus NA

distinct(sub_np1, sampling_agency_name)

distinct(sub_np1, date) %>% arrange(., date) %>% 
  print(n = Inf) #660 dates from 1990-02-11 to 2024-12-17

sub_np1 %>% 
  group_by(organization_name) %>% 
  summarise(count = n_distinct(monitoring_location_name))

sub_np1 %>% 
  group_by(organization_id) %>% 
  summarise(., count = dplyr::n()) %>% 
  print(n=Inf)


sub_np1 %>% 
  group_by(organization_name) %>% 
  summarise(., count = dplyr::n()) %>% 
  print(n=Inf)

distinct(sub_np1, monitoring_location_id) %>% arrange(., monitoring_location_id) %>% 
  print(n = Inf)

distinct(sub_np1, monitoring_location_name) %>% arrange(., monitoring_location_name) %>%   print(n = Inf)

distinct(sub_np1, bay) %>% arrange(., bay) %>%   print(n = Inf)
# return 15 bays

distinct(sub_np1, sta_nr) %>% arrange(., sta_nr) %>% print(n=Inf)
# returns 29 distinct numeric codes plus an NA
# assumption is that if have numeric code for sta_nr then station is 
# a Baywatch station based on RMA_Baywatch_Active_Stations_Revised_112015.pdf
# even though organization ID, name, and sampling agency fields are blank

group_by(sub_np1, sta_nr) %>% 
  summarize(., count = dplyr::n()) %>% 
  arrange(., sta_nr) %>% 
  print(n=Inf)

  
# nd_ds1 <- group_by(sub_np_stations, year) %>% 
#   summarize(.,  count = dplyr::n(),
#             na_count = sum(is.na(tn))) %>%
#   mutate(., pct_na = (na_count/count)*100) %>% 
#   print(n = Inf)

# distinct(sub_np1, dep_analyte_name)

# distinct monitoring stations
# drops by 1 from sub_np1 I think b/c of duplicate
sub_np_stations <- unique.data.frame(sub_np1) # returns 5736 obs/stations & geometries

sub_np_stations %>% 
  group_by(organization_name) %>% 
  summarise(count = n_distinct(monitoring_location_name))

distinct(sub_np_stations, sector)


sub_np_stations %>% 
  group_by(sta_nr) %>% 
  summarise(count = n_distinct(sta_nr)) %>% 
  print(n=Inf)


sub_np_stations %>% 
  group_by(sta_nr) %>% 
  summarise(., count = dplyr::n()) %>% 
  print(n=Inf)
```


## Facet Map by Organization
This can possible identify which organizations have unique station locations and which organizations have overlapping locations. A lot of overlap between Florida Lakewatch and Missing organization name.
```{r facetp_map_org}

tm_shape(subbay_1) +
  tm_polygons(fill = "white", lwd = 2) +
  tm_title("Detectable and NA Entries of Total Nitrogen (mg/L)")  + 
  tm_shape(sub_np_stations) +
  tm_symbols(fill = "tn") +
  tm_facets(by = "organization_name")

```

## na_sta_nr Station Data
Can I identify unique stations identifiers for these observations where sta_nr is NA? Have 1,491 observations with sta_nr being NA. One concern is how to check that 861 Lakewatch observations are not duplicates of entries where sta_nr had a numerical coded entry. Another concern is how to know if a tn, tp, or chl entry is NA does that indicate a nondetect or no sample taken?
```{r na_sta_nr}
filter(sub_np_stations, is.na(sta_nr)) %>% 
  distinct(., organization_name)

na_sta_nr1 <- filter(sub_np_stations, is.na(sta_nr))


# FDEP NORTHWEST REGIONAL OPERATIONS CENTER       5
# FDEP TALLAHASSEE REGIONAL OPERATIONS CENTER    15
# FDEP WATERSHED ASSESSMENT SECTION             610
# FLORIDA LAKEWATCH                             861 

na_sta_nr1 %>% 
  group_by(organization_name) %>% 
  summarise(., count = dplyr::n()) %>% 
  print(n=Inf)

filter(sub_np_stations, !is.na(sta_nr)) %>%
         distinct(., organization_name)

filter(sub_np_stations, !is.na(sta_nr)) %>%
         group_by(., organization_name)  %>% 
  summarise(., count = dplyr::n()) %>% 
  print(n=Inf)

filter(sub_np_stations, organization_name == "FLORIDA LAKEWATCH") %>%
         group_by(., sta_nr)  %>% 
  summarise(., count = dplyr::n()) %>% 
  print(n=Inf)


fdep_nw1 <- filter(sub_np_stations, organization_name == "FDEP NORTHWEST REGIONAL OPERATIONS CENTER")

mapview(fdep_nw1)
# tn is NA for all 5 observations, but tp and chl detectable concentrations
# two East Bay stations, one near southeast of downtown PC, the other near
# Dupont Bridge, each with an April 30, 2019 and December 7, 2020 collection.
# One West bay station sampled on June 6, 2018.

fdep_tlh1 <- filter(sub_np_stations, organization_name == "FDEP TALLAHASSEE REGIONAL OPERATIONS CENTER")

mapview(fdep_tlh1)
# tn is NA for all 15 observations, but tp and chl detectable concentrations
# One East Bay station at Sandy Creek, and five observations
# One North Bay station at Fannin Bayou, and five observations
# Another North Bay station near Lynn Haven Golf course with five observations
# Samples collected from 2021-2024

fdep_was1 <- filter(sub_np_stations, organization_name == "FDEP WATERSHED ASSESSMENT SECTION")
# 610 observations with sampling agency being FDEP Central Panhandle Aquatic Preserves and believe monitoring locations confined to aquatic preserve boundaries on either side of the pass. There are 11 distinct monitoring location names. 84 distinct dates, monthly, starting 2016-07-20, skip to 2017-07-10 through 2024-12-17. Only 4 stations span the full time period, which include Hurricane Michael

summary(fdep_was1$tp)
summary(fdep_was1$chl)
summary(fdep_was1$tn)

mapview(fdep_was1)

distinct(fdep_was1, sampling_agency_name)
distinct(fdep_was1, monitoring_location_name)
distinct(fdep_was1, monitoring_location_id)
distinct(fdep_was1, date) %>% 
  arrange(.) %>% 
  print(n = Inf)

ggplot(fdep_was1, aes(x = date, y = tp)) + geom_point() + 
  facet_wrap(vars(monitoring_location_name))

ggplot(fdep_was1, aes(x = date, y = tp)) + geom_point() + 
  facet_wrap(vars(monitoring_location_name), scales = "free_y")

ggplot(fdep_was1, aes(x = date, y = chl)) + geom_point() + 
  facet_wrap(vars(monitoring_location_name))

ggplot(fdep_was1, aes(x = date, y = chl)) + geom_point() + 
  facet_wrap(vars(monitoring_location_name), scales = "free_y")


fl_lw1 <- filter(sub_np_stations, organization_name == "FLORIDA LAKEWATCH")
# note sta_nr entries are NA for these 610 observations. But, do their coordinates put them on top of the other Lakewatch stations? Have 37 distinct station IDs. Dates spanning from 2018-12-07 to 2024-04-29

summary(fl_lw1$tp)
summary(fl_lw1$chl)
summary(fl_lw1$tn)

mapview(fl_lw1)

distinct(fl_lw1, sampling_agency_name)
distinct(fl_lw1, monitoring_location_name) %>%  print(n = Inf)
distinct(fl_lw1, monitoring_location_id) %>%  print(n = Inf)
distinct(fl_lw1, date) %>% 
  arrange(.) %>% 
  print(n = Inf)


ggplot(fl_lw1, aes(x = date, y = tp)) + geom_point() + 
  facet_wrap(vars(monitoring_location_name))

ggplot(fl_lw1, aes(x = date, y = tp)) + geom_point() + 
  facet_wrap(vars(monitoring_location_name), scales = "free_y")

ggplot(fl_lw1, aes(x = date, y = tn)) + geom_point() + 
  facet_wrap(vars(monitoring_location_name))

ggplot(fl_lw1, aes(x = date, y = tn)) + geom_point() + 
  facet_wrap(vars(monitoring_location_name), scales = "free_y")

ggplot(fl_lw1, aes(x = date, y = chl)) + geom_point() + 
  facet_wrap(vars(monitoring_location_name))

ggplot(fl_lw1, aes(x = date, y = chl)) + geom_point() + 
  facet_wrap(vars(monitoring_location_name), scales = "free_y")


```

## Compare sta_nr coded vs Lakewatch with sta_nr as NA
```{r compare}

sta_nr_coded1 <- filter(sub_np_stations, !is.na(sta_nr))

summary(sta_nr_coded1$date)
# older dates 1990-02-11 to 2006-12-21 with sta_nr numeric coded

summary(fl_lw1$date)
# more recent dates 2018-12-07 to 2024-04-09 with sta_nr coded NA

mapview(sta_nr_coded1)

mapview(fl_lw1) 

mapview(fl_lw1, color = "magenta") + mapview(sta_nr_coded1, color = "green")


# tm_1 <- tm_shape(subbay_1) +
#   tm_polygons(fill = "white", lwd = 2) +
#   tm_shape(sta_nr_coded1) +
#   tm_symbols(size = 1, fill = "magenta")
# 
# 
# tm_2 <- tm_shape(subbay_1) +
#   tm_polygons(fill = "white", lwd = 2) +
#   tm_shape(fl_lw1) +
#   tm_symbols(size = 0.5, fill = "black")

# mapview(DFobs3a, zcol = "vsci", at=seq(0, 90, 15), alpha.regions = .8, legend = TRUE,layer.name = 'Observed Sites VSCI', popup = popupTable(aug_d_predict1ws_vsci, zcol = c("station_id_2", "vsci"))) + mapview(vahusb_poly, zcol="subbasin")


# mapview(sites_stratv2) %>%
#   addStaticLabels(label = sites_stratv2$abbreviation,
#                   noHide = TRUE,
#                   direction = 'top',
#                   textOnly = TRUE,
#                   textsize = "20px")



sta_nr_labels <- mapview(sta_nr_coded1) %>% 
  addStaticLabels(label = sta_nr_coded1$sta_nr,
                  noHide = TRUE,
                  direction = 'top',
                  textOnly = TRUE,
                  textsize = "20px") 


# https://stackoverflow.com/questions/31336898/how-to-save-leaflet-in-r-map-as-png-or-jpg-file
# mapshot(sta_nr_labels, file =" ~/output_maps/sta_nr_labels.png")
# getting error about install_phantomjs

# in Viewer selected Export as webpage and put in Baywatch_documents as such a large file


tm_1 <- mapview(fl_lw1, color = "magenta") + mapview(subbay_1)
tm_2 <- mapview(sta_nr_coded1, color = "green", popup =
                  popupTable(sta_nr_coded1, zcol = "sta_nr")) + mapview(subbay_1)

sync(tm_1, tm_2, ncol=1)
```

## Baywatch Station IDs
Baywatch Station IDs provided by Ryann Rossi.
```{r baywatch_ids}

bw_id1 <- read_csv("C:/Users/atweg/Documents/Mike/SASJBEP/Baywatch_documents/baywatch_station_ids.csv")
bw_kd1 <- clean_names(bw_id1)

glimpse(bw_id1)

# create sf object
coords <- data.frame(longitude  = bw_id1$Long, latitude = bw_id1$Lat)
head(coords)

# See Marc Weber workshop above about Coordinate Reference Systems (CRS)

bw_sf1 <- st_as_sf(bw_id1, coords = c("Long", "Lat"), crs =4269) %>% st_transform((5070))

mapview(bw_sf1)

bw_sf1 <- bw_sf1 %>% st_transform(3087)
st_crs(bw_sf1)
st_crs(fl_lw1)

bw_m1 <- mapview(bw_sf1, popup =
                  popupTable(bw_sf1, zcol = "Station_Number"))


lw_m1 <- mapview(fl_lw1, popup =
                  popupTable(fl_lw1, zcol = "sta_nr"))

sync(bw_m1, lw_m1, ncol = 1)

distinct(fl_lw1, monitoring_location_name) %>% 
  arrange(.) %>% 
  print(n=Inf)
# returns 37 location names

bw_sf2 <- st_join(fl_lw1, bw_sf1, join = st_covers)

names(bw_sf2)
head(bw_sf2)
```


## Subbay Stations
```{r subbay_tn}

# tn
dplyr::select(sub_np_stations, tn) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn)))

dplyr::select(sub_np_stations, tn, SEGMENT_NA) %>% 
group_by(., SEGMENT_NA) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)

mapview(sub_np_stations)

sub_np_stations %>% 
  group_by(SEGMENT_NA) %>% 
  summarise(count = n_distinct(u_of_f_number))

# tn_sub1 <- filter(sub_np1, dep_analyte_name == "Nitrogen- Total") 
# tn data is only from Florida Lakewatch

# distinct(tn_sub1, organization_name)

# distinct tn monitoring stations
# tn_stations <- unique.data.frame(tn_sub1)

# mapview(tn_sub1)

# mapview(tn_stations)
```

## North Bay
Believe it has smallest n of stations based on SEGEMENT_NA
Does not appear to be unique station identifier as have NA entries for U of F number, in addition to North Bay-1, North Bay-2, North Bay-3, and Fannin Bayou-1
No samples collected in 2006, though station coordinated entered. No samples collected from 2007-2017 and no coordinates entered.
```{r northbay}
nb_ds1 <- dplyr::filter(sub_np_stations, SEGMENT_NA == "North Bay")

distinct(nb_ds1, u_of_f_number)
distinct(nb_ds1, monitoring_location_name)
distinct(nb_ds1, sta_nr)
# does 1.1, 1.2,1.3, and 1.7 corresponds to North Bay 1, 2, 3 and Fannin Bayou?


dplyr::select(nb_ds1, tn, monitoring_location_name) %>% 
group_by(., monitoring_location_name) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)

# why 620 observations not assigned a monitoring location name

dplyr::select(nb_ds1, tn, sta_nr) %>% 
group_by(., sta_nr) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)
# now returns only 178 with sta_nr NA

dplyr::select(nb_ds1, tn, year) %>% 
group_by(., year) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)

dplyr::select(nb_ds1, tn, date_factor) %>% 
group_by(., date_factor) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)


mapview(nb_ds1)
# why does map show 8 stations

nb_ds2 <- dplyr::select(nb_ds1, tn, monitoring_location_name, geometry)

nb_geom <- dplyr::select(nb_ds1, geometry)

distinct(nb_geom)
mapview(nb_geom)
```

## tmap North Bay
Does tn is NA means no sample taken or sample taken by concentration was not detectable? Filter to just year 1999 gives 43 observations, but with some tn being NA. Filter 1999 and (!(is.na(tn))) gives 12 observations. tmap does not like missing values so no longer have Fannin Bayou. Was Fannin Bayou not sampled or was it sampled and had non-detecable concentration. Use tm_symbols(fill = "tn") to get nondetectable NA's mapped as missing data. Also, using date_factor showed two April samplings. No entries for 2007-2017 so no map?
```{r tmap_nb}

distinct(nb_ds1, year) %>% print(n=Inf)

# (dg_sf1, (!is.na(log_n2o))
nb_ds3 <- filter(nb_ds1, year == "2024")
# nb_ds4 <- filter(nb_ds1, year == "1990" & (!(is.na(tn))))

tmap_mode("plot")
tmap_mode("view")
tn_map_1990_v1 <- tm_shape(subbay_1) +
  tm_polygons(fill = "white", lwd = 2) +
  tm_title("Detectable and Nondetectable Total Nitrogen (mg/L) in North Bay, 2024")  + 
  tm_shape(nb_ds3) +
  tm_symbols(fill = "tn") +
  tm_facets(by = "date_factor")

tn_map_1990_v1

tmap_mode("plot")
tn_bubble <- tm_shape(subbay_1) +
  tm_fill("#dddddd") + 
tm_shape(nb_ds3) +
  tm_symbols(fill =  "tn",
             # size.scale = tm_scale_continuous_log(values.scale = "dep_result_value_number"),
             size.legend = tm_legend("TN (mg/L)"))

#daily loop
tn_bubble + tm_animate(frame = "date_factor", fps = 1, play = "loop")

# monthly loop
tn_bubble + tm_animate(frame = "yrmm_factor", fps = 1, play = "loop")



  tm_compass(position = c("LEFT", "BOTTOM")) +
  tm_scalebar(position = c("LEFT", "BOTTOM")) +
```

## West Bay
Filtering misses most upstream Crooked Creek and Burnt Mill Creek stations.
```{r westbay}
wb_ds1 <- dplyr::filter(sub_np_stations, SEGMENT_NA == "West Bay")

distinct(wb_ds1, u_of_f_number)
distinct(wb_ds1, monitoring_location_name)
distinct(wb_ds1, sta_nr)
# does 1.1, 1.2,1.3, and 1.7 corresponds to North Bay 1, 2, 3 and Fannin Bayou?

dplyr::select(wb_ds1, tn, monitoring_location_name) %>% 
group_by(., monitoring_location_name) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)

# why 575 observations not assigned a monitoring location name

dplyr::select(wb_ds1, tn, sta_nr) %>% 
group_by(., sta_nr) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)
# now returns only 1555 with sta_nr NA

dplyr::select(wb_ds1, tn, year) %>% 
group_by(., year) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)

dplyr::select(wb_ds1, tn, date_factor) %>% 
group_by(., date_factor) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)


mapview(wb_ds1)
# why does map show 8 stations

wb_ds2 <- dplyr::select(wb_ds1, tn, monitoring_location_name, geometry)

wb_geom <- dplyr::select(wb_ds1, geometry)

distinct(wb_geom)
mapview(wb_geom)
```

## tmap West Bay
```{r tmap_wb}

distinct(wb_ds1, year) %>% print(n=Inf)

# (dg_sf1, (!is.na(log_n2o))
wb_ds3 <- filter(wb_ds1, year == "2024")
# wb_ds4 <- filter(wb_ds1, year == "1990" & (!(is.na(tn))))

tmap_mode("plot")
# tmap_mode("view")
tn_map_1990_v1 <- tm_shape(subbay_1) +
  tm_polygons(fill = "white", lwd = 2) +
  tm_title("Detectable and Nondetectable Total Nitrogen (mg/L) in West Bay, 2024")  + 
  tm_shape(wb_ds3) +
  tm_symbols(fill = "tn") +
  tm_facets(by = "date_factor")

tn_map_1990_v1

tmap_mode("plot")
tn_bubble <- tm_shape(subbay_1) +
  tm_fill("#dddddd") + 
tm_shape(wb_ds3) +
  tm_symbols(fill =  "tn",
             # size.scale = tm_scale_continuous_log(values.scale = "dep_result_value_number"),
             size.legend = tm_legend("TN (mg/L)"))

#daily loop
tn_bubble + tm_animate(frame = "date_factor", fps = 1, play = "loop")

# monthly loop
tn_bubble + tm_animate(frame = "yrmm_factor", fps = 1, play = "loop")



  tm_compass(position = c("LEFT", "BOTTOM")) +
  tm_scalebar(position = c("LEFT", "BOTTOM")) +
```

## East Bay
```{r westbay}
eb_ds1 <- dplyr::filter(sub_np_stations, SEGMENT_NA == "East Bay")

distinct(eb_ds1, u_of_f_number)
distinct(eb_ds1, monitoring_location_name)
distinct(eb_ds1, sta_nr)
# does 1.1, 1.2,1.3, and 1.7 corresponds to North Bay 1, 2, 3 and Fannin Bayou?

dplyr::select(eb_ds1, tn, monitoring_location_name) %>% 
group_by(., monitoring_location_name) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)

# why 575 observations not assigned a monitoring location name

dplyr::select(eb_ds1, tn, sta_nr) %>% 
group_by(., sta_nr) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)
# now returns only 1555 with sta_nr NA

dplyr::select(eb_ds1, tn, year) %>% 
group_by(., year) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)

dplyr::select(eb_ds1, tn, date_factor) %>% 
group_by(., date_factor) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)


mapview(eb_ds1)
# why does map show 8 stations

eb_ds2 <- dplyr::select(eb_ds1, tn, monitoring_location_name, geometry)

eb_geom <- dplyr::select(eb_ds1, geometry)

distinct(eb_geom)
mapview(eb_geom)
```

## tmap East Bay

```{r tmap_eb}

distinct(eb_ds1, year) %>% print(n=Inf)

# (dg_sf1, (!is.na(log_n2o))
eb_ds3 <- filter(eb_ds1, year == "2024")
# wb_ds4 <- filter(wb_ds1, year == "1990" & (!(is.na(tn))))

tmap_mode("plot")
# tmap_mode("view")
tn_map_1990_v1 <- tm_shape(subbay_1) +
  tm_polygons(fill = "white", lwd = 2) +
  tm_title("Detectable and Nondetectable Total Nitrogen (mg/L) in East Bay, 2024")  + 
  tm_shape(eb_ds3) +
  tm_symbols(fill = "tn") +
  tm_facets(by = "date_factor")

tn_map_1990_v1

tmap_mode("plot")
tn_bubble <- tm_shape(subbay_1) +
  tm_fill("#dddddd") + 
tm_shape(eb_ds3) +
  tm_symbols(fill =  "tn",
             # size.scale = tm_scale_continuous_log(values.scale = "dep_result_value_number"),
             size.legend = tm_legend("TN (mg/L)"))

#daily loop
tn_bubble + tm_animate(frame = "date_factor", fps = 1, play = "loop")

# monthly loop
tn_bubble + tm_animate(frame = "yrmm_factor", fps = 1, play = "loop")



  tm_compass(position = c("LEFT", "BOTTOM")) +
  tm_scalebar(position = c("LEFT", "BOTTOM")) +
```

## St Andrews Bay
```{r saintandrewsbay}
sb_ds1 <- dplyr::filter(sub_np_stations, SEGMENT_NA == "St. Andrew Bay")

distinct(sb_ds1, u_of_f_number)
distinct(sb_ds1, monitoring_location_name) %>% print(n = Inf)
distinct(sb_ds1, sta_nr)
# does 1.1, 1.2,1.3, and 1.7 corresponds to North Bay 1, 2, 3 and Fannin Bayou?


dplyr::select(sb_ds1, tn, monitoring_location_name) %>% 
group_by(., monitoring_location_name) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)

# why 620 observations not assigned a monitoring location name

dplyr::select(sb_ds1, tn, sta_nr) %>% 
group_by(., sta_nr) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)
# now returns only 178 with sta_nr NA

dplyr::select(sb_ds1, tn, year) %>% 
group_by(., year) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)

dplyr::select(sb_ds1, tn, date_factor) %>% 
group_by(., date_factor) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)


mapview(sb_ds1)
# why does map show 8 stations

sb_ds2 <- dplyr::select(sb_ds1, tn, monitoring_location_name, geometry)

sb_geom <- dplyr::select(sb_ds1, geometry)

distinct(sb_geom)
mapview(sb_geom)
```

##tmap St. Andrews Bay

```{r tmap_sb}

distinct(sb_ds1, year) %>% print(n=Inf)

# (dg_sf1, (!is.na(log_n2o))
sb_ds3 <- filter(sb_ds1, year == "2024")
# wb_ds4 <- filter(wb_ds1, year == "1990" & (!(is.na(tn))))

tmap_mode("plot")
# tmap_mode("view")
tn_map_1990_v1 <- tm_shape(subbay_1) +
  tm_polygons(fill = "white", lwd = 2) +
  tm_title("Detectable and Nondetectable Total Nitrogen (mg/L) in St. Andrews Bay, 2024")  + 
  tm_shape(sb_ds3) +
  tm_symbols(fill = "tn") +
  tm_facets(by = "date_factor")

tn_map_1990_v1

tmap_mode("plot")
tn_bubble <- tm_shape(subbay_1) +
  tm_fill("#dddddd") + 
tm_shape(sb_ds3) +
  tm_symbols(fill =  "tn",
             # size.scale = tm_scale_continuous_log(values.scale = "dep_result_value_number"),
             size.legend = tm_legend("TN (mg/L)"))

#daily loop
tn_bubble + tm_animate(frame = "date_factor", fps = 1, play = "loop")

# monthly loop
tn_bubble + tm_animate(frame = "yrmm_factor", fps = 1, play = "loop")



  tm_compass(position = c("LEFT", "BOTTOM")) +
  tm_scalebar(position = c("LEFT", "BOTTOM")) +
```


## Nondetects Map
```{r nondetects}

na_tn1 <- filter(sub_np_stations, is.na(tn)) # returns 2941 obs

# tn NA's occurring when tp is detected
na_tn2 <- filter(sub_np_stations, is.na(tn) & !is.na(tp))

nd_ds1 <- group_by(sub_np_stations, year) %>% 
  summarize(.,  count = dplyr::n(),
            na_count = sum(is.na(tn))) %>%
  mutate(., pct_na = (na_count/count)*100) %>% 
  print(n = Inf)

  
tm_shape(subbay_1) +
  tm_polygons(fill = "white", lwd = 2) +
  tm_title("Percentage Nondetectable Total Nitrogen in St. Andrews Bay")  + 
  tm_shape(nd_ds1) +
  tm_symbols(fill = "pct_na") +
  tm_facets(by = "year")

nd_ds2 <- sub_np_stations %>% group_by(year,SEGMENT_NA) %>% 
  summarize(.,  count = dplyr::n(),
            na_count = sum(is.na(tn))) %>%
  mutate(., pct_na = (na_count/count)*100) %>% 
  print(n = Inf)

tm_shape(subbay_1) +
  tm_polygons(fill = "white", lwd = 2) +
  tm_title("Nondetectable Total Nitrogen (mg/L) in St. Andrews Bay")  + 
  tm_shape(nd_ds2) +
  tm_symbols(fill = "pct_na") +
  tm_facets(by = "year")


ggplot(nd_ds2, aes(x = year, y = pct_na, color = SEGMENT_NA)) + geom_point()

ggplot(nd_ds2, aes(x = year, y = pct_na, color = SEGMENT_NA)) + geom_point() +
  geom_smooth() +
  facet_wrap(~SEGMENT_NA)

nd_plot <- ggplot(nd_ds2, aes(x = year, y = pct_na, color = SEGMENT_NA)) + geom_point() +
  geom_smooth() +
  facet_wrap(~SEGMENT_NA)

ggplotly(nd_plot)

nd_1990 <- filter(nd_ds, year < 2000)
```


## ECDF of Sub-Bays
```{r ecdf}

ggplot(sub_np_stations, aes(tn, color = SEGMENT_NA)) + stat_ecdf()

ggplot(sub_np_stations, aes(tn, color = SEGMENT_NA)) + stat_ecdf() + facet_wrap(vars(year))

```

