---
title: "Linked Micromap"
author: "Mike McManus"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Libraries

```{r library}
library(mapview)
library(micromap)
```

## Log

2026-04-24 linked micromap of logged chla with 2 perceptual groups

2026-04-22 pasted workshop code

## Data

Edit data, both polygons and stats, to just the areas of interest. Also, merge the statistics to the polygon data.
```{r data}
names(polygons8)
View(polygons8)

polygons6 <- filter(polygons8, SEGMENT_NA != "Crooked Creek" & SEGMENT_NA != "Burnt Mill Creek")

#chla
chla_2019v1 <- filter(chla_2019, SEGMENT_NA != "Crooked Creek")
names(chla_2019v1)

# merge polygon to chla stats
# list sf object first, then dataframe of stats to create polygons with stats
chla_2019v2 <- merge(polygons6, chla_2019v1, by = "SEGMENT_NA" )
summary(chla_2019v2)
# note that geometry is polygon

# tn
tn_2019v1 <- filter(tn_2019, SEGMENT_NA != "Crooked Creek")

chlatn_2019v2 <- merge(chla_2019v2, tn_2019v1, by = "SEGMENT_NA" )

# tp
tp_2019v1 <- filter(tp_2019, SEGMENT_NA != "Crooked Creek")

nutr_2019 <- merge(chlatn_2019v2, tp_2019v1, by = "SEGMENT_NA" )

```

## Draft Linked Micromap
Make draft linked micromaps to get code working, and note any statistical and spatial patterns.

Note cannot have a column called "median" in data. Also, for chla data needed log transformation, as did tn and tp.  Getting 3 micromaps, one uncolored, one with 6 colors, and one all gray when grouping was 6. With grouping = 3 get 2 micromaps corresponding to the 2 perceptual groups.

For chla, interesting how little variation West Bay shows. For tn, Grand Lagoon has highest median and largest variation, and West Bay has an outlier that almost exceed Grand Lagoon. For tp, Grand Lagoon shows previous pattern similar to tn. Is tp, also showing a salinity graident correlation with high tp in higher salinity waters?
```{r draft_linked_micromap}
#Draft plot
mmplot(map.data=nutr_2019,
       panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       list('chla_lmin', 'chla_lq1', 'chla_lmed', 'chla_lq3', 'chla_lmax'),
                       NA),
       grouping = 3,
       ord.by='chla_lmed', rev.ord=TRUE,
       median.row=FALSE)
       #map.link=c('SEGMENT_NA'))

mmplot(map.data=nutr_2019,
       panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       list('tn_lmin', 'tn_lq1', 'tn_lmed', 'tn_lq3', 'tn_lmax'),
                       NA),
       grouping = 3,
       ord.by='tn_lmed', rev.ord=TRUE,
       median.row=FALSE)


mmplot(map.data=nutr_2019,
       panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 3,
       ord.by='tp_lmed', rev.ord=TRUE,
       median.row=FALSE,
       print.file="mm_tp2019.jpeg",res=100)

```


## Refine Linked Micromaps
Use more detailed coding arguments to make better looking linked micromaps.
```{r refine_micromap_tp}
# TP Refined Plot
mmplot(map.data=nutr_2019,
       panel.types=c('dot_legend', 'labels', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       'tp_count',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 3,
       ord.by='tp_lmed',
       rev.ord=TRUE,
       median.row=FALSE,
       plot.pGrp.spacing=1,
       
       plot.height=6, plot.width=9, #for poster use plot.height=6, plot.width=9,
       colors=brewer.pal(5, "Spectral"), #for journal use plot.height=9, plot.width=6.8
              
       panel.att=list(list(1, panel.width=1, point.type=20, point.size=2,
                           point.border=TRUE), 
                      
                      list(2, header='St.Andrew Bay Segments', panel.width=1.2,#increased panel.width to 1.1 from .8
                           align='left', text.size=.8),
                      
                      list(3, header='Samples',
                           panel.width=.35,
                           align='right',
                           text.size=.8),
                      
                      list(4, header='2019 Detectable Total Phoshporous',
                           graph.bgcolor='lightgray',
                           xaxis.ticks=c(-6.0, -5.0, -4.0, -3.0, -2.0),
                           xaxis.labels=c(-6,-5, -4, -3, -2),
                           xaxis.labels.size=1.25,
                           graph.bar.size = 0.5,
                           #xaxis.labels.angle=90,
                           xaxis.title='Log(TP, mg/L)', xaxis.title.size=1.25, panel.width=2),
                      
                      list(5, header='Light Gray Means Highlighted Above',
                           inactive.border.color=gray(.7), inactive.border.size=2,
                            panel.width=1.5)),
print.file = "output_micromap/mmplot_tp2019.jpeg", res=100)



```

The `echo: false` option disables the printing of code (only output is displayed).
