---
title: "Linked Micromap"
author: "Mike McManus"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Libraries

```{r library}
library(mapview)
library(micromap)
```

## Log

2026-04-22 pasted workshop code

## Data

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r data}
names(polygons8)
View(polygons8)

polygons6 <- filter(polygons8, SEGMENT_NA != "Crooked Creek" & SEGMENT_NA != "Burnt Mill Creek")

chla_2019v1 <- filter(chla_2019, SEGMENT_NA != "Crooked Creek")
names(chla_2019v1)

# list sf object first, then dataframe of stats to create polygons with stats
chla_2019v2 <- merge(polygons6, chla_2019v1, by = "SEGMENT_NA" )
summary(chla_2019v2)

```

## Linked Micromap
Note cannot have a column called "median" in data.
Also, for chla data need log transformation
Getting 3 micromaps, one uncolored, one with 6 colors, and one all gray

```{r linked_micromap}
#Draft plot
mmplot(map.data=chla_2019v2,
       panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       list('min', 'q1', 'med', 'q3', 'max'),
                       NA),
       grouping = 6,
       ord.by='med', rev.ord=TRUE,
       median.row=FALSE)
       #map.link=c('SEGMENT_NA'))

#Refined Plot
mmplot(stat.data=wv.data.ds1,
       map.data=wv.map.table,
       panel.types=c('dot_legend', 'labels', 'box_summary', 'dot_cl', 'map'),
       panel.data=list(NA,
                       'alt.name',
                       list('cond_min', 'condq1', 'cond_med1', 'condq3', 'cond_max'),
                       list('Cond_med', 'Cond_LCB95', 'Cond_UCB95'),                       
                       NA),
       ord.by='Cond_med', rev.ord=TRUE,
       grouping=5, plot.pGrp.spacing=1,
       median.row=FALSE,
       map.link=c('Random_Wat', 'ID'),
       
       plot.height=9, plot.width=6.8, #for poster use plot.height=6, plot.width=9,
       colors=brewer.pal(5, "Spectral"), #for journal use plot.height=9, plot.width=6.8
              
       panel.att=list(list(1, panel.width=1, point.type=20, point.size=2,
                           point.border=TRUE), 
                      
                      list(2, header='WVDEP\n Subbasins', panel.width=1.25,#increased panel.width to 1.1 from .8
                           align='left', text.size=.8),
                      
                      list(3, header='Observed Conductivity \n 4th Order Streams or Less',
                           graph.bgcolor='lightgray',
                           xaxis.ticks=c(0, 1000, 2000, 3000, 4000, 5000),
                           xaxis.labels=c(0, 1, 2, 3, 4, 5),
                           xaxis.labels.size=1.25,
                           #xaxis.labels.angle=90,
                           xaxis.title='Conductivity (x1000)', xaxis.title.size=1.25, panel.width=2),
                      
                      list(4, header='Estimated Conductivity \n 4th Order Streams or Less',
                           graph.bgcolor='lightgray',
                           xaxis.ticks=list(0, 150, 300, 450, 600, 750),
                           xaxis.labels=list(0, 150, 300, 450, 600, 750),
                           xaxis.labels.size=1.25,
                           xaxis.title='Conductivity',
                           add.line=129, add.line.col="black", add.line.typ="dashed",
                           xaxis.title.size=1.25,panel.width=2),
                      
                      list(5, header='Light Gray Means\n Highlighted Above',
                           inactive.border.color=gray(.7), inactive.border.size=2,
                            panel.width=1.5)))

```

The `echo: false` option disables the printing of code (only output is displayed).
