---
output: html_document
editor_options: 
  chunk_output_type: console
---
## Libraries

```{r libraries}
library(tidyverse)
library(sf)
library(mapview)
library(janitor)
library(leaflet)
library(leafpop) # for popups in 
library(leafsync) # sync mapview maps
library(patchwork)
library(nhdplusTools)
library(tmap)
library(forcats)
library(gifski) #for tmap animation
```

## Log

2025-12-27 read in win_baywatch_sabnut.csv and using code and data frame names from script_01_esda.qmd

## Reading in Data and Formatting Date Variables

One csv file, win_baywatch_sabnut.csv, uploaded by Ryann Rossi, SASJBEP, on December 23, 2025. Ryan pulled the data from WIN and did some data cleaning. Note activity_start_date comes in as character.

Read in Bays_Bayous_Lakes_V1 shapefile that Rose uploaded to repository.

```{r data}
np_ds1 <- read_csv("data/win_baywatch_sabnut.csv")
names(np_ds1)
np_ds1 <- clean_names(np_ds1)
# spec(np_ds1)
# names(np_ds1)


# See Long and Lat in decimal degrees and date as character
View(np_ds1) 
str(np_ds1)
glimpse(np_ds1)
class(np_ds1$date)

#zdates <- distinct(np_ds1, activity_start_date)
#ym_ds1 <- lubridate::mdy(zdates$activity_start_date)
#floor_date(ym_ds1, "month")
#saveRDS(zdates, "zdates")

# with activity start date still character have to assign factors
# have to create activity start date as factor for tmap animation
np_ds1$date_factor <- np_ds1$date
np_ds1$yrmm_factor <- np_ds1$date

# convert to date
np_ds1$date <- lubridate::mdy(np_ds1$date)
class(np_ds1$date)
np_ds1$yrmm <- floor_date(np_ds1$date, "month")
head(np_ds1$yrmm)
class(np_ds1$yrmm)

# have to create activity start date as factor for tmap animation
np_ds1 <- arrange(np_ds1, date)
np_ds1$date_factor <- as_factor(np_ds1$date_factor)
head(np_ds1$date_factor)
levels(np_ds1$date_factor)


np_ds1$yrmm_factor <- as_factor(as.character(np_ds1$yrmm))
head(np_ds1$yrmm_factor)
levels(np_ds1$yrmm_factor)

# filter(dg_sf1, (!is.na(log_n2o))

np_ds1 <- filter(np_ds1, (!is.na(long)))
# drops to 10,369 obs and summary(mp_ds2$lat) returns no NA

```

## Create sf object
Have NA for some lat and long entries so remove these
```{r create_sf}

# create sf object
coords <- data.frame(longitude  = np_ds1$long, latitude = np_ds1$lat)
head(coords)

# See Marc Weber workshop above about Coordinate Reference Systems (CRS)

np_sf1 <- st_as_sf(np_ds1, coords = c("long", "lat"), crs =4269) %>% st_transform((5070))

class(np_sf1)
# read in shapefile
nnc_1 <- st_read("data_gis/NNC_Bay_Segments.shp")
class(nnc_1)
summary(nnc_1)
# see that espg code is 3087
# asking do the two sf objects have the same CRS
st_crs(np_sf1) == st_crs(nnc_1)

np_sf1 <- np_sf1 %>% st_transform(3087)
st_crs(np_sf1) == st_crs(nnc_1)

mapview(np_sf1) + mapview(nnc_1)

# mapview(hab_sf1, zcol = "Depth", cex = "Depth", alpha.regions = .8, legend = TRUE, popup = popupTable(hab_sf1, zcol = c("Depth", "Site"))) + mapview(bigbend_seagrass)

bbl_1 <- st_read("data_gis/Bays_Bayous_Lakes_V1.shp")
class(bbl_1)
summary(bbl_1)

bbl_1 <- bbl_1 %>% st_transform(3087)

mapview(bbl_1)

mapview(np_sf1) + mapview(bbl_1)

```

## Data Structure and Summaries

Six distinct organizations collected monitoring data with Florida Lakewatch being a synonym for St. Andrew Baywatch. An NA organization is listed and missing is NWFWMD. Need to compare Ryann and Rose's queries. Get 1020 dates spanning 1998-07-10 to 2025-04-23. Have 130 distinct monitoring location IDs, including NA. Have 210 distinct monitoring location names. There are 7894 distinct coordinates/locations.

But, I am wondering if DEP assigns coordinates that are not the actual locations of the monitoring station, but instead are always to same place (the centroid) of the spatial feature. Also, are there any analyte code remarks about questionable values?

```{r data_structure}
names(np_ds1)
distinct(np_ds1, organization_name)

distinct(np_ds1, sampling_agency_name) # shows ST. ANDREW BAY RESOURCE MANAGEMENT ASSOCIATION INC.

distinct(np_ds1, date) %>% arrange(., date) %>% 
  print(n = Inf)

np_ds1 %>% 
  group_by(organization_name) %>% 
  summarise(count = n_distinct(monitoring_location_name))

distinct(np_ds1, monitoring_location_id) %>% arrange(., monitoring_location_id) %>% 
  print(n = Inf)

distinct(np_ds1, monitoring_location_name) %>% arrange(., monitoring_location_name) %>% 
  print(n = Inf)

distinct(np_ds1, chl) %>% arrange(., chl) %>% print(n = Inf)
distinct(np_ds1, tn) %>% arrange(., tn) %>% print(n = Inf)
distinct(np_ds1, tp) %>% arrange(., tp) %>% print(n = Inf)

# distinct monitoring stations
np_stations <- unique.data.frame(np_sf1) # returns 10,367 obs/stations & geometries

np_stations %>% 
  group_by(organization_name) %>% 
  summarise(count = n_distinct(monitoring_location_name))

# distinct bay
distinct(np_ds1, bay) %>% print(n = Inf)

np_ds1 %>% 
  group_by(bay) %>% 
  summarise(count = n_distinct(monitoring_location_name)) %>% 
  print(n = Inf)

```

Interesting that total phosphorus has monitoring names that include format of DEP's status survey of zone 1 (Z1) for NWFMD region followed by resource type large lake (LL), but that's not the case for total nitrogen.

```{r summaries}
# tn
dplyr::select(np_ds1, tn) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn)))

dplyr::select(np_ds1, tn, bay) %>% 
group_by(., bay) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)

# tp
dplyr::select(np_ds1, tp) %>% 
summarize(., min = fivenum(tp)[1],
              q1 = fivenum(tp)[2],
              median = fivenum(tp)[3],
              q3 = fivenum(tp)[4],
              max = fivenum(tp)[5],
              count = dplyr::n(),
              mean = mean(tp, na.rm = TRUE),
              na_count = sum(is.na(tp)))

dplyr::select(np_ds1, tp, bay) %>% 
group_by(., bay) %>% 
summarize(., min = fivenum(tp)[1],
              q1 = fivenum(tp)[2],
              median = fivenum(tp)[3],
              q3 = fivenum(tp)[4],
              max = fivenum(tp)[5],
              count = dplyr::n(),
              mean = mean(tp, na.rm = TRUE),
              na_count = sum(is.na(tp))) %>% 
print(n = Inf)

# chl
dplyr::select(np_ds1, chl) %>% 
summarize(., min = fivenum(chl)[1],
              q1 = fivenum(chl)[2],
              median = fivenum(chl)[3],
              q3 = fivenum(chl)[4],
              max = fivenum(chl)[5],
              count = dplyr::n(),
              mean = mean(chl, na.rm = TRUE),
              na_count = sum(is.na(chl)))

dplyr::select(np_ds1, chl, bay) %>% 
group_by(., bay) %>% 
summarize(., min = fivenum(chl)[1],
              q1 = fivenum(chl)[2],
              median = fivenum(chl)[3],
              q3 = fivenum(chl)[4],
              max = fivenum(chl)[5],
              count = dplyr::n(),
              mean = mean(chl, na.rm = TRUE),
              na_count = sum(is.na(chl))) %>% 
print(n = Inf)
  
```

## Spatial Intersect of Sub-bays

Does spatial intersection result in, for example, of upper Crooked Creek being excluded? Yes, it does. Do Rose and Ryann needed to include, possibly, NHDPlus polygons of Crooked Creek and Burnt Mill Creek so as not to "lose" stations in those locations?
Spatial intersection with sub-bay drops obs to 5737
```{r st_intersect}
# four sub_bays
subbay_1 <- dplyr::filter(bbl_1, SEGMENT_NA == "West Bay" | SEGMENT_NA == "North Bay" | SEGMENT_NA == "St. Andrew Bay" | SEGMENT_NA == "East Bay")

mapview(subbay_1)
sub_np1 <- st_intersection(np_sf1, subbay_1)
head(sub_np1)


mapview(sub_np1) + mapview(bbl_1)
mapview(sub_np1) + mapview(subbay_1)

```

## Sub-bay Data Structure
Note sector does not always match SEGMENT_NA

```{r subbay_data_struct}
names(sub_np1)
distinct(sub_np1, organization_name) # 4 organizations

distinct(sub_np1, sampling_agency_name)

distinct(sub_np1, date) %>% arrange(., date) %>% 
  print(n = Inf) #660 dates from 1990-02-11 to 2024-12-17

sub_np1 %>% 
  group_by(organization_name) %>% 
  summarise(count = n_distinct(monitoring_location_name))

distinct(sub_np1, monitoring_location_id) %>% arrange(., monitoring_location_id) %>% 
  print(n = Inf)

distinct(sub_np1, monitoring_location_name) %>% arrange(., monitoring_location_name) %>%   print(n = Inf)

distinct(sub_np1, bay) %>% arrange(., bay) %>%   print(n = Inf)
# return 15 bays

# distinct(sub_np1, dep_analyte_name)

# distinct monitoring stations
sub_np_stations <- unique.data.frame(sub_np1) # returns 5736 obs/stations & geometries

sub_np_stations %>% 
  group_by(organization_name) %>% 
  summarise(count = n_distinct(monitoring_location_name))

distinct(sub_np_stations, sector)
```

## Subbay Stations
```{r subbay_tn}

# tn
dplyr::select(sub_np_stations, tn) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn)))

dplyr::select(sub_np_stations, tn, SEGMENT_NA) %>% 
group_by(., SEGMENT_NA) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)

mapview(sub_np_stations)

sub_np_stations %>% 
  group_by(SEGMENT_NA) %>% 
  summarise(count = n_distinct(u_of_f_number))

# tn_sub1 <- filter(sub_np1, dep_analyte_name == "Nitrogen- Total") 
# tn data is only from Florida Lakewatch

# distinct(tn_sub1, organization_name)

# distinct tn monitoring stations
# tn_stations <- unique.data.frame(tn_sub1)

# mapview(tn_sub1)

# mapview(tn_stations)
```

## North Bay
Believe it has smallest n of stations based on SEGEMENT_NA
Does not appear to be unique station identifier as have NA entries for U of F number, in addition to North Bay-1, North Bay-2, North Bay-3, and Fannin Bayou-1
No samples collected in 2006, though station coordinated entered. No samples collected from 2007-2017 and no coordinates entered.
```{r northbay}
nb_ds1 <- dplyr::filter(sub_np_stations, SEGMENT_NA == "North Bay")

distinct(nb_ds1, u_of_f_number)
distinct(nb_ds1, monitoring_location_name)
distinct(nb_ds1, sta_nr)
# does 1.1, 1.2,1.3, and 1.7 corresponds to North Bay 1, 2, 3 and Fannin Bayou?

dplyr::select(nb_ds1, tn, monitoring_location_name) %>% 
group_by(., monitoring_location_name) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)

# why 620 observations not assigned a monitoring location name

dplyr::select(nb_ds1, tn, sta_nr) %>% 
group_by(., sta_nr) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)
# now returns only 178 with sta_nr NA

dplyr::select(nb_ds1, tn, year) %>% 
group_by(., year) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)

dplyr::select(nb_ds1, tn, date_factor) %>% 
group_by(., date_factor) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)


mapview(nb_ds1)
# why does map show 8 stations

nb_ds2 <- dplyr::select(nb_ds1, tn, monitoring_location_name, geometry)

nb_geom <- dplyr::select(nb_ds1, geometry)

distinct(nb_geom)
mapview(nb_geom)
```

## tmap North Bay
Does tn is NA means no sample taken or sample taken by concentration was not detectable? Filter to just year 1999 gives 43 observations, but with some tn being NA. Filter 1999 and (!(is.na(tn))) gives 12 observations. tmap does not like missing values so no longer have Fannin Bayou. Was Fannin Bayou not sampled or was it sampled and had non-detecable concentration. Use tm_symbols(fill = "tn") to get nondetectable NA's mapped as missing data. Also, using date_factor showed two April samplings. No entries for 2007-2017 so no map?
```{r tmap_nb}

distinct(nb_ds1, year) %>% print(n=Inf)

# (dg_sf1, (!is.na(log_n2o))
nb_ds3 <- filter(nb_ds1, year == "2024")
# nb_ds4 <- filter(nb_ds1, year == "1990" & (!(is.na(tn))))

tmap_mode("plot")
tmap_mode("view")
tn_map_1990_v1 <- tm_shape(subbay_1) +
  tm_polygons(fill = "white", lwd = 2) +
  tm_title("Detectable and Nondetectable Total Nitrogen (mg/L) in North Bay, 2024")  + 
  tm_shape(nb_ds3) +
  tm_symbols(fill = "tn") +
  tm_facets(by = "date_factor")

tn_map_1990_v1

tmap_mode("plot")
tn_bubble <- tm_shape(subbay_1) +
  tm_fill("#dddddd") + 
tm_shape(nb_ds3) +
  tm_symbols(fill =  "tn",
             # size.scale = tm_scale_continuous_log(values.scale = "dep_result_value_number"),
             size.legend = tm_legend("TN (mg/L)"))

#daily loop
tn_bubble + tm_animate(frame = "date_factor", fps = 1, play = "loop")

# monthly loop
tn_bubble + tm_animate(frame = "yrmm_factor", fps = 1, play = "loop")



  tm_compass(position = c("LEFT", "BOTTOM")) +
  tm_scalebar(position = c("LEFT", "BOTTOM")) +
```

## West Bay
Filtering misses most upstream Crooked Creek and Burnt Mill Creek stations.
```{r westbay}
wb_ds1 <- dplyr::filter(sub_np_stations, SEGMENT_NA == "West Bay")

distinct(wb_ds1, u_of_f_number)
distinct(wb_ds1, monitoring_location_name)
distinct(wb_ds1, sta_nr)
# does 1.1, 1.2,1.3, and 1.7 corresponds to North Bay 1, 2, 3 and Fannin Bayou?

dplyr::select(wb_ds1, tn, monitoring_location_name) %>% 
group_by(., monitoring_location_name) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)

# why 575 observations not assigned a monitoring location name

dplyr::select(wb_ds1, tn, sta_nr) %>% 
group_by(., sta_nr) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)
# now returns only 1555 with sta_nr NA

dplyr::select(wb_ds1, tn, year) %>% 
group_by(., year) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)

dplyr::select(wb_ds1, tn, date_factor) %>% 
group_by(., date_factor) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)


mapview(wb_ds1)
# why does map show 8 stations

wb_ds2 <- dplyr::select(wb_ds1, tn, monitoring_location_name, geometry)

wb_geom <- dplyr::select(wb_ds1, geometry)

distinct(wb_geom)
mapview(wb_geom)
```

## tmap West Bay
```{r tmap_nw}

distinct(wb_ds1, year) %>% print(n=Inf)

# (dg_sf1, (!is.na(log_n2o))
wb_ds3 <- filter(wb_ds1, year == "1990")
# wb_ds4 <- filter(wb_ds1, year == "1990" & (!(is.na(tn))))

tmap_mode("plot")
# tmap_mode("view")
tn_map_1990_v1 <- tm_shape(subbay_1) +
  tm_polygons(fill = "white", lwd = 2) +
  tm_title("Detectable and Nondetectable Total Nitrogen (mg/L) in West Bay, 1990")  + 
  tm_shape(wb_ds3) +
  tm_symbols(fill = "tn") +
  tm_facets(by = "date_factor")

tn_map_1990_v1

tmap_mode("plot")
tn_bubble <- tm_shape(subbay_1) +
  tm_fill("#dddddd") + 
tm_shape(wb_ds3) +
  tm_symbols(fill =  "tn",
             # size.scale = tm_scale_continuous_log(values.scale = "dep_result_value_number"),
             size.legend = tm_legend("TN (mg/L)"))

#daily loop
tn_bubble + tm_animate(frame = "date_factor", fps = 1, play = "loop")

# monthly loop
tn_bubble + tm_animate(frame = "yrmm_factor", fps = 1, play = "loop")



  tm_compass(position = c("LEFT", "BOTTOM")) +
  tm_scalebar(position = c("LEFT", "BOTTOM")) +
```
