---
title: "Linked Micromap 2020"
author: "Mike McManus"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Libraries

```{r libraries}
library(tidyverse)
library(sf)
library(mapview)
library(janitor)
library(leaflet)
library(leafpop) # for popups in 
library(leafsync) # sync mapview maps
library(leafem) # add labels to map
library(patchwork)
library(nhdplusTools)
library(tmap)
library(forcats)
library(gifski) #for tmap animation
library(plotly)
library(webshot) # save mapview object
library(nngeo)
library(micromap)
library(tidyverse)
```

## Log

2026-05-30 calculated distances from Warren Bayou point/station to polygons in bayous_wb1

2026-05-06 renamed with YYYY in script name

## Reading in Data and Formatting Date Variables

One csv file, win_baywatch_sabnut.csv, uploaded by Ryann Rossi, SASJBEP, on December 23, 2025. Ryan pulled the data from WIN and did some data cleaning. Note activity_start_date comes in as character.

Read in Bays_Bayous_Lakes_V4 shapefile that Rose uploaded to repository. Can I merge wq data, np_ds1, with Baywatch station IDs, bw_id1, by coordinates?

```{r data}
np_ds1 <- read_csv("data/win_baywatch_sabnut.csv")
names(np_ds1)
np_ds1 <- clean_names(np_ds1)
# spec(np_ds1)
# names(np_ds1)


# See Long and Lat in decimal degrees and date as character
View(np_ds1) 
str(np_ds1)
glimpse(np_ds1)
class(np_ds1$date)

#zdates <- distinct(np_ds1, activity_start_date)
#ym_ds1 <- lubridate::mdy(zdates$activity_start_date)
#floor_date(ym_ds1, "month")
#saveRDS(zdates, "zdates")

# with activity start date still character have to assign factors
# have to create activity start date as factor for tmap animation
np_ds1$date_factor <- np_ds1$date
np_ds1$yrmm_factor <- np_ds1$date

# convert to date
np_ds1$date <- lubridate::mdy(np_ds1$date)
class(np_ds1$date)
np_ds1$yrmm <- floor_date(np_ds1$date, "month")
head(np_ds1$yrmm)
class(np_ds1$yrmm)

# have to create activity start date as factor for tmap animation
np_ds1 <- arrange(np_ds1, date)
np_ds1$date_factor <- as_factor(np_ds1$date_factor)
head(np_ds1$date_factor)
levels(np_ds1$date_factor)


np_ds1$yrmm_factor <- as_factor(as.character(np_ds1$yrmm))
head(np_ds1$yrmm_factor)
levels(np_ds1$yrmm_factor)

# filter(dg_sf1, (!is.na(log_n2o))

np_ds1 <- filter(np_ds1, (!is.na(long)))
# drops to 10,369 obs and summary(mp_ds2$lat) returns no NA

unique(np_ds1$organization_name)
distinct(np_ds1,bay) %>%  dplyr::arrange(., bay) %>% print(n=Inf)
# 35 distinct bay entries
distinct(np_ds1, monitoring_location_id) %>%  dplyr::arrange(., ) %>% print(n=Inf)
# 130 distinct monitoring location id entries

z1 <- filter(np_ds1, is.na(organization_name))
# return 7,162 obs
summary(z1$sta_nr) # code min 1.1 to max 9.5
# believe having a numeric sta_nr code means this is a Lakewatch/Baywatch station and have long/lat coordinates for these stations

z2 <- filter(np_ds1, organization_name == "FLORIDA LAKEWATCH")
# return 2,380
summary(z2$sta_nr) # returns all NAs
# want to get numeric codes assigned to sta_nr, struggling to get that done


```

## Explore Duplicates and NAs

```{r dupes_na}
# z1 <- get_dupes(np_ds1)
# 
# z2 <- filter(np_ds1, (is.na(tn) & !is.na(tp)))
# 
# z3 <- filter(np_ds1, is.na(tn))
```

## Create sf object

Have NA for some lat and long entries so remove these. Also, Numeric Nutrient Criteria (NNC) polygons omit bayous, creeks, and lakes, but includes intracoastal waterway and St. Joe Bay. But bayous, bays, and lakes does have all those polygons. Although, a Burnt Mill Creek station does not hit that polygon.

```{r create_sf}

# create sf object
coords <- data.frame(longitude  = np_ds1$long, latitude = np_ds1$lat)
head(coords)

# See Marc Weber workshop above about Coordinate Reference Systems (CRS)

np_sf1 <- st_as_sf(np_ds1, coords = c("long", "lat"), crs =4269) %>% st_transform((5070))

class(np_sf1)
# read in shapefile
nnc_1 <- st_read("data_gis/NNC_Bay_Segments.shp")
class(nnc_1)
summary(nnc_1)
distinct(nnc_1, SEGMENT_NA)
distinct(nnc_1, WATERBODY_)

mapview(nnc_1)
# Numeric

# see that espg code is 3087
# asking do the two sf objects have the same CRS
st_crs(np_sf1) == st_crs(nnc_1)

np_sf1 <- np_sf1 %>% st_transform(3087)
st_crs(np_sf1) == st_crs(nnc_1)

mapview(np_sf1) + mapview(nnc_1)

# mapview(hab_sf1, zcol = "Depth", cex = "Depth", alpha.regions = .8, legend = TRUE, popup = popupTable(hab_sf1, zcol = c("Depth", "Site"))) + mapview(bigbend_seagrass)


```

## Read Bays, Bayous and Lakes V4

Rose Horn took 2 polygons for Alligator Bayou, which borders the U.S. Coast Guard Station on St. Andrew Bay, and splits them into separate sampling areas. Also, in version the new update reflects the removal of small edge portions of St. Andrew Bay that were along the edges of East Bay and Grand Lagoon. Looks like I should pick-up Upper and Lower Burnt Mill Creek Stations.

Need to see if I have 8 polygon names for: East Bay, Grand Lagoon, North Bay, St. Andrew Bay, West Bay, Crooked and Burnt Mill Creeks, and Deer Point Lake as I want to use those 6 polygons in linked micromap. Perhaps by using SEGMENT_NA and Major_Seg I can get the polygons I want.

```{r bbl_v4}

bbl_4 <- st_read("data_gis/Bays_Bayous_Lakes_V4/Bays_Bayous_Lakes_V4.shp")

bbl_4 <- bbl_4 %>% st_transform(3087)

summary(bbl_4)
mapview(bbl_4)


distinct(bbl_4, SEGMENT_NA) %>% dplyr::arrange(., SEGMENT_NA)
# have 190 distinct SEGMENT_NA, but 191 obs suggesting one polygon is sharing a name

# Is Major_Seg a new field from Rose?
distinct(bbl_4, Major_Seg) %>% dplyr::arrange(., Major_Seg)
# returns 8 Major_Seg


mapview(np_sf1) + mapview(bbl_4)
# a burnt mill creek station from 2024-04-25 is off polygon

# these 8 polygons might be what I use for a linked micromap
# polygons8 <- dplyr::filter(bbl_4, SEGMENT_NA == "West Bay" | SEGMENT_NA == "North Bay" | SEGMENT_NA == "St. Andrew Bay" | SEGMENT_NA == "East Bay" | SEGMENT_NA == "Burnt Mill Creek" | SEGMENT_NA == "Crooked Creek" | SEGMENT_NA == "Grand Lagoon" |SEGMENT_NA == "Deer Point Lake")
# 
# mapview(polygons8)

# mapview(np_sf1) + mapview(polygons8)

```

## West Bay, Bayous, and Creeks

The two code chunks below illustrate 2 ways to format data to make linked micromaps. The first way involves filtering from the points/stations data so it only has points within the polygons of interest. Also, the filtered data need to have variable name, in this case "bay", to match to the polgyon names entered under SEGMENT_NA. A five-number summary is made of the WQ data grouped by bay. Next, the appropriate polygons are selected and, if needed, values under SEGMENT_NA are replaced to match the bay entries of the points. The polygon data and five-number summary data are merged using the shared variable names of SEGMENT_NA and bay.

Mapvview of mm_stations1 shows monitoring location id BAY-WE-STBAY-5 clearly in SAB, but its bay is labeled as West Bay. Since December 1, 2018, BAY-WE-STBAY-5 has had either the wrong coordinate entered or the wrong bay or both, and it has 28 entries.

Note monitoring location id is all upper case so always use that. Monitoring location name is snake-cased so avoid it. Note subtle behavior of filter and filter_out with NAs, <https://dplyr.tidyverse.org/reference/filter.html>.

```{r wb_stats_spatial_join}

mm_station1 <- dplyr::filter(np_sf1, bay == "Crooked Creek" | bay == "Burnt Mill Creek" | bay == "Warren Bayou" | bay == "West Bay")
# return 1778 observations, of which 1418 observations have sta_nr numeric entry, and 360 have monitoring location id entry

distinct(mm_station1, monitoring_location_id) %>% arrange(., monitoring_location_id) %>% print(n=Inf)
# returns 21 observations including NA, 3 Burnt Mill stations, 3 Crooked stations, 1 Warren Bayou, 11 BAY-WE-STBAY stations, 1 BAY-WEST-BAY-12, and 1 NWFYI0005

ggplot(mm_station1 |> filter(year >=2020), aes(x=yrmm_factor, y=log(tp))) + geom_point() + facet_wrap(vars(monitoring_location_id))

mapview(mm_station1) + bayous_wb1

odd_coordinates <- dplyr::filter(mm_station1, monitoring_location_id == "BAY-WE-STBAY-3" | monitoring_location_id == "BAY-WE-STBAY-5")

mapview(odd_coordinates) + bayous_wb1

mm_station1 <- dplyr::filter_out(mm_station1, monitoring_location_id == "BAY-WE-STBAY-5")
# believe this filtering out removes all 28 obs for monitoring location BAY-WE-STBAY-5 so now have 1750 obs

distinct(mm_station1, monitoring_location_id) %>% arrange(., monitoring_location_id) %>% print(n=Inf)
# now returns to 20 obs and no BAY-WE-STBAY-5 observation


mm_station1 %>% 
  group_by(bay) %>% 
  summarize(., count = dplyr::n())
# want to add a bay entry of "West Bay ICWW" for BAY-WEST BAY-12


mm_station2 <- mm_station1 |>
  mutate(
    bay = case_when(monitoring_location_id == "BAY-WEST-BAY-12" ~ "West Bay ICWW",
                       TRUE ~ bay
                       )
    )

mm_station2 %>% 
  group_by(bay) %>% 
  summarize(., count = dplyr::n())

bayous_wb1 <- dplyr::filter(bbl_4, SEGMENT_NA == "Burnt Mill Creek" | SEGMENT_NA == "Crooked Creek" | SEGMENT_NA == "Warren Bayou" | SEGMENT_NA == "Gulf ICWW between Choctawhatchee Bay and St. Andrew Bay" | SEGMENT_NA == "West Bay")


# do replace values so SEGMENT_NA will match to bay in points data frame
bayous_wb2 <- bayous_wb1 |> 
  mutate(
  SEGMENT_NA = SEGMENT_NA |>
    replace_values(
      "Gulf ICWW between Choctawhatchee Bay and St. Andrew Bay" ~ "West Bay ICWW"
    )
  )  

# getting 5-number summary from points data
tp_wb2020 <- st_set_geometry(mm_station2, NULL) %>% 
  dplyr::filter(., year == 2020 & sampling_agency_name == "ST. ANDREW BAY RESOURCE MANAGEMENT ASSOCIATION INC.") %>% 
  dplyr::select(., tp, bay) %>% 
  group_by(., bay) %>% 
  summarize(., tp_min = fivenum(tp)[1],
            tp_q1 = fivenum(tp)[2],
            tp_med = fivenum(tp)[3],
            tp_q3 = fivenum(tp)[4],
            tp_max = fivenum(tp)[5],
            tp_lmin = log(fivenum(tp)[1]),
            tp_lq1 = log(fivenum(tp)[2]),
            tp_lmed = log(fivenum(tp)[3]),
            tp_lq3 = log(fivenum(tp)[4]),
            tp_lmax = log(fivenum(tp)[5]),
            tp_count = dplyr::n(),
            tp_mean = mean(tp, na.rm = TRUE),
            tp_na_count = sum(is.na(tp))) %>% 
  print(n = Inf)


tp_wb2020v2 <- merge(bayous_wb2, tp_wb2020, by.x = "SEGMENT_NA", by.y = "bay" )


mmplot(map.data=tp_wb2020v2,
       panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 1,
       ord.by='tp_lmed', rev.ord=TRUE,
       median.row=FALSE)


# TP Refined Plot for West Bay Creeks and Bayous
mmplot(map.data=tp_wb2020v2,
       panel.types=c('dot_legend', 'labels', 'labels', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       'tp_count',
                       'tp_na_count',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 2,
       ord.by='tp_lmed',
       rev.ord=TRUE,
       median.row=FALSE,
       plot.pGrp.spacing=1,
       
       plot.height=6, plot.width=9, #for poster use plot.height=6, plot.width=9,
       colors=brewer.pal(4, "Spectral"), #for journal use plot.height=9, plot.width=6.8
              
       panel.att=list(list(1, panel.width=1, point.type=20, point.size=2,
                           point.border=TRUE), 
                      
                      list(2, header='West Bay Segment,\n Creeks & Bayous', panel.width=1.2,#increased panel.width to 1.1 from .8
                           align='left', text.size=.8),
                      
                      list(3, header='Total #\n of Samples',
                           panel.width=.35,
                           align='right',
                           text.size=.8),
                      
                      list(4, header='# of NA\n Samples',
                           panel.width=.4,
                           align='right',
                           text.size=.8),
                      
                      list(5, header='2020 Detectable Total Phoshporous',
                           graph.bgcolor='lightgray',
                           xaxis.ticks=c(-6.0, -5.0, -4.0, -3.0, -2.0),
                           xaxis.labels=c(-6,-5, -4, -3, -2),
                           xaxis.labels.size=1.25,
                           graph.bar.size = 0.5,
                           #xaxis.labels.angle=90,
                           add.line=-4.07, add.line.col="black",
                           xaxis.title='Log(TP, mg/L)', xaxis.title.size=1.25, panel.width=1.75),
                      
                      list(6, header='Light Gray Means Highlighted Above',
                           inactive.border.color=gray(.7), inactive.border.size=2,
                            panel.width=1.75)),
print.file = "output_micromap/mmplot_tp_wb2020v2.jpeg", res=100)



```

I do not recommend this second method as it assumes points and polygons are exactly geograpically matched. Consequently, I have commented out this code chunk. The second way to format the data assumes the points and polgyon datasets are in exact alignment. That is every point/station is in its appropriate polygon. A spatial intersection is done of the points/stations to the polygons so that now all the polygon attributes, such as SEGMENT_NA, are assigned to the points. After this spatial intersection is done, then a five-number summary of the WQ data can be made and the linked micromaps created.

Spatial intersection ends up omitting Bay-West Bay-3 from April 25, 2024 as that monitoring location does not intersect polygons for some reason. Need to think of summary statistics for Burnt Mill Creek, Crooked Creek and GICWW-West given their small sample sizes relative to West Bay main segment.

Also, Warren Bayou falls in polygon of West Bay segment so that's why Warren Bayou not showing up in linked micromap. The spatial intersection I do puts the Warren Bayou point into the West Bay segment. Code is included to measure how far the Warren Bayou station is from the Warren Bayou polygon.

```{r wb_stats_spatial_interseciton}

# bayous_wb1 <- dplyr::filter(bbl_4, SEGMENT_NA == "Burnt Mill Creek" | SEGMENT_NA == "Crooked Creek" | SEGMENT_NA == "Warren Bayou" | SEGMENT_NA == "Gulf ICWW between Choctawhatchee Bay and St. Andrew Bay" | SEGMENT_NA == "West Bay")
# 
# # mm_station1 <- dplyr::filter(np_sf1, bay != "Lake Powell" & bay != "Powell Creek" & bay != "Lullwater" & bay != "Lake Marin" & bay != "Marin")
# 
# 
# warren_v1 <- dplyr::filter(mm_station1, bay == "Warren Bayou")
# #returns 162 observations
# mapview(warren_v1)
# summary(warren_v1$tp)
# summary(warren_v1$tn)
# summary(warren_v1$chl)
# 
# # pull out only geometry of Warren Bayou station
# warren_v2 <- st_geometry(warren_v1) %>% unique.data.frame(.)
# class(warren_v2)
# 
# st_crs(warren_v2) == st_crs(bayous_wb1)
# 
# # https://stackoverflow.com/questions/46214479/distance-between-a-set-of-points-and-a-polygon-with-sf-in-r
# st_distance(warren_v2, bayous_wb1)
# # returns 5 distances for each of the 5 polygons in bayous_wb1 and distances are in order of the polygons. Have a 0 m distance for West Bay as the warren_v2 point is in that polygon. The distance to the 3rd polygon, SEGMENT_NA is Warren Bayou, is 7.0 m.
# 
# mm_wbstation2 <- st_intersection(mm_station1, bayous_wb1)
# 
# 
# mapview(mm_wbstation2) + mapview(bayous_wb1, col.regions = "yellow")
# 
# names(mm_wbstation2)
# 
# mm_wbstation2 %>% 
#   group_by(SEGMENT_NA) %>% 
#   summarize(., count = dplyr::n())
# 
# mm_wbstation2 %>% 
#   group_by(SEGMENT_NA, year) %>% 
#   summarize(., count = dplyr::n()) %>% 
#   print(n=Inf)
# 
# # uses 2020 data as have samples from all 4 polygons
# 
# tp_wb2020 <- st_set_geometry(mm_wbstation2, NULL) %>% 
#   dplyr::filter(., year == 2020 & sampling_agency_name == "ST. ANDREW BAY RESOURCE MANAGEMENT ASSOCIATION INC.") %>% 
#   dplyr::select(., tp, SEGMENT_NA) %>% 
#   group_by(., SEGMENT_NA) %>% 
#   summarize(., tp_min = fivenum(tp)[1],
#             tp_q1 = fivenum(tp)[2],
#             tp_med = fivenum(tp)[3],
#             tp_q3 = fivenum(tp)[4],
#             tp_max = fivenum(tp)[5],
#             tp_lmin = log(fivenum(tp)[1]),
#             tp_lq1 = log(fivenum(tp)[2]),
#             tp_lmed = log(fivenum(tp)[3]),
#             tp_lq3 = log(fivenum(tp)[4]),
#             tp_lmax = log(fivenum(tp)[5]),
#             tp_count = dplyr::n(),
#             tp_mean = mean(tp, na.rm = TRUE),
#             tp_na_count = sum(is.na(tp))) %>% 
#   print(n = Inf)
# 
# tp_wb2020v2 <- merge(bayous_wb1, tp_wb2020, by = "SEGMENT_NA" )
# summary(tp_wb2020v2)
# 
# tp_wb2020v3 <- tp_wb2020v2 |> 
#   mutate(
#   SEGMENT_NA = SEGMENT_NA |>
#     replace_values(
#       "Gulf ICWW between Choctawhatchee Bay and St. Andrew Bay" ~ "West Bay ICWW"
#     )
#   )  
#     
# 
# mmplot(map.data=tp_wb2020v3,
#        panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
#        panel.data=list(NA,
#                        'SEGMENT_NA',
#                        list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
#                        NA),
#        grouping = 1,
#        ord.by='tp_lmed', rev.ord=TRUE,
#        median.row=FALSE)
# 
# 
# # TP Refined Plot for West Bay Creeks and Bayous
# mmplot(map.data=tp_wb2020v3,
#        panel.types=c('dot_legend', 'labels', 'labels', 'labels', 'box_summary', 'map'),
#        panel.data=list(NA,
#                        'SEGMENT_NA',
#                        'tp_count',
#                        'tp_na_count',
#                        list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
#                        NA),
#        grouping = 2,
#        ord.by='tp_lmed',
#        rev.ord=TRUE,
#        median.row=FALSE,
#        plot.pGrp.spacing=1,
#        
#        plot.height=6, plot.width=9, #for poster use plot.height=6, plot.width=9,
#        colors=brewer.pal(4, "Spectral"), #for journal use plot.height=9, plot.width=6.8
#               
#        panel.att=list(list(1, panel.width=1, point.type=20, point.size=2,
#                            point.border=TRUE), 
#                       
#                       list(2, header='West Bay Segment,\n Creeks & Bayous', panel.width=1.2,#increased panel.width to 1.1 from .8
#                            align='left', text.size=.8),
#                       
#                       list(3, header='Total #\n of Samples',
#                            panel.width=.35,
#                            align='right',
#                            text.size=.8),
#                       
#                       list(4, header='# of NA\n Samples',
#                            panel.width=.4,
#                            align='right',
#                            text.size=.8),
#                       
#                       list(5, header='2020 Detectable Total Phoshporous',
#                            graph.bgcolor='lightgray',
#                            xaxis.ticks=c(-6.0, -5.0, -4.0, -3.0, -2.0),
#                            xaxis.labels=c(-6,-5, -4, -3, -2),
#                            xaxis.labels.size=1.25,
#                            graph.bar.size = 0.5,
#                            #xaxis.labels.angle=90,
#                            add.line=-4.07, add.line.col="black",
#                            xaxis.title='Log(TP, mg/L)', xaxis.title.size=1.25, panel.width=1.75),
#                       
#                       list(6, header='Light Gray Means Highlighted Above',
#                            inactive.border.color=gray(.7), inactive.border.size=2,
#                             panel.width=1.75)),
# print.file = "output_micromap/mmplot_tp_wb2020.jpeg", res=100)

```

## North Bay and Bayous

Seems like looking Upper Goose Bay when I summarize by bay, and that's because Goose Bay station and Upper Goose Bay station are grouped together under bay as Goose Bay. From polygons, I need to select all the North Bay bayous to show how many there are and to show which ones have been sampled and not sampled.

Very ugly draft linked micromap shows that 1) Upper Goose Bay station stats should be assigned to Upper Goose Bay polygon, and 2) that a dozen bayous in North Bay do not have any samples.
```{r nb_stats_spatial_join}
nb_station1 <- dplyr::filter(np_sf1, bay == "Robinson Bayou" | bay == "Goose Bayou" | bay == "Upper Goose Bayou" | bay == "Upper Goose Bayou"| bay == "Fannin Bayou" |bay == "North Bay")
# return 1148 observations, 

distinct(nb_station1, monitoring_location_id) %>% arrange(., monitoring_location_id) %>% print(n=Inf)
# returns 12 observations including 1 NA, 1 Fannin Bayou,  1 Goose Bayou station, 4 North Bay stations, 1 Robinson Bayou, 1 Upper Goose Bayou station, and 3 FDEP Stations, G3TLHR0021, G3TLHR0031, and G3TLHR0055

ggplot(nb_station1 |> filter(year >=2020), aes(x=yrmm_factor, y=log(tp))) + geom_point() + facet_wrap(vars(monitoring_location_id))

ggplot(nb_station1 |> filter(year >=2020), aes(y=bay, x=log(tp))) + geom_boxplot() + facet_wrap(vars(yrmm_factor))

mapview(nb_station1) + bbl_4

# no odd coordinate placement in North Bay
# so commented out code
# odd_coordinates <- dplyr::filter(mm_station1, monitoring_location_id == "BAY-WE-STBAY-3" | monitoring_location_id == "BAY-WE-STBAY-5")
# 
# mapview(odd_coordinates) + bayous_wb1
# 
# mm_station1 <- dplyr::filter_out(mm_station1, monitoring_location_id == "BAY-WE-STBAY-5")
# believe this filtering out removes all 28 obs for monitoring location BAY-WE-STBAY-5 so now have 1750 obs

distinct(nb_station1, monitoring_location_id) %>% arrange(., monitoring_location_id) %>% print(n=Inf)
# still returns 12 obs

nb_station1 %>% 
  group_by(bay) %>% 
  summarize(., count = dplyr::n())

nb_station1 %>% 
  dplyr::filter(., year == 2020) %>% 
  group_by(bay) %>% 
  summarize(., count = dplyr::n())

nb_station1 %>% 
  dplyr::filter(., year >= 2020) %>% 
  group_by(year) %>% 
  summarize(., count = dplyr::n())

# this creates an Upper Goose Bayou so stats can be assigned to that station
nb_station2 <- nb_station1 |>
  mutate(
    bay = case_when(monitoring_location_id == "BAY-UPPER-GOOSE-BAYOU-1" ~ "Upper Goose Bayou",
                       TRUE ~ bay
                       )
    )

nb_station2 %>% 
  dplyr::filter(., year == 2020) %>% 
  group_by(bay) %>% 
  summarize(., count = dplyr::n())

# filter all North Bay polygons
bayous_nb1 <- dplyr::filter(bbl_4, SEGMENT_NA == "Fannin Bayou" | SEGMENT_NA == "Goose Bayou" | SEGMENT_NA == "Robinson Bayou" | SEGMENT_NA == "North Bay" | SEGMENT_NA == "Mud Bayou" | SEGMENT_NA == "Alligator Bayou_GeneratingPlant" | SEGMENT_NA == "Newman Bayou" | SEGMENT_NA == "Sheephead Bayou" | SEGMENT_NA == "Gainer Bayou" | SEGMENT_NA == "Hodges Bayou" |  SEGMENT_NA == "Mill Bayou" | SEGMENT_NA == "Beatty Bayou" | SEGMENT_NA == "Anderson Bayou" | SEGMENT_NA == "Upper Goose Bayou" | SEGMENT_NA == "Lynn Haven Bayou")

mapview(bayous_nb1)

mapview(nb_station1) + mapview(bayous_nb1)

# do replace values so SEGMENT_NA will match to bay in points data frame
# bayous_wb2 <- bayous_wb1 |> 
#   mutate(
#   SEGMENT_NA = SEGMENT_NA |>
#     replace_values(
#       "Gulf ICWW between Choctawhatchee Bay and St. Andrew Bay" ~ "West Bay ICWW"
#     )
#   )  

# getting 5-number summary from points data
tp_nb2020 <- st_set_geometry(nb_station2, NULL) %>% 
  dplyr::filter(., year == 2020 & sampling_agency_name == "ST. ANDREW BAY RESOURCE MANAGEMENT ASSOCIATION INC.") %>% 
  dplyr::select(., tp, bay) %>% 
  group_by(., bay) %>% 
  summarize(., tp_min = fivenum(tp)[1],
            tp_q1 = fivenum(tp)[2],
            tp_med = fivenum(tp)[3],
            tp_q3 = fivenum(tp)[4],
            tp_max = fivenum(tp)[5],
            tp_lmin = log(fivenum(tp)[1]),
            tp_lq1 = log(fivenum(tp)[2]),
            tp_lmed = log(fivenum(tp)[3]),
            tp_lq3 = log(fivenum(tp)[4]),
            tp_lmax = log(fivenum(tp)[5]),
            tp_count = dplyr::n(),
            tp_mean = mean(tp, na.rm = TRUE),
            tp_na_count = sum(is.na(tp))) %>% 
  print(n = Inf)

# use left_join as have a many polygons (n=16) to few stats (n=5 obs) and want to keep all the polygons whether polygons have stats or not
tp_nb2020v2 <- left_join(bayous_nb1, tp_nb2020, by = join_by(SEGMENT_NA == bay), keep = TRUE)

# use right_join so cut down polygons to n=5 to match stats
tp_nb2020v3 <- right_join(bayous_nb1, tp_nb2020, by = join_by(SEGMENT_NA == bay), keep = TRUE)

# using tp_nb2020v2 shows how many bayous do not have samples
mmplot(map.data=tp_nb2020v2,
       panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 5,
       ord.by='tp_lmed', rev.ord=TRUE,
       median.row=FALSE)

# using tp_nb2020v3 shows only polygons with stats
mmplot(map.data=tp_nb2020v3,
       panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 1,
       ord.by='tp_lmed', rev.ord=TRUE,
       median.row=FALSE)
```

## Station Issues

Specific follow-up needed for some stations, such as coordinates that put a station on land, monitoring location ID BAY-WE-STBAY-3 near Burnt Mill Creek. Also, monitoring location ID BAY-WE-STBAY-5 that has a bay assignment of West Bay but coordinates put it in St. Andrew Bay.

Note that Sheepshead Bayou in North Bay actually has Major_Seg of "East Bay" assigned to it. Also, that in North Bay, Fannin Bayou has 3 arms to that bayou.
