---
title: "Linked Micromap 2020"
author: "Mike McManus"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Libraries

```{r libraries}
library(tidyverse)
library(sf)
library(mapview)
library(janitor)
library(leaflet)
library(leafpop) # for popups in 
library(leafsync) # sync mapview maps
library(leafem) # add labels to map
library(patchwork)
library(nhdplusTools)
library(tmap)
library(forcats)
library(gifski) #for tmap animation
library(plotly)
library(webshot) # save mapview object
library(nngeo)
library(micromap)
library(tidyverse)
```

## Log

2026-05-30 calculated distances from Warren Bayou point/station to polygons in bayous_wb1

2026-05-06 renamed with YYYY in script name

## Reading in Data and Formatting Date Variables

One csv file, win_baywatch_sabnut.csv, uploaded by Ryann Rossi, SASJBEP, on December 23, 2025. Ryan pulled the data from WIN and did some data cleaning. Note activity_start_date comes in as character.

Read in Bays_Bayous_Lakes_V4 shapefile that Rose uploaded to repository. Can I merge wq data, np_ds1, with Baywatch station IDs, bw_id1, by coordinates?

```{r data}
np_ds1 <- read_csv("data/win_baywatch_sabnut.csv")
names(np_ds1)
np_ds1 <- clean_names(np_ds1)
# spec(np_ds1)
# names(np_ds1)


# See Long and Lat in decimal degrees and date as character
View(np_ds1) 
str(np_ds1)
glimpse(np_ds1)
class(np_ds1$date)

#zdates <- distinct(np_ds1, activity_start_date)
#ym_ds1 <- lubridate::mdy(zdates$activity_start_date)
#floor_date(ym_ds1, "month")
#saveRDS(zdates, "zdates")

# with activity start date still character have to assign factors
# have to create activity start date as factor for tmap animation
np_ds1$date_factor <- np_ds1$date
np_ds1$yrmm_factor <- np_ds1$date

# convert to date
np_ds1$date <- lubridate::mdy(np_ds1$date)
class(np_ds1$date)
np_ds1$yrmm <- floor_date(np_ds1$date, "month")
head(np_ds1$yrmm)
class(np_ds1$yrmm)

# have to create activity start date as factor for tmap animation
np_ds1 <- arrange(np_ds1, date)
np_ds1$date_factor <- as_factor(np_ds1$date_factor)
head(np_ds1$date_factor)
levels(np_ds1$date_factor)


np_ds1$yrmm_factor <- as_factor(as.character(np_ds1$yrmm))
head(np_ds1$yrmm_factor)
levels(np_ds1$yrmm_factor)

# filter(dg_sf1, (!is.na(log_n2o))

np_ds1 <- filter(np_ds1, (!is.na(long)))
# drops to 10,369 obs and summary(mp_ds2$lat) returns no NA

unique(np_ds1$organization_name)
distinct(np_ds1,bay) %>%  dplyr::arrange(., bay) %>% print(n=Inf)
# 35 distinct bay entries
distinct(np_ds1, monitoring_location_id) %>%  dplyr::arrange(., ) %>% print(n=Inf)
# 130 distinct monitoring location id entries

z1 <- filter(np_ds1, is.na(organization_name))
# return 7,162 obs
summary(z1$sta_nr) # code min 1.1 to max 9.5
# believe having a numeric sta_nr code means this is a Lakewatch/Baywatch station and have long/lat coordinates for these stations

z2 <- filter(np_ds1, organization_name == "FLORIDA LAKEWATCH")
# return 2,380
summary(z2$sta_nr) # returns all NAs
# want to get numeric codes assigned to sta_nr, struggling to get that done


```

## Explore Duplicates and NAs

```{r dupes_na}
# z1 <- get_dupes(np_ds1)
# 
# z2 <- filter(np_ds1, (is.na(tn) & !is.na(tp)))
# 
# z3 <- filter(np_ds1, is.na(tn))
```

## Create sf object

Have NA for some lat and long entries so remove these. Also, Numeric Nutrient Criteria (NNC) polygons omit bayous, creeks, and lakes, but includes intracoastal waterway and St. Joe Bay. But bayous, bays, and lakes does have all those polygons. Although, a Burnt Mill Creek station does not hit that polygon.

```{r create_sf}

# create sf object
coords <- data.frame(longitude  = np_ds1$long, latitude = np_ds1$lat)
head(coords)

# See Marc Weber workshop above about Coordinate Reference Systems (CRS)

np_sf1 <- st_as_sf(np_ds1, coords = c("long", "lat"), crs =4269) %>% st_transform((5070))

class(np_sf1)
# read in shapefile
nnc_1 <- st_read("data_gis/NNC_Bay_Segments.shp")
class(nnc_1)
summary(nnc_1)
distinct(nnc_1, SEGMENT_NA)
distinct(nnc_1, WATERBODY_)

mapview(nnc_1)
# Numeric

# see that espg code is 3087
# asking do the two sf objects have the same CRS
st_crs(np_sf1) == st_crs(nnc_1)

np_sf1 <- np_sf1 %>% st_transform(3087)
st_crs(np_sf1) == st_crs(nnc_1)

mapview(np_sf1) + mapview(nnc_1)

# mapview(hab_sf1, zcol = "Depth", cex = "Depth", alpha.regions = .8, legend = TRUE, popup = popupTable(hab_sf1, zcol = c("Depth", "Site"))) + mapview(bigbend_seagrass)


```

## Read Bays, Bayous and Lakes V4

Rose Horn took 2 polygons for Alligator Bayou, which borders the U.S. Coast Guard Station on St. Andrew Bay, and splits them into separate sampling areas. Also, in version the new update reflects the removal of small edge portions of St. Andrew Bay that were along the edges of East Bay and Grand Lagoon. Looks like I should pick-up Upper and Lower Burnt Mill Creek Stations.

Need to see if I have 8 polygon names for: East Bay, Grand Lagoon, North Bay, St. Andrew Bay, West Bay, Crooked and Burnt Mill Creeks, and Deer Point Lake as I want to use those 6 polygons in linked micromap. Perhaps by using SEGMENT_NA and Major_Seg I can get the polygons I want.

```{r bbl_v4}

bbl_4 <- st_read("data_gis/Bays_Bayous_Lakes_V4/Bays_Bayous_Lakes_V4.shp")
summary(bbl_4)
mapview(bbl_4)

bbl_4 <- bbl_4 %>% st_transform(3087)

distinct(bbl_4, SEGMENT_NA) %>% dplyr::arrange(., SEGMENT_NA)
# have 190 distinct SEGMENT_NA, but 191 obs suggesting one polygon is sharing a name

# Is Major_Seg a new field from Rose?
distinct(bbl_4, Major_Seg) %>% dplyr::arrange(., Major_Seg)
# returns 8 Major_Seg


mapview(np_sf1) + mapview(bbl_4)
# a burnt mill creek station from 2024-04-25 is off polygon

# these 8 polygons might be what I use for a linked micromap
polygons8 <- dplyr::filter(bbl_4, SEGMENT_NA == "West Bay" | SEGMENT_NA == "North Bay" | SEGMENT_NA == "St. Andrew Bay" | SEGMENT_NA == "East Bay" | SEGMENT_NA == "Burnt Mill Creek" | SEGMENT_NA == "Crooked Creek" | SEGMENT_NA == "Grand Lagoon" |SEGMENT_NA == "Deer Point Lake")

mapview(polygons8)

mapview(np_sf1) + mapview(polygons8)

```

## Stations for Linked Micromap

Inconsistent data entry for Lake Marin or just Marin. The "mm\_" naming convention refers to getting the data structured for a linked micromap.

```{r micromap_stations}

mm_station1 <- dplyr::filter(np_sf1, bay != "Lake Powell" & bay != "Powell Creek" & bay != "Lullwater" & bay != "Lake Marin" & bay != "Marin")

names(mm_station1)

mapview(mm_station1)

distinct(mm_station1, bay) %>% dplyr::arrange(., bay) %>% print(n=Inf) 

# np_ds1 %>% 
#   group_by(organization_name) %>% 
#   summarise(count = n_distinct(monitoring_location_name))

mapview(mm_station1) + mapview(polygons8, col.regions = "yellow")

```

## Spatial Intersection of MM Stations to Polygons 8

Spatial intersection so station points get polygon attributes, specifically SEGMENT_NA. For linked micromap the data frame of summary statistics has to have the same ID linking variable as in the polygon data frame. Note this intersection will drop many bayou and creek stations, but I am okay as this is just for a draft linked micromap.

In West Bay, the spatial intersection omits the intracoastal waterway, and upper stations in Crooked and Burnt Mill Creeks. In North Bay, omit many of the bayous, and likewise for East Bay.

```{r mmstations_intersect_polygons}

mm_station2 <- st_intersection(mm_station1, polygons8)

mm_stat1map <- mapview(mm_station1)
mm_stat2map <- mapview(mm_station2, col.regions = "yellow")

sync(mm_stat1map, mm_stat2map, ncol = 1)

mapview(mm_station2) + mapview(polygons8, col.regions = "yellow")

names(mm_station2)

mm_station2 %>% 
  group_by(SEGMENT_NA) %>% 
  summarize(., count = dplyr::n())

mm_station2 %>% 
  group_by(SEGMENT_NA, year) %>% 
  summarize(., count = dplyr::n()) %>% 
  print(n=Inf)

summary(mm_station2$sta_nr)
# returns 1727 NAs so would likely need to do nearest neighbor analysis to get numeric codes assigned to sta_nr

# na's for sta_nr appear in 2016 and later years
# will need to work this out so can get an accurate count of distinct within each
# of the polygons
mm_station2 %>% 
  group_by(SEGMENT_NA, year) %>% 
  summarize(., na_count = sum(is.na(sta_nr))) %>% 
print(n = Inf)

# np_ds1 %>% 
#   group_by(organization_name) %>% 
#   summarise(count = n_distinct(monitoring_location_name))

# dplyr::select(sub_np_stations, tn, SEGMENT_NA) %>% 
# group_by(., SEGMENT_NA) %>% 
# summarize(., min = fivenum(tn)[1],
#               q1 = fivenum(tn)[2],
#               median = fivenum(tn)[3],
#               q3 = fivenum(tn)[4],
#               max = fivenum(tn)[5],
#               count = dplyr::n(),
#               mean = mean(tn, na.rm = TRUE),
#               na_count = sum(is.na(tn))) %>% 
# print(n = Inf)
distinct(mm_station2, SEGMENT_NA) %>% dplyr::arrange(., SEGMENT_NA)

distinct(mm_station2, organization_id)
distinct(mm_station2, organization_name)
distinct(mm_station2, year) %>% dplyr::arrange(., year) %>% print(n=Inf)
distinct(mm_station2, sampling_agency_name)

#2019-2023 covers St. Andrew Baywatch WQ report
dplyr::filter(mm_station2, year == 2020 & sampling_agency_name == "ST. ANDREW BAY RESOURCE MANAGEMENT ASSOCIATION INC.") %>% 
dplyr::select(., tn, SEGMENT_NA) %>% 
group_by(., SEGMENT_NA) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)


# get years that match St. Andrew Baywatch WQ Report
mm_station3 <- dplyr::filter(mm_station2, sampling_agency_name == "ST. ANDREW BAY RESOURCE MANAGEMENT ASSOCIATION INC.")
# returns 1082 obs

distinct(mm_station3, organization_id)
distinct(mm_station3, organization_name)
distinct(mm_station3, sampling_agency_name)
distinct(mm_station3, year)

# mm_stations3$yr <- as.factor(year)

# note in 2018 facet no East Bay samples for tn, tp and chl
# for chl all missing values in 2021 and some in 2023 and 2024
tm_shape(polygons8) +
  tm_polygons(fill = "white", lwd = 2) +
  tm_title("Detectable and NA Entries of Total Nitrogen (mg/L)")  + 
  tm_shape(mm_station3) +
  tm_symbols(fill = "tn") +
  tm_facets(by = "year")
```

## Statistics for Linked Micromaps

Get 5-number summary statistics to use in making boxplot for linked micromaps. Also, getting stats on log-transformed data.

```{r stats_micromap}
# remove geometry from summary stats
chla_2020 <- st_set_geometry(mm_station2, NULL) %>% 
dplyr::filter(., year == 2020 & sampling_agency_name == "ST. ANDREW BAY RESOURCE MANAGEMENT ASSOCIATION INC.") %>% 
dplyr::select(., chl, SEGMENT_NA) %>% 
group_by(., SEGMENT_NA) %>% 
summarize(., chla_min = fivenum(chl)[1],
              chla_q1 = fivenum(chl)[2],
              chla_med = fivenum(chl)[3],
              chla_q3 = fivenum(chl)[4],
              chla_max = fivenum(chl)[5],
              chla_lmin = log(fivenum(chl)[1]),
              chla_lq1 = log(fivenum(chl)[2]),
              chla_lmed = log(fivenum(chl)[3]),
              chla_lq3 = log(fivenum(chl)[4]),
              chla_lmax = log(fivenum(chl)[5]),
              chla_count = dplyr::n(),
              chla_mean = mean(chl, na.rm = TRUE),
              chla_na_count = sum(is.na(chl))) %>% 
print(n = Inf)

tn_2020 <- st_set_geometry(mm_station2, NULL) %>% 
  dplyr::filter(., year == 2020 & sampling_agency_name == "ST. ANDREW BAY RESOURCE MANAGEMENT ASSOCIATION INC.") %>% 
  dplyr::select(., tn, SEGMENT_NA) %>% 
  group_by(., SEGMENT_NA) %>% 
  summarize(., tn_min = fivenum(tn)[1],
            tn_q1 = fivenum(tn)[2],
            tn_med = fivenum(tn)[3],
            tn_q3 = fivenum(tn)[4],
            tn_max = fivenum(tn)[5],
            tn_lmin = log(fivenum(tn)[1]),
            tn_lq1 = log(fivenum(tn)[2]),
            tn_lmed = log(fivenum(tn)[3]),
            tn_lq3 = log(fivenum(tn)[4]),
            tn_lmax = log(fivenum(tn)[5]),
            tn_count = dplyr::n(),
            tn_mean = mean(tn, na.rm = TRUE),
            tn_na_count = sum(is.na(tn))) %>% 
  print(n = Inf)

tp_2020 <- st_set_geometry(mm_station2, NULL) %>% 
  dplyr::filter(., year == 2020 & sampling_agency_name == "ST. ANDREW BAY RESOURCE MANAGEMENT ASSOCIATION INC.") %>% 
  dplyr::select(., tp, SEGMENT_NA) %>% 
  group_by(., SEGMENT_NA) %>% 
  summarize(., tp_min = fivenum(tp)[1],
            tp_q1 = fivenum(tp)[2],
            tp_med = fivenum(tp)[3],
            tp_q3 = fivenum(tp)[4],
            tp_max = fivenum(tp)[5],
            tp_lmin = log(fivenum(tp)[1]),
            tp_lq1 = log(fivenum(tp)[2]),
            tp_lmed = log(fivenum(tp)[3]),
            tp_lq3 = log(fivenum(tp)[4]),
            tp_lmax = log(fivenum(tp)[5]),
            tp_count = dplyr::n(),
            tp_mean = mean(tp, na.rm = TRUE),
            tp_na_count = sum(is.na(tp))) %>% 
  print(n = Inf)



```

## Edit Polygons and Stats

Edit data, both polygons and stats, to just the areas of interest. Dropping Burnt Mill Creek and Crooked Creek polygons and stats as have too few observations. Also, merge the chlorophyll, tn, and tp statistics to the polygon data as a single sf object.

```{r polygons_stats}
names(polygons8)
# View(polygons8)

polygons6 <- filter(polygons8, SEGMENT_NA != "Crooked Creek" & SEGMENT_NA != "Burnt Mill Creek")

#chla
chla_2020v1 <- filter(chla_2020, SEGMENT_NA != "Crooked Creek")
names(chla_2020v1)

# merge polygon to chla stats
# list sf object first, then dataframe of stats to create polygons with stats
chla_2020v2 <- merge(polygons6, chla_2020v1, by = "SEGMENT_NA" )
summary(chla_2020v2)
# note that geometry is polygon

# tn
tn_2020v1 <- filter(tn_2020, SEGMENT_NA != "Crooked Creek")

chlatn_2020v2 <- merge(chla_2020v2, tn_2020v1, by = "SEGMENT_NA" )

# tp
tp_2020v1 <- filter(tp_2020, SEGMENT_NA != "Crooked Creek")

nutr_2020 <- merge(chlatn_2020v2, tp_2020v1, by = "SEGMENT_NA" )

```

## Draft Linked Micromap

Make draft linked micromaps to get code working, and note any statistical and spatial patterns.

For 2020 data, tp, tn, and chla all have NA entries.

```{r draft_linked_micromap}
#Draft plot

mmplot(map.data=nutr_2020,
       panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 3,
       ord.by='tp_lmed', rev.ord=TRUE,
       median.row=FALSE)

mmplot(map.data=nutr_2020,
       panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       list('tn_lmin', 'tn_lq1', 'tn_lmed', 'tn_lq3', 'tn_lmax'),
                       NA),
       grouping = 3,
       ord.by='tn_lmed', rev.ord=TRUE,
       median.row=FALSE)


mmplot(map.data=nutr_2020,
       panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       list('chla_lmin', 'chla_lq1', 'chla_lmed', 'chla_lq3', 'chla_lmax'),
                       NA),
       grouping = 3,
       ord.by='chla_lmed', rev.ord=TRUE,
       median.row=FALSE)

```

## Refine Linked Micromaps

Use more detailed coding arguments to make better looking linked micromaps.

Refined TP See if tp scale from 2019 also works for 2020, and 2019 scale works.

```{r refine_micromap_tp}
# TP Refined Plot
mmplot(map.data=nutr_2020,
       panel.types=c('dot_legend', 'labels', 'labels', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       'tp_count',
                       'tp_na_count',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 3,
       ord.by='tp_lmed',
       rev.ord=TRUE,
       median.row=FALSE,
       plot.pGrp.spacing=1,
       
       plot.height=6, plot.width=9, #for poster use plot.height=6, plot.width=9,
       colors=brewer.pal(5, "Spectral"), #for journal use plot.height=9, plot.width=6.8
              
       panel.att=list(list(1, panel.width=1, point.type=20, point.size=2,
                           point.border=TRUE), 
                      
                      list(2, header='St.Andrew Bay Segments', panel.width=1.2,#increased panel.width to 1.1 from .8
                           align='left', text.size=.8),
                      
                      list(3, header='Total #\n of Samples',
                           panel.width=.35,
                           align='right',
                           text.size=.8),
                      
                      list(4, header='# of NA\n Samples',
                           panel.width=.4,
                           align='right',
                           text.size=.8),
                      
                      list(5, header='2020 Detectable Total Phoshporous',
                           graph.bgcolor='lightgray',
                           xaxis.ticks=c(-6.0, -5.0, -4.0, -3.0, -2.0),
                           xaxis.labels=c(-6,-5, -4, -3, -2),
                           xaxis.labels.size=1.25,
                           graph.bar.size = 0.5,
                           #xaxis.labels.angle=90,
                           xaxis.title='Log(TP, mg/L)', xaxis.title.size=1.25, panel.width=1.75),
                      
                      list(6, header='Light Gray Means Highlighted Above',
                           inactive.border.color=gray(.7), inactive.border.size=2,
                            panel.width=1.75)),
print.file = "output_micromap/mmplot_tp2020.jpeg", res=100)



```

Refined TN Looks like 2019 tn scale should work for 2020 tn data.

```{r refined_micromap_tn}
# TN Refined Plot
mmplot(map.data=nutr_2020,
       panel.types=c('dot_legend', 'labels', 'labels', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       'tn_count',
                       'tn_na_count',
                       list('tn_lmin', 'tn_lq1', 'tn_lmed', 'tn_lq3', 'tn_lmax'),
                       NA),
       grouping = 3,
       ord.by='tn_lmed',
       rev.ord=TRUE,
       median.row=FALSE,
       plot.pGrp.spacing=1,
       
       plot.height=6, plot.width=9, #for poster use plot.height=6, plot.width=9,
       colors=brewer.pal(5, "Spectral"), #for journal use plot.height=9, plot.width=6.8
              
       panel.att=list(list(1, panel.width=1, point.type=20, point.size=2,
                           point.border=TRUE), 
                      
                      list(2, header='St.Andrew Bay Segments', panel.width=1.2,#increased panel.width to 1.1 from .8
                           align='left', text.size=.8),
                      
                      list(3, header='Total #\n of Samples',
                           panel.width=.35,
                           align='right',
                           text.size=.8),
                      
                      list(4, header='# of NA\n Samples',
                           panel.width=.4,
                           align='right',
                           text.size=.8),
                      
                      list(5, header='2020 Detectable Total Nitrogen',
                           graph.bgcolor='lightgray',
                           xaxis.ticks=c(-2.5, -2.0, -1.5, -1.0, -0.5, -0.0),
                           xaxis.labels=c(-2.5, -2.0, -1.5, -1.0, -0.5, -0.0),
                           xaxis.labels.size=1.25,
                           graph.bar.size = 0.5,
                           #xaxis.labels.angle=90,
                           xaxis.title='Log(TN, mg/L)', xaxis.title.size=1.25, panel.width=1.75),
                      
                      list(6, header='Light Gray Means Highlighted Above',
                           inactive.border.color=gray(.7), inactive.border.size=2,
                            panel.width=1.75)),
# to run so just appears in Plots pane have 3 parentheses after the last panel.width
# to run and save the image delete 3rd parenthesis and insert comma       
print.file = "output_micromap/mmplot_tn2020.jpeg", res=100)


```

Refined Chlorophyll a Note need to add column "\# of NA Samples". Try 2019 scale first, but might need to drop it down as max log of chla in 2020 is near 3.0

```{r refined_micromap_chla}
# Chla Refined Plot
mmplot(map.data=nutr_2020,
       panel.types=c('dot_legend', 'labels', 'labels', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       'chla_count',
                       'chla_na_count',
                       list('chla_lmin', 'chla_lq1', 'chla_lmed', 'chla_lq3', 'chla_lmax'),
                       NA),
       grouping = 3,
       ord.by='chla_lmed',
       rev.ord=TRUE,
       median.row=FALSE,
       plot.pGrp.spacing=1,
       
       plot.height=6, plot.width=9, #for poster use plot.height=6, plot.width=9,
       colors=brewer.pal(5, "Spectral"), #for journal use plot.height=9, plot.width=6.8
              
       panel.att=list(list(1, panel.width=1, point.type=20, point.size=2,
                           point.border=TRUE), 
                      
                      list(2, header='St.Andrew \n Bay Segments', panel.width=1,#increased panel.width to 1.1 from .8
                           align='left', text.size=.8),
                      
                      list(3, header='Total #\n of Samples',
                           panel.width=.5,
                           align='right',
                           text.size=.8),
                      
                      list(4, header='# of NA\n Samples',
                           panel.width=.4,
                           align='right',
                           text.size=.8),
                      
                      list(5, header='2020 Detectable\n Chlorophyll a',
                           graph.bgcolor='lightgray',
                           xaxis.ticks=c(0,1.5,3,4.5),
                           xaxis.labels=c(0,1.5,3,4.5),
                           xaxis.labels.size=1.25,
                           graph.bar.size = 0.5,
                           #xaxis.labels.angle=90,
                           xaxis.title='Log(µg/L)', xaxis.title.size=1.25, panel.width=1.75),
                      
                      list(6, header='Light Gray Means\n Highlighted Above',
                           inactive.border.color=gray(.7), inactive.border.size=2,
                            panel.width=1.75)),
# to run so just appears in Plots pane have 3 parentheses after the last panel.width
# to run and save the image replace 3rd parenthesis with comma       
print.file = "output_micromap/mmplot_chla2020.jpeg", res=100)


```

## West Bay, Bayous, and Creeks

The two code chunks below illustrate 2 ways to format data to make linked micromaps. The first way involves filtering from the points/stations data so it only has points within the polygons of interest. Also, the filtered data need to have variable name, in this case "bay", to match to the polgyon names entered under SEGMENT_NA. A five-number summary is made of the WQ data grouped by bay. Next, the appropriate polygons are selected and, if needed, values under SEGMENT_NA are replaced to match the bay entries of the points. The polygon data and five-number summary data are merged using the shared variable names of SEGMENT_NA and bay.

Mapvview of mm_stations1 shows monitoring location id BAY-WE-STBAY-5 clearly in SAB, but its bay is labeled as West Bay. Since December 1, 2018, BAY-WE-STBAY-5 has had either the wrong coordinate entered or the wrong bay or both, and it has 28 entries.

Note monitoring location id is all upper case so always use that. Monitoring location name is snake-cased so avoid it. Note subtle behavior of filter and filter_out with NAs, <https://dplyr.tidyverse.org/reference/filter.html>.

```{r wb_stats_spatial_join}

mm_station1 <- dplyr::filter(np_sf1, bay == "Crooked Creek" | bay == "Burnt Mill Creek" | bay == "Warren Bayou" | bay == "West Bay")
# return 1778 observations, of which 1418 observations have sta_nr numeric entry, and 360 have monitoring location id entry

distinct(mm_station1, monitoring_location_id) %>% arrange(., monitoring_location_id) %>% print(n=Inf)
# returns 21 observations including NA, 3 Burnt Mill stations, 3 Crooked stations, 1 Warren Bayou, 11 BAY-WE-STBAY stations, 1 BAY-WEST-BAY-12, and 1 NWFYI0005

ggplot(mm_station1 |> filter(year >=2020), aes(x=yrmm_factor, y=log(tp))) + geom_point() + facet_wrap(vars(monitoring_location_id))

mapview(mm_station1) + bayous_wb1

odd_coordinates <- dplyr::filter(mm_station1, monitoring_location_id == "BAY-WE-STBAY-3" | monitoring_location_id == "BAY-WE-STBAY-5")

mapview(odd_coordinates) + bayous_wb1

mm_station1 <- dplyr::filter_out(mm_station1, monitoring_location_id == "BAY-WE-STBAY-5")
# believe this filtering out removes all 28 obs for monitoring location BAY-WE-STBAY-5 so now have 1750 obs

distinct(mm_station1, monitoring_location_id) %>% arrange(., monitoring_location_id) %>% print(n=Inf)
# now returns to 20 obs and no BAY-WE-STBAY-5 observation


mm_station1 %>% 
  group_by(bay) %>% 
  summarize(., count = dplyr::n())
# want to add a bay entry of "West Bay ICWW" for BAY-WEST BAY-12


mm_station2 <- mm_station1 |>
  mutate(
    bay = case_when(monitoring_location_id == "BAY-WEST-BAY-12" ~ "West Bay ICWW",
                       TRUE ~ bay
                       )
    )

mm_station2 %>% 
  group_by(bay) %>% 
  summarize(., count = dplyr::n())

bayous_wb1 <- dplyr::filter(bbl_4, SEGMENT_NA == "Burnt Mill Creek" | SEGMENT_NA == "Crooked Creek" | SEGMENT_NA == "Warren Bayou" | SEGMENT_NA == "Gulf ICWW between Choctawhatchee Bay and St. Andrew Bay" | SEGMENT_NA == "West Bay")


# do replace values so SEGMENT_NA will match to bay in points data frame
bayous_wb2 <- bayous_wb1 |> 
  mutate(
  SEGMENT_NA = SEGMENT_NA |>
    replace_values(
      "Gulf ICWW between Choctawhatchee Bay and St. Andrew Bay" ~ "West Bay ICWW"
    )
  )  

# getting 5-number summary from points data
tp_wb2020 <- st_set_geometry(mm_station2, NULL) %>% 
  dplyr::filter(., year == 2020 & sampling_agency_name == "ST. ANDREW BAY RESOURCE MANAGEMENT ASSOCIATION INC.") %>% 
  dplyr::select(., tp, bay) %>% 
  group_by(., bay) %>% 
  summarize(., tp_min = fivenum(tp)[1],
            tp_q1 = fivenum(tp)[2],
            tp_med = fivenum(tp)[3],
            tp_q3 = fivenum(tp)[4],
            tp_max = fivenum(tp)[5],
            tp_lmin = log(fivenum(tp)[1]),
            tp_lq1 = log(fivenum(tp)[2]),
            tp_lmed = log(fivenum(tp)[3]),
            tp_lq3 = log(fivenum(tp)[4]),
            tp_lmax = log(fivenum(tp)[5]),
            tp_count = dplyr::n(),
            tp_mean = mean(tp, na.rm = TRUE),
            tp_na_count = sum(is.na(tp))) %>% 
  print(n = Inf)


tp_wb2020v2 <- merge(bayous_wb2, tp_wb2020, by.x = "SEGMENT_NA", by.y = "bay" )


mmplot(map.data=tp_wb2020v2,
       panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 1,
       ord.by='tp_lmed', rev.ord=TRUE,
       median.row=FALSE)


# TP Refined Plot for West Bay Creeks and Bayous
mmplot(map.data=tp_wb2020v2,
       panel.types=c('dot_legend', 'labels', 'labels', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       'tp_count',
                       'tp_na_count',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 2,
       ord.by='tp_lmed',
       rev.ord=TRUE,
       median.row=FALSE,
       plot.pGrp.spacing=1,
       
       plot.height=6, plot.width=9, #for poster use plot.height=6, plot.width=9,
       colors=brewer.pal(4, "Spectral"), #for journal use plot.height=9, plot.width=6.8
              
       panel.att=list(list(1, panel.width=1, point.type=20, point.size=2,
                           point.border=TRUE), 
                      
                      list(2, header='West Bay Segment,\n Creeks & Bayous', panel.width=1.2,#increased panel.width to 1.1 from .8
                           align='left', text.size=.8),
                      
                      list(3, header='Total #\n of Samples',
                           panel.width=.35,
                           align='right',
                           text.size=.8),
                      
                      list(4, header='# of NA\n Samples',
                           panel.width=.4,
                           align='right',
                           text.size=.8),
                      
                      list(5, header='2020 Detectable Total Phoshporous',
                           graph.bgcolor='lightgray',
                           xaxis.ticks=c(-6.0, -5.0, -4.0, -3.0, -2.0),
                           xaxis.labels=c(-6,-5, -4, -3, -2),
                           xaxis.labels.size=1.25,
                           graph.bar.size = 0.5,
                           #xaxis.labels.angle=90,
                           add.line=-4.07, add.line.col="black",
                           xaxis.title='Log(TP, mg/L)', xaxis.title.size=1.25, panel.width=1.75),
                      
                      list(6, header='Light Gray Means Highlighted Above',
                           inactive.border.color=gray(.7), inactive.border.size=2,
                            panel.width=1.75)),
print.file = "output_micromap/mmplot_tp_wb2020v2.jpeg", res=100)



```

The second way to format the data assumes the points and polgyon datasets are in exact alignment. That is every point/station is in its appropriate polygon. A spatial intersection is done of the points/stations to the polygons so that now all the polygon attributes, such as SEGMENT_NA, are assigned to the points. After this spatial intersection is done, then a five-number summary of the WQ data can be made and the linked micromaps created.

Spatial intersection ends up omitting Bay-West Bay-3 from April 25, 2024 as that monitoring location does not intersect polygons for some reason. Need to think of summary statistics for Burnt Mill Creek, Crooked Creek and GICWW-West given their small sample sizes relative to West Bay main segment.

Also, Warren Bayou falls in polygon of West Bay segment so that's why Warren Bayou not showing up in linked micromap. The spatial intersection I do puts the Warren Bayou point into the West Bay segment. Code is included to measure how far the Warren Bayou station is from the Warren Bayou polygon.

```{r wb_stats_spatial_interseciton}

bayous_wb1 <- dplyr::filter(bbl_4, SEGMENT_NA == "Burnt Mill Creek" | SEGMENT_NA == "Crooked Creek" | SEGMENT_NA == "Warren Bayou" | SEGMENT_NA == "Gulf ICWW between Choctawhatchee Bay and St. Andrew Bay" | SEGMENT_NA == "West Bay")

# mm_station1 <- dplyr::filter(np_sf1, bay != "Lake Powell" & bay != "Powell Creek" & bay != "Lullwater" & bay != "Lake Marin" & bay != "Marin")


warren_v1 <- dplyr::filter(mm_station1, bay == "Warren Bayou")
#returns 162 observations
mapview(warren_v1)
summary(warren_v1$tp)
summary(warren_v1$tn)
summary(warren_v1$chl)

# pull out only geometry of Warren Bayou station
warren_v2 <- st_geometry(warren_v1) %>% unique.data.frame(.)
class(warren_v2)

st_crs(warren_v2) == st_crs(bayous_wb1)

# https://stackoverflow.com/questions/46214479/distance-between-a-set-of-points-and-a-polygon-with-sf-in-r
st_distance(warren_v2, bayous_wb1)
# returns 5 distances for each of the 5 polygons in bayous_wb1 and distances are in order of the polygons. Have a 0 m distance for West Bay as the warren_v2 point is in that polygon. The distance to the 3rd polygon, SEGMENT_NA is Warren Bayou, is 7.0 m.

mm_wbstation2 <- st_intersection(mm_station1, bayous_wb1)


mapview(mm_wbstation2) + mapview(bayous_wb1, col.regions = "yellow")

names(mm_wbstation2)

mm_wbstation2 %>% 
  group_by(SEGMENT_NA) %>% 
  summarize(., count = dplyr::n())

mm_wbstation2 %>% 
  group_by(SEGMENT_NA, year) %>% 
  summarize(., count = dplyr::n()) %>% 
  print(n=Inf)

# uses 2020 data as have samples from all 4 polygons

tp_wb2020 <- st_set_geometry(mm_wbstation2, NULL) %>% 
  dplyr::filter(., year == 2020 & sampling_agency_name == "ST. ANDREW BAY RESOURCE MANAGEMENT ASSOCIATION INC.") %>% 
  dplyr::select(., tp, SEGMENT_NA) %>% 
  group_by(., SEGMENT_NA) %>% 
  summarize(., tp_min = fivenum(tp)[1],
            tp_q1 = fivenum(tp)[2],
            tp_med = fivenum(tp)[3],
            tp_q3 = fivenum(tp)[4],
            tp_max = fivenum(tp)[5],
            tp_lmin = log(fivenum(tp)[1]),
            tp_lq1 = log(fivenum(tp)[2]),
            tp_lmed = log(fivenum(tp)[3]),
            tp_lq3 = log(fivenum(tp)[4]),
            tp_lmax = log(fivenum(tp)[5]),
            tp_count = dplyr::n(),
            tp_mean = mean(tp, na.rm = TRUE),
            tp_na_count = sum(is.na(tp))) %>% 
  print(n = Inf)

tp_wb2020v2 <- merge(bayous_wb1, tp_wb2020, by = "SEGMENT_NA" )
summary(tp_wb2020v2)

tp_wb2020v3 <- tp_wb2020v2 |> 
  mutate(
  SEGMENT_NA = SEGMENT_NA |>
    replace_values(
      "Gulf ICWW between Choctawhatchee Bay and St. Andrew Bay" ~ "West Bay ICWW"
    )
  )  
    

mmplot(map.data=tp_wb2020v3,
       panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 1,
       ord.by='tp_lmed', rev.ord=TRUE,
       median.row=FALSE)


# TP Refined Plot for West Bay Creeks and Bayous
mmplot(map.data=tp_wb2020v3,
       panel.types=c('dot_legend', 'labels', 'labels', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       'tp_count',
                       'tp_na_count',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 2,
       ord.by='tp_lmed',
       rev.ord=TRUE,
       median.row=FALSE,
       plot.pGrp.spacing=1,
       
       plot.height=6, plot.width=9, #for poster use plot.height=6, plot.width=9,
       colors=brewer.pal(4, "Spectral"), #for journal use plot.height=9, plot.width=6.8
              
       panel.att=list(list(1, panel.width=1, point.type=20, point.size=2,
                           point.border=TRUE), 
                      
                      list(2, header='West Bay Segment,\n Creeks & Bayous', panel.width=1.2,#increased panel.width to 1.1 from .8
                           align='left', text.size=.8),
                      
                      list(3, header='Total #\n of Samples',
                           panel.width=.35,
                           align='right',
                           text.size=.8),
                      
                      list(4, header='# of NA\n Samples',
                           panel.width=.4,
                           align='right',
                           text.size=.8),
                      
                      list(5, header='2020 Detectable Total Phoshporous',
                           graph.bgcolor='lightgray',
                           xaxis.ticks=c(-6.0, -5.0, -4.0, -3.0, -2.0),
                           xaxis.labels=c(-6,-5, -4, -3, -2),
                           xaxis.labels.size=1.25,
                           graph.bar.size = 0.5,
                           #xaxis.labels.angle=90,
                           add.line=-4.07, add.line.col="black",
                           xaxis.title='Log(TP, mg/L)', xaxis.title.size=1.25, panel.width=1.75),
                      
                      list(6, header='Light Gray Means Highlighted Above',
                           inactive.border.color=gray(.7), inactive.border.size=2,
                            panel.width=1.75)),
print.file = "output_micromap/mmplot_tp_wb2020.jpeg", res=100)

```

## North Bay and Bayous

```{r nb_stats_spatial_join}

```

## Station Issues
Specific follow-up needed for some stations, such as coordinates that put a station on land, monitoring location ID BAY-WE-STBAY-3 near Burnt Mill Creek. Also, monitoring location ID BAY-WE-STBAY-5 that has a bay assignment of West Bay but coordinates put it in St. Andrew Bay.
