---
title: "ESDA of St. Andrew Bay Watershed Water Quality"
author: "Mike McManus and Ruth Horn"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Exploratory Spatial Data Analysis

The purpose of this script to examine statistical and spatial patterns in the water qualiy of St. Andrews Bay (SAB) watershed using monitoring data collected by several agencies and organizations.

## Libraries

```{r libraries}
library(tidyverse)
library(sf)
library(mapview)
library(janitor)

```

## Log

2025-11-22 Unsure the DEP longitude and latitude are actual coordinates of monitoring stations

2025-11-21 Bring in csv files

## Reading in Data

Two csv file, one on nutrients and other on conductivity, provided by Rose Horn, SASJBEP, by email on November 13, 2025. Rose pulled the data from WIN. Note activity_start_date comes in as character.

```{r data}
np_ds1 <- read_csv("data/WQ_Values_Station_Baysection_NP.csv")

spec(np_ds1)
names(np_ds1)
np_ds1 <- clean_names(np_ds1)

# See Long and Lat in decimal degrees
View(np_ds1) 
str(np_ds1)
glimpse(np_ds1)

np_ds1$activity_start_date <- lubridate::mdy(np_ds1$activity_start_date)
str(np_ds1)

# create sf object
coords <- data.frame(longitude  = np_ds1$dep_longitude, latitude = np_ds1$dep_latitude)
head(coords)

# See Marc Weber workshop above about Coordinate Reference Systems (CRS)

np_sf1 <- st_as_sf(np_ds1, coords = c("dep_longitude", "dep_latitude"), crs =4269) %>% st_transform((5070))

class(np_sf1)
# read in shapefile
nnc_1 <- st_read("data_gis/NNC_Bay_Segments.shp")
class(nnc_1)
summary(nnc_1)
# see that espg code is 3087
# asking do the two sf objects have the same CRS
st_crs(np_sf1) == st_crs(nnc_1)

np_sf1 <- np_sf1 %>% st_transform(3087)
st_crs(np_sf1) == st_crs(nnc_1)

mapview(np_sf1) + mapview(nnc_1)

# mapview(hab_sf1, zcol = "Depth", cex = "Depth", alpha.regions = .8, legend = TRUE, popup = popupTable(hab_sf1, zcol = c("Depth", "Site"))) + mapview(bigbend_seagrass)

```

## Data Structure and Summaries
Seven distinct organizations collected monitoring data with Florida Lakewatch being a synonym for St. Andrew Baywatch. Get 890 dates spanning 2013-01-24 to 2023-12-20. There are 7894 distinct coordinates/locations.

But, I am wondering if DEP assigns coordinates that are not the actual locations of the monitoring station, but instead are always to same place (the centroid) of the spatial feature.
```{r data_structure}
names(np_ds1)
distinct(np_ds1, organization_name)

distinct(np_ds1, activity_start_date ) %>% arrange(., activity_start_date) %>% 
  print(n = Inf)

np_ds1 %>% 
  group_by(organization_name) %>% 
  summarise(count = n_distinct(monitoring_location_name))



# distinct monitoring stations
np_stations <- unique.data.frame(np_sf1) # returns 7894 obs/stations & geometries

np_stations %>% 
  group_by(organization_name) %>% 
  summarise(count = n_distinct(monitoring_location_name))


```

The `echo: false` option disables the printing of code (only output is displayed).
