---
title: "Bay-Segment-Specific Linked Micromaps TP 2020"
author: "Mike McManus"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Libraries

```{r libraries}
library(tidyverse)
library(sf)
library(mapview)
library(janitor)
library(leaflet)
library(leafpop) # for popups in 
library(leafsync) # sync mapview maps
library(leafem) # add labels to map
library(patchwork)
library(nhdplusTools)
library(tmap)
library(forcats)
library(gifski) #for tmap animation
library(plotly)
library(webshot) # save mapview object
library(nngeo)
library(micromap)
library(tidyverse)
```

## Log

20206-07-29 bbl_v7 from Rose Horn was uploaded to repository, and methods and metadata were uploaded to Methods folder.

2026-07-20 bbl_v5 that Rose Horn uploaded to repository. No changes to refined West Bay micromap based on bbl_v5. The all polygons West Bay micromap now has Johnson Bayou_WestBay label and polygon displayed.

2026-06-25 update West Bay polygons to include bayous that do not have samples as I did with other bay segments. Need to check bayous in other bay segments following this approach: got names below by filtering on Major_Seg West Bay in View mode of bbl_v4

2026-05-30 calculated distances from Warren Bayou point/station to polygons in bayous_wb1

2026-05-06 renamed with YYYY in script name

## Reading in Data and Formatting Date Variables

One csv file, win_baywatch_sabnut.csv, uploaded by Ryann Rossi, SASJBEP, on December 23, 2025. Ryan pulled the data from WIN and did some data cleaning. Note activity_start_date comes in as character.

Read in Bays_Bayous_Lakes_V4 shapefile that Rose uploaded to repository. Can I merge wq data, np_ds1, with Baywatch station IDs, bw_id1, by coordinates?

```{r data}
np_ds1 <- read_csv("data/win_baywatch_sabnut.csv")
names(np_ds1)
np_ds1 <- clean_names(np_ds1)
# spec(np_ds1)
# names(np_ds1)


# See Long and Lat in decimal degrees and date as character
View(np_ds1) 
str(np_ds1)
glimpse(np_ds1)
class(np_ds1$date)

#zdates <- distinct(np_ds1, activity_start_date)
#ym_ds1 <- lubridate::mdy(zdates$activity_start_date)
#floor_date(ym_ds1, "month")
#saveRDS(zdates, "zdates")

# with activity start date still character have to assign factors
# have to create activity start date as factor for tmap animation
np_ds1$date_factor <- np_ds1$date
np_ds1$yrmm_factor <- np_ds1$date

# convert to date
np_ds1$date <- lubridate::mdy(np_ds1$date)
class(np_ds1$date)
np_ds1$yrmm <- floor_date(np_ds1$date, "month")
head(np_ds1$yrmm)
class(np_ds1$yrmm)

# have to create activity start date as factor for tmap animation
np_ds1 <- arrange(np_ds1, date)
np_ds1$date_factor <- as_factor(np_ds1$date_factor)
head(np_ds1$date_factor)
levels(np_ds1$date_factor)


np_ds1$yrmm_factor <- as_factor(as.character(np_ds1$yrmm))
head(np_ds1$yrmm_factor)
levels(np_ds1$yrmm_factor)

# filter(dg_sf1, (!is.na(log_n2o))

np_ds1 <- filter(np_ds1, (!is.na(long)))
# drops to 10,369 obs and summary(mp_ds2$lat) returns no NA

unique(np_ds1$organization_name)
distinct(np_ds1,bay) %>%  dplyr::arrange(., bay) %>% print(n=Inf)
# 35 distinct bay entries
distinct(np_ds1, monitoring_location_id) %>%  dplyr::arrange(., ) %>% print(n=Inf)
# 130 distinct monitoring location id entries

z1 <- filter(np_ds1, is.na(organization_name))
# return 7,162 obs
summary(z1$sta_nr) # code min 1.1 to max 9.5
# believe having a numeric sta_nr code means this is a Lakewatch/Baywatch station and have long/lat coordinates for these stations

z2 <- filter(np_ds1, organization_name == "FLORIDA LAKEWATCH")
# return 2,380
summary(z2$sta_nr) # returns all NAs
# want to get numeric codes assigned to sta_nr, struggling to get that done


```

## Explore Duplicates and NAs

```{r dupes_na}
# z1 <- get_dupes(np_ds1)
# 
# z2 <- filter(np_ds1, (is.na(tn) & !is.na(tp)))
# 
# z3 <- filter(np_ds1, is.na(tn))
```

## Create sf object

Have NA for some lat and long entries so remove these. Also, Numeric Nutrient Criteria (NNC) polygons omit bayous, creeks, and lakes, but includes intracoastal waterway and St. Joe Bay. But bayous, bays, and lakes does have all those polygons. Although, a Burnt Mill Creek station does not hit that polygon.

```{r create_sf}

# create sf object
coords <- data.frame(longitude  = np_ds1$long, latitude = np_ds1$lat)
head(coords)

# See Marc Weber workshop above about Coordinate Reference Systems (CRS)

np_sf1 <- st_as_sf(np_ds1, coords = c("long", "lat"), crs =4269) %>% st_transform((5070))

class(np_sf1)
# read in shapefile
nnc_1 <- st_read("data_gis/NNC_Bay_Segments.shp")
class(nnc_1)
summary(nnc_1)
distinct(nnc_1, SEGMENT_NA)
distinct(nnc_1, WATERBODY_)

mapview(nnc_1)
# Numeric

# see that espg code is 3087
# asking do the two sf objects have the same CRS
st_crs(np_sf1) == st_crs(nnc_1)

np_sf1 <- np_sf1 %>% st_transform(3087)
st_crs(np_sf1) == st_crs(nnc_1)

mapview(np_sf1) + mapview(nnc_1)

# mapview(hab_sf1, zcol = "Depth", cex = "Depth", alpha.regions = .8, legend = TRUE, popup = popupTable(hab_sf1, zcol = c("Depth", "Site"))) + mapview(bigbend_seagrass)


```

## Read Bays, Bayous and Lakes V4
Using bbl_7 from Rose Horn.

Using bbl_5 that Rose Horn sent, which has adjustments to distinguish polygons that had the same name. Still wonder why have Shape_Leng, Shape_Le_1 and Shape_Le_2.

Rose Horn took 2 polygons for Alligator Bayou, which borders the U.S. Coast Guard Station on St. Andrew Bay, and splits them into separate sampling areas. Also, in version the new update reflects the removal of small edge portions of St. Andrew Bay that were along the edges of East Bay and Grand Lagoon. Looks like I should pick-up Upper and Lower Burnt Mill Creek Stations.

Need to see if I have 8 polygon names for: East Bay, Grand Lagoon, North Bay, St. Andrew Bay, West Bay, Crooked and Burnt Mill Creeks, and Deer Point Lake as I want to use those 6 polygons in linked micromap. Perhaps by using SEGMENT_NA and Major_Seg I can get the polygons I want.

```{r bbl_v7}

bbl_7 <- st_read("data_gis/Bays_Bayous_Lakes_V7.shp")

bbl_7 <- bbl_7 %>% st_transform(3087)

summary(bbl_7)
mapview(bbl_7)


distinct(bbl_7, SEGMENT_NA) %>% dplyr::arrange(., SEGMENT_NA)
# now with bbl_7 have 194 distinct segment names and 194 obs, I believe, because Harrison Bayou was split in to Harrison Bayou and Little Johnson Bayou. with bbl_5 have 193 distinct segment names and 193 obs. Previously in bbl_4 had 190 distinct SEGMENT_NA, but 191 obs suggesting one polygon is sharing a name

distinct(bbl_7, Major_Seg) %>% dplyr::arrange(., Major_Seg)
# returns 8 Major_Seg


mapview(np_sf1) + mapview(bbl_7)
# a burnt mill creek station from 2024-04-25 is off polygon

# these 8 polygons might be what I use for a linked micromap
# polygons8 <- dplyr::filter(bbl_5, SEGMENT_NA == "West Bay" | SEGMENT_NA == "North Bay" | SEGMENT_NA == "St. Andrew Bay" | SEGMENT_NA == "East Bay" | SEGMENT_NA == "Burnt Mill Creek" | SEGMENT_NA == "Crooked Creek" | SEGMENT_NA == "Grand Lagoon" |SEGMENT_NA == "Deer Point Lake")
# 
# mapview(polygons8)

# mapview(np_sf1) + mapview(polygons8)

```

## West Bay, Bayous, and Creeks

Note that in bbl_7, both Harrison Bayou and Little Johnson Bayou in West Bay have same area. I think Rose Horn may have forgotten to recalculate areas once she edited the original Harrison Bayou.

The two code chunks below illustrate 2 ways to format data to make linked micromaps. The first way involves filtering from the points/stations data so it only has points within the polygons of interest. Also, the filtered data need to have variable name, in this case "bay", to match to the polgyon names entered under SEGMENT_NA. A five-number summary is made of the WQ data grouped by bay. Next, the appropriate polygons are selected and, if needed, values under SEGMENT_NA are replaced to match the bay entries of the points. The polygon data and five-number summary data are merged using the shared variable names of SEGMENT_NA and bay.

Mapvview of mm_stations1 shows monitoring location id BAY-WE-STBAY-5 clearly in SAB, but its bay is labeled as West Bay. Since December 1, 2018, BAY-WE-STBAY-5 has had either the wrong coordinate entered or the wrong bay or both, and it has 28 entries.

Note monitoring location id is all upper case so always use that. Monitoring location name is snake-cased so avoid it. Note subtle behavior of filter and filter_out with NAs, <https://dplyr.tidyverse.org/reference/filter.html>.

With bbl_5,believe Johnson Bayou_WestBay also needs an edit of Major Seg from, currently, St. Andrew Bay to West Bay. Also, I list a Little Johnson Bayou, but no such name occurs in filtered polygons. Did Little Johnson Bayou appear in Google Maps? Yes, it does, and Little Johnson Bayou in Google Maps is adjacent to Harrison Bayou. But in bbl_5, Harrison Bayou polygon incorporates Little Johnson Bayou.

```{r wb_stats_spatial_join}

wb_station1 <- dplyr::filter(np_sf1, bay == "Crooked Creek" | bay == "Burnt Mill Creek" | bay == "Warren Bayou" | bay == "West Bay")
# return 1778 observations, of which 1418 observations have sta_nr numeric entry, and 360 have monitoring location id entry

distinct(wb_station1, monitoring_location_id) %>% arrange(., monitoring_location_id) %>% print(n=Inf)
# returns 21 observations including NA, 3 Burnt Mill stations, 3 Crooked stations, 1 Warren Bayou, 11 BAY-WE-STBAY stations, 1 BAY-WEST-BAY-12, and 1 NWFYI0005

ggplot(wb_station1 |> filter(year >=2020), aes(x=yrmm_factor, y=log(tp))) + geom_point() + facet_wrap(vars(monitoring_location_id))

# could high and variable Burnt Mill and Crooked Creeks in 2020-10-01 be due to Hurricane Sally in September of that year
ggplot(wb_station1 |> filter(year >=2020), aes(y=bay, x=log(tp))) + geom_boxplot() + facet_wrap(vars(yrmm_factor))

mapview(wb_station1)

# check script 03 2020 b/c I thought BAY-WE-STBAY-5 had a bay assignment of SAB there, but here it has bay assignment of West Bay
odd_coordinates <- dplyr::filter(wb_station1, monitoring_location_id == "BAY-WE-STBAY-3" | monitoring_location_id == "BAY-WE-STBAY-5")

mapview(odd_coordinates)

wb_station2 <- dplyr::filter_out(wb_station1, monitoring_location_id == "BAY-WE-STBAY-5")
# believe this filtering out removes all 28 obs for monitoring location BAY-WE-STBAY-5 so now have 1750 obs

distinct(wb_station2, monitoring_location_id) %>% arrange(., monitoring_location_id) %>% print(n=Inf)
# now returns to 20 obs and no BAY-WE-STBAY-5 observation

ggplot(wb_station2 |> filter(year >=2020), aes(x=yrmm_factor, y=log(tp))) + geom_point() + facet_wrap(vars(monitoring_location_id))

ggplot(wb_station2 |> filter(year >=2020), aes(y=bay, x=log(tp))) + geom_boxplot() + facet_wrap(vars(yrmm_factor))


wb_station2 %>% 
  group_by(bay) %>% 
  summarize(., count = dplyr::n())
# want to add a bay entry of "West Bay ICWW" for BAY-WEST BAY-12

# evaluate samples within bays sampled in 2020
wb_station2 %>% 
  dplyr::filter(., year == 2020) %>% 
  group_by(bay) %>% 
  summarize(., count = dplyr::n())

# evaluate which years have the most samples
wb_station2 %>% 
  dplyr::filter(., year >= 2020) %>% 
  group_by(year) %>% 
  summarize(., count = dplyr::n())


wb_station3 <- wb_station2 |>
  mutate(
    bay = case_when(monitoring_location_id == "BAY-WEST-BAY-12" ~ "West Bay ICWW",
                       TRUE ~ bay
                       )
    )

wb_station3 %>% 
  group_by(bay) %>% 
  summarize(., count = dplyr::n())
#now have 5 bays

# got names below by filtering on Major_Seg West Bay in View mode of bbl_v7 and adding Gulf ICCW to list
bayous_wb1 <- dplyr::filter(bbl_7, SEGMENT_NA == "Burnt Mill Creek" | SEGMENT_NA == "Crooked Creek" | SEGMENT_NA == "Warren Bayou" | SEGMENT_NA == "Gulf ICWW between Choctawhatchee Bay and St. Andrew Bay" | SEGMENT_NA == "West Bay" | SEGMENT_NA == "Johnson Bayou_WestBay" | SEGMENT_NA == "Basin Bayou" | SEGMENT_NA == "Botheration Bayou" | SEGMENT_NA == "Doyle Bayou" | SEGMENT_NA == "Harrison Bayou" | SEGMENT_NA == "Little Johnson Bayou" |SEGMENT_NA == "Ward Creek")
# returns 12 obs matching 12 segment names
mapview(bayous_wb1)
# no longer see other Johnson Bayou in St. Andrew Bay

mapview(wb_station3) + bayous_wb1
# still see sta_nr 2.5, which is BAY-WE-STBAY-5, b/c of 2006-12-01 collection but will loose that when I filter for year == 2020 in 5-number summary stats

# do replace values so SEGMENT_NA will match to bay in points data frame
bayous_wb2 <- bayous_wb1 |> 
  mutate(
  SEGMENT_NA = SEGMENT_NA |>
    replace_values(
      "Gulf ICWW between Choctawhatchee Bay and St. Andrew Bay" ~ "West Bay ICWW"
    )
  )  

# getting 5-number summary from points data
tp_wb2020 <- st_set_geometry(wb_station3, NULL) %>% 
  dplyr::filter(., year == 2020 & sampling_agency_name == "ST. ANDREW BAY RESOURCE MANAGEMENT ASSOCIATION INC.") %>% 
  dplyr::select(., tp, bay) %>% 
  group_by(., bay) %>% 
  summarize(., tp_min = fivenum(tp)[1],
            tp_q1 = fivenum(tp)[2],
            tp_med = fivenum(tp)[3],
            tp_q3 = fivenum(tp)[4],
            tp_max = fivenum(tp)[5],
            tp_lmin = log(fivenum(tp)[1]),
            tp_lq1 = log(fivenum(tp)[2]),
            tp_lmed = log(fivenum(tp)[3]),
            tp_lq3 = log(fivenum(tp)[4]),
            tp_lmax = log(fivenum(tp)[5]),
            tp_count = dplyr::n(),
            tp_mean = mean(tp, na.rm = TRUE),
            tp_na_count = sum(is.na(tp))) %>% 
  print(n = Inf)

# use left_join as have a many polygons (n=12) to few stats (n=5 obs) and want to keep all the polygons whether polygons have stats or not
tp_wb2020v1 <- left_join(bayous_wb2, tp_wb2020, by = join_by(SEGMENT_NA == bay), keep = TRUE)

# use right_join so cut down polygons to n=5 to match stats
tp_wb2020v2 <- right_join(bayous_wb2, tp_wb2020, by = join_by(SEGMENT_NA == bay), keep = TRUE)

# this micromap shows Little John Bayou
mmplot(map.data=tp_wb2020v1,
       panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 5,
       ord.by='tp_lmed', rev.ord=TRUE,
       median.row=FALSE)

mmplot(map.data=tp_wb2020v2,
       panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 1,
       ord.by='tp_lmed', rev.ord=TRUE,
       median.row=FALSE)



# TP Refined Plot for West Bay Creeks and Bayous
mmplot(map.data=tp_wb2020v2,
       panel.types=c('dot_legend', 'labels', 'labels', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       'tp_count',
                       'tp_na_count',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 2,
       ord.by='tp_lmed',
       rev.ord=TRUE,
       median.row=FALSE,
       plot.pGrp.spacing=1,
       
       plot.height=6, plot.width=9, #for poster use plot.height=6, plot.width=9,
       colors=brewer.pal(4, "Spectral"), #for journal use plot.height=9, plot.width=6.8
              
       panel.att=list(list(1, panel.width=1, point.type=20, point.size=2,
                           point.border=TRUE), 
                      
                      list(2, header='West Bay Segment,\n Creeks & Bayous', panel.width=1,#increased panel.width to 1.1 from .8
                           align='left', text.size=.8),
                      
                      list(3, header='Total #\n of Samples',
                           panel.width=.45,
                           align='right',
                           text.size=.8),
                      
                      list(4, header='# of NA\n Samples',
                           panel.width=.4,
                           align='right',
                           text.size=.8),
                      
                      list(5, header='2020 Detectable Total Phoshporous',
                           graph.bgcolor='lightgray',
                           xaxis.ticks=c(-6.0, -5.0, -4.0, -3.0, -2.0),
                           xaxis.labels=c(-6,-5, -4, -3, -2),
                           xaxis.labels.size=1.25,
                           graph.bar.size = 0.5,
                           #xaxis.labels.angle=90,
                           add.line=-4.07, add.line.col="black",
                           xaxis.title='Log(TP, mg/L)', xaxis.title.size=1.25, panel.width=1.75),
                      
                      list(6, header='Light Gray Means Highlighted Above',
                           inactive.border.color=gray(.7), inactive.border.size=2,
                            panel.width=1.75)),
print.file = "output_micromap/mmplot_tp_wb2020v4.jpeg", res=100)



```

I do not recommend this second method as it assumes points and polygons are exactly geographically matched. Consequently, I have commented out this code chunk. The second way to format the data assumes the points and polgyon datasets are in exact alignment. That is every point/station is in its appropriate polygon. A spatial intersection is done of the points/stations to the polygons so that now all the polygon attributes, such as SEGMENT_NA, are assigned to the points. After this spatial intersection is done, then a five-number summary of the WQ data can be made and the linked micromaps created.

Spatial intersection ends up omitting Bay-West Bay-3 from April 25, 2024 as that monitoring location does not intersect polygons for some reason. Need to think of summary statistics for Burnt Mill Creek, Crooked Creek and GICWW-West given their small sample sizes relative to West Bay main segment.

Also, Warren Bayou falls in polygon of West Bay segment so that's why Warren Bayou not showing up in linked micromap. The spatial intersection I do puts the Warren Bayou point into the West Bay segment. Code is included to measure how far the Warren Bayou station is from the Warren Bayou polygon.

```{r wb_stats_spatial_interseciton}

# bayous_wb1 <- dplyr::filter(bbl_4, SEGMENT_NA == "Burnt Mill Creek" | SEGMENT_NA == "Crooked Creek" | SEGMENT_NA == "Warren Bayou" | SEGMENT_NA == "Gulf ICWW between Choctawhatchee Bay and St. Andrew Bay" | SEGMENT_NA == "West Bay")
# 
# # mm_station1 <- dplyr::filter(np_sf1, bay != "Lake Powell" & bay != "Powell Creek" & bay != "Lullwater" & bay != "Lake Marin" & bay != "Marin")
# 
# 
# warren_v1 <- dplyr::filter(mm_station1, bay == "Warren Bayou")
# #returns 162 observations
# mapview(warren_v1)
# summary(warren_v1$tp)
# summary(warren_v1$tn)
# summary(warren_v1$chl)
# 
# # pull out only geometry of Warren Bayou station
# warren_v2 <- st_geometry(warren_v1) %>% unique.data.frame(.)
# class(warren_v2)
# 
# st_crs(warren_v2) == st_crs(bayous_wb1)
# 
# # https://stackoverflow.com/questions/46214479/distance-between-a-set-of-points-and-a-polygon-with-sf-in-r
# st_distance(warren_v2, bayous_wb1)
# # returns 5 distances for each of the 5 polygons in bayous_wb1 and distances are in order of the polygons. Have a 0 m distance for West Bay as the warren_v2 point is in that polygon. The distance to the 3rd polygon, SEGMENT_NA is Warren Bayou, is 7.0 m.
# 
# mm_wbstation2 <- st_intersection(mm_station1, bayous_wb1)
# 
# 
# mapview(mm_wbstation2) + mapview(bayous_wb1, col.regions = "yellow")
# 
# names(mm_wbstation2)
# 
# mm_wbstation2 %>% 
#   group_by(SEGMENT_NA) %>% 
#   summarize(., count = dplyr::n())
# 
# mm_wbstation2 %>% 
#   group_by(SEGMENT_NA, year) %>% 
#   summarize(., count = dplyr::n()) %>% 
#   print(n=Inf)
# 
# # uses 2020 data as have samples from all 4 polygons
# 
# tp_wb2020 <- st_set_geometry(mm_wbstation2, NULL) %>% 
#   dplyr::filter(., year == 2020 & sampling_agency_name == "ST. ANDREW BAY RESOURCE MANAGEMENT ASSOCIATION INC.") %>% 
#   dplyr::select(., tp, SEGMENT_NA) %>% 
#   group_by(., SEGMENT_NA) %>% 
#   summarize(., tp_min = fivenum(tp)[1],
#             tp_q1 = fivenum(tp)[2],
#             tp_med = fivenum(tp)[3],
#             tp_q3 = fivenum(tp)[4],
#             tp_max = fivenum(tp)[5],
#             tp_lmin = log(fivenum(tp)[1]),
#             tp_lq1 = log(fivenum(tp)[2]),
#             tp_lmed = log(fivenum(tp)[3]),
#             tp_lq3 = log(fivenum(tp)[4]),
#             tp_lmax = log(fivenum(tp)[5]),
#             tp_count = dplyr::n(),
#             tp_mean = mean(tp, na.rm = TRUE),
#             tp_na_count = sum(is.na(tp))) %>% 
#   print(n = Inf)
# 
# tp_wb2020v2 <- merge(bayous_wb1, tp_wb2020, by = "SEGMENT_NA" )
# summary(tp_wb2020v2)
# 
# tp_wb2020v3 <- tp_wb2020v2 |> 
#   mutate(
#   SEGMENT_NA = SEGMENT_NA |>
#     replace_values(
#       "Gulf ICWW between Choctawhatchee Bay and St. Andrew Bay" ~ "West Bay ICWW"
#     )
#   )  
#     
# 
# mmplot(map.data=tp_wb2020v3,
#        panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
#        panel.data=list(NA,
#                        'SEGMENT_NA',
#                        list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
#                        NA),
#        grouping = 1,
#        ord.by='tp_lmed', rev.ord=TRUE,
#        median.row=FALSE)
# 
# 
# # TP Refined Plot for West Bay Creeks and Bayous
# mmplot(map.data=tp_wb2020v3,
#        panel.types=c('dot_legend', 'labels', 'labels', 'labels', 'box_summary', 'map'),
#        panel.data=list(NA,
#                        'SEGMENT_NA',
#                        'tp_count',
#                        'tp_na_count',
#                        list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
#                        NA),
#        grouping = 2,
#        ord.by='tp_lmed',
#        rev.ord=TRUE,
#        median.row=FALSE,
#        plot.pGrp.spacing=1,
#        
#        plot.height=6, plot.width=9, #for poster use plot.height=6, plot.width=9,
#        colors=brewer.pal(4, "Spectral"), #for journal use plot.height=9, plot.width=6.8
#               
#        panel.att=list(list(1, panel.width=1, point.type=20, point.size=2,
#                            point.border=TRUE), 
#                       
#                       list(2, header='West Bay Segment,\n Creeks & Bayous', panel.width=1.2,#increased panel.width to 1.1 from .8
#                            align='left', text.size=.8),
#                       
#                       list(3, header='Total #\n of Samples',
#                            panel.width=.35,
#                            align='right',
#                            text.size=.8),
#                       
#                       list(4, header='# of NA\n Samples',
#                            panel.width=.4,
#                            align='right',
#                            text.size=.8),
#                       
#                       list(5, header='2020 Detectable Total Phoshporous',
#                            graph.bgcolor='lightgray',
#                            xaxis.ticks=c(-6.0, -5.0, -4.0, -3.0, -2.0),
#                            xaxis.labels=c(-6,-5, -4, -3, -2),
#                            xaxis.labels.size=1.25,
#                            graph.bar.size = 0.5,
#                            #xaxis.labels.angle=90,
#                            add.line=-4.07, add.line.col="black",
#                            xaxis.title='Log(TP, mg/L)', xaxis.title.size=1.25, panel.width=1.75),
#                       
#                       list(6, header='Light Gray Means Highlighted Above',
#                            inactive.border.color=gray(.7), inactive.border.size=2,
#                             panel.width=1.75)),
# print.file = "output_micromap/mmplot_tp_wb2020.jpeg", res=100)

```

## North Bay, Bayous, and Deer Point Lake
Using bbl_7 to make linked micromap.

Does 2020-10-01 boxplot capture effects of Hurricane Sally?

Seems like losing Upper Goose Bay when I summarize by bay, and that's because Goose Bay station and Upper Goose Bay station are grouped together under bay as Goose Bay polygon. From polygons, I need to select all the North Bay bayous to show how many there are and to show which ones have been sampled (n=6) and not sampled (n=17).

Very ugly draft linked micromap shows that 1) Upper Goose Bay station stats should be assigned to Upper Goose Bay polygon, and 2) that 11 bayous in North Bay do not have any samples.

If used FDEP data would pick-up a lot more stations in bayous upstream of Bailey Bridge and downstream of dam/bridge.

Filtering by Deer Point Lake picks up St. Andrew Bay Resource Management Association stations, NWFWMD/FDEP probabilistic stations (including a Lake Martin station), and FDEP Northwest Regional Office HAB station. The probabilistic stations are from 2011, 2013-2018.

In 2020, St. Andrew Bay Resource Management Association sampled Deer Point Lake in March-September, November, and December, with only 2 stations sampled in August so 26 observations. As have a single polygon, Deer Point Lake, making a linked micromap is not appropriate so I combined it with North Bay. In Deer Point Lake station 1 is furthest downstream, closest to dam, and station 3 is furthest upstream just above where Bayou George enters lake.

Likely worthwhile to see if FDEP probabilistic large river and small stream status monitoring has sampling points on Econfina Creek, Bayou George, Bear Creek, and Cedar Creek.

In bbl_5, Massalina Bayou polygon is mistakenly assigned to North Bay. Also, have an FDEP large lake station point in Massalina Bayou and it is has a bay assignment of North Bay.

Also, need to bring both Robinson polygon and Robinson Bayou station into North Bay.

```{r nbdp_stats_spatial_join}
nbdp_station1 <- dplyr::filter(np_sf1, bay == "Robinson Bayou" | bay == "Goose Bayou" | bay == "Upper Goose Bayou" | bay == "Fannin Bayou" | bay == "North Bay" | bay == "Deer Point Lake")
# return 1531 observations

distinct(nbdp_station1, monitoring_location_id) %>% arrange(., monitoring_location_id) %>% print(n=Inf)
# returns 31 observations including 1 NA, 15 numerically coded stations, 3 Deer Point Lake stations, 1 Fannin Bayou,  1 Goose Bayou station, 4 North Bay stations, 1 Robinson Bayou, 1 Upper Goose Bayou station, and 3 FDEP Stations, G3TLHR0021, G3TLHR0031, and G3TLHR0055


nbdp_station1 %>% 
  group_by(bay) %>% 
  summarize(., count = dplyr::n())

nbdp_station1 %>% 
  group_by(sampling_agency_name) %>% 
  summarize(., count = dplyr::n())


# evaluate samples within bays sampled in 2020
nbdp_station1 %>% 
  dplyr::filter(., year == 2020) %>% 
  group_by(bay) %>% 
  summarize(., count = dplyr::n())

# evaluate which years have the most samples
nbdp_station1 %>% 
  dplyr::filter(., year >= 2020) %>% 
  group_by(year) %>% 
  summarize(., count = dplyr::n())


ggplot(nbdp_station1 |> filter(year >=2020), aes(x=yrmm_factor, y=log(tp))) + geom_point() + facet_wrap(vars(monitoring_location_id))

ggplot(nbdp_station1 |> filter(year >=2020), aes(y=bay, x=log(tp))) + geom_boxplot() + facet_wrap(vars(yrmm_factor))

mapview(nbdp_station1, col.regions = "red") + bbl_7

# shows odd coordinate placement for monitoring location id 57719, a FDEP Large Lake sample supposedly for Lake Martin and bay was assigned as Deer Point Lake. This station point will be filtered out by sampling_agency_name == "ST. ANDREW BAY RESOURCE MANAGEMENT ASSOCIATION INC."

odd_coords_nb <- dplyr::filter(nbdp_station1, monitoring_location_id == "57719")
 
mapview(odd_coords_nb, col.regions = "red") + bbl_7
# 
# mm_station1 <- dplyr::filter_out(mm_station1, monitoring_location_id == "BAY-WE-STBAY-5")
# believe this filtering out removes all 28 obs for monitoring location BAY-WE-STBAY-5 so now have 1750 obs

distinct(nbdp_station1, monitoring_location_id) %>% arrange(., monitoring_location_id) %>% print(n=Inf)
# still returns 31 obs

# this creates an Upper Goose Bayou so stats can be assigned to that station
nbdp_station2 <- nbdp_station1 |>
  mutate(
    bay = case_when(monitoring_location_id == "BAY-UPPER-GOOSE-BAYOU-1" ~ "Upper Goose Bayou",
                       TRUE ~ bay
                       )
    )

nbdp_station2 %>% 
  dplyr::filter(., year == 2020) %>% 
  group_by(bay) %>% 
  summarize(., count = dplyr::n())
# now returns 6 rows/obs of summary stats for 6 bays

# filter for North Bay and include Robinson Bay, which is near North Bay and St. Andrew Bay border
bayous_nb1 <- dplyr::filter(bbl_7, Major_Seg == "North Bay" | SEGMENT_NA == "Robinson Bayou")
# using bbl_7 returns 17 obs/polygon


# this method too laborious and missed some polygons, such as Maxwell and McKitchens
# filter all North Bay polygons
# bayous_nb1 <- dplyr::filter(bbl_4, SEGMENT_NA == "Fannin Bayou" | SEGMENT_NA == "Goose Bayou" | SEGMENT_NA == "Robinson Bayou" | SEGMENT_NA == "North Bay" | SEGMENT_NA == "Mud Bayou" | SEGMENT_NA == "Alligator Bayou_GeneratingPlant" | SEGMENT_NA == "Newman Bayou" | SEGMENT_NA == "Sheephead Bayou" | SEGMENT_NA == "Gainer Bayou" | SEGMENT_NA == "Hodges Bayou" |  SEGMENT_NA == "Mill Bayou" | SEGMENT_NA == "Beatty Bayou" | SEGMENT_NA == "Anderson Bayou" | SEGMENT_NA == "Upper Goose Bayou" | SEGMENT_NA == "Lynn Haven Bayou")

mapview(nbdp_station2, col.regions = "green") + mapview(bayous_nb1)
# shows odd Massalina station

# do replace values so SEGMENT_NA will match to bay in points data frame
# bayous_wb2 <- bayous_wb1 |> 
#   mutate(
#   SEGMENT_NA = SEGMENT_NA |>
#     replace_values(
#       "Gulf ICWW between Choctawhatchee Bay and St. Andrew Bay" ~ "West Bay ICWW"
#     )
#   )  

# getting 5-number summary from points data
tp_nbdp2020 <- st_set_geometry(nbdp_station2, NULL) %>% 
  dplyr::filter(., year == 2020 & sampling_agency_name == "ST. ANDREW BAY RESOURCE MANAGEMENT ASSOCIATION INC.") %>% 
  dplyr::select(., tp, bay) %>% 
  group_by(., bay) %>% 
  summarize(., tp_min = fivenum(tp)[1],
            tp_q1 = fivenum(tp)[2],
            tp_med = fivenum(tp)[3],
            tp_q3 = fivenum(tp)[4],
            tp_max = fivenum(tp)[5],
            tp_lmin = log(fivenum(tp)[1]),
            tp_lq1 = log(fivenum(tp)[2]),
            tp_lmed = log(fivenum(tp)[3]),
            tp_lq3 = log(fivenum(tp)[4]),
            tp_lmax = log(fivenum(tp)[5]),
            tp_count = dplyr::n(),
            tp_mean = mean(tp, na.rm = TRUE),
            tp_na_count = sum(is.na(tp))) %>% 
  print(n = Inf)
# returns 6 obs/rows of summary stats for 6 bays

# use left_join as have a many polygons (n=11) to few stats (n=6 obs) and want to keep all the polygons whether polygons have stats or not
tp_nbdp2020v2 <- left_join(bayous_nb1, tp_nbdp2020, by = join_by(SEGMENT_NA == bay), keep = TRUE)

# use right_join so cut down polygons to n=6 to match stats
tp_nbdp2020v3 <- right_join(bayous_nb1, tp_nbdp2020, by = join_by(SEGMENT_NA == bay), keep = TRUE)

# using tp_nbdp2020v2 shows how 11 bayous that do not have samples
mmplot(map.data=tp_nbdp2020v2,
       panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 5,
       ord.by='tp_lmed', rev.ord=TRUE,
       median.row=FALSE)

# using tp_nbdp2020v3 shows only the 6 polygons with stats
mmplot(map.data=tp_nbdp2020v3,
       panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 3,
       ord.by='tp_lmed', rev.ord=TRUE,
       median.row=FALSE)


# TP Refined Plot for North Bay, North Bay Bayous, and Deer Point Lake
mmplot(map.data=tp_nbdp2020v3,
       panel.types=c('dot_legend', 'labels', 'labels', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       'tp_count',
                       'tp_na_count',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 3,
       ord.by='tp_lmed',
       rev.ord=TRUE,
       median.row=FALSE,
       plot.pGrp.spacing=1,
       
       plot.height=6, plot.width=9, #for poster use plot.height=6, plot.width=9,
       colors=brewer.pal(4, "Spectral"), #for journal use plot.height=9, plot.width=6.8
              
       panel.att=list(list(1, panel.width=1, point.type=20, point.size=2,
                           point.border=TRUE), 
                      
                      list(2, header='North Bay, Bayous\n & Deer Point Lake', panel.width=1,#increased panel.width to 1.1 from .8
                           align='left', text.size=.8),
                      
                      list(3, header='Total # of\n Samples',
                           panel.width=.5,
                           align='right',
                           text.size=.8),
                      
                      list(4, header='# of NA\n Samples',
                           panel.width=.4,
                           align='right',
                           text.size=.8),
                      
                      list(5, header='2020 Detectable Total Phoshporous',
                           graph.bgcolor='lightgray',
                           xaxis.ticks=c(-6.0, -5.0, -4.0, -3.0, -2.0),
                           xaxis.labels=c(-6,-5, -4, -3, -2),
                           xaxis.labels.size=1.25,
                           graph.bar.size = 0.5,
                           #xaxis.labels.angle=90,
                           add.line=-4.27, add.line.col="black",
                           xaxis.title='Log(TP, mg/L)', xaxis.title.size=1.25, panel.width=1.7),
                      
                      list(6, header='Light Gray Means Highlighted Above',
                           inactive.border.color=gray(.7), inactive.border.size=2,panel.width=1.7)),
print.file = "output_micromap/mmplot_tp_nbdp2020v9.jpeg", res=100)



```

## East Bay, Bayous, and Creeks

Filter bbl_7 by Major Segment for East Bay and get 42 polygons.
Removed Massalina Bayou station and Massalina polygon as they are assigned to be in St. Andrew Bayou major segment. Also, dropping Lake Martin station and polygon. Need to check with Ryan Rossi about that.

Consequently, the linked micromap will go from 8 boxplots and polygons down to six.
```{r eb_stats_spatial_join}

eb_station1 <- dplyr::filter(np_sf1, bay == "Watson Bayou" | bay == "Pitts Bayou"| bay == "Callaway Bayou" | bay == "Laird Bayou" | bay == "Sandy Creek" | bay == "Parker Bayou"| bay == "Eagle Nest Bayou" | bay == "East Bay" | monitoring_location_id == "NWFYI0018")
# returns 1733 obs

mapview(eb_station1, col.regions = "red") + bbl_7

distinct(eb_station1, monitoring_location_id) %>% arrange(., monitoring_location_id) %>% print(n=Inf)
# returns 22 observations including 1 NA, 2 Callaway Bayou stations,  5 East Bayou stations, 3 Laird Bayou stations, 1 Parker Bayou station, 1 Pitts Bayou Station, 1 Watson Bayou station, 7 FDEP Stations, that begin with G3TLHR, and 1 NWFY10018 station

ggplot(eb_station1 |> filter(year >=2020), aes(x=yrmm_factor, y=log(tp))) + geom_point() + facet_wrap(vars(monitoring_location_id))

ggplot(eb_station1 |> filter(year >=2020), aes(y=bay, x=log(tp))) + geom_boxplot() + facet_wrap(vars(yrmm_factor))

# evaluate samples within bays across all years
# note 2 samples from St. Andrew Bay
eb_station1 %>% 
  group_by(bay) %>% 
  summarize(., count = dplyr::n())

# evaluate samples within bays sampled in 2020
# note 1 sample from St. Andrew Bay
eb_station1 %>% 
  dplyr::filter(., year == 2020) %>% 
  group_by(bay) %>% 
  summarize(., count = dplyr::n())

# evaluate which years have the most samples
eb_station1 %>% 
  dplyr::filter(., year >= 2020) %>% 
  group_by(year) %>% 
  summarize(., count = dplyr::n())

# Bunkers Cove, Lannie Rowe, and Lake Martin dropped in bbl_7 and Massalina Bayou moved to St. Andrew Bay segement because close to border of SAB and East Bay. See metadata and methods from 07/28/2026 in Methods folder provided by Rose Horn

bayous_eb1 <- dplyr::filter(bbl_7, Major_Seg == "East Bay")
# returns 42 polygons

# this method to laborious
# bayous_eb1 <- dplyr::filter(bbl_7, SEGMENT_NA == "Auburn Cove" |SEGMENT_NA == "Baby Bayou" | SEGMENT_NA == "Baker Bayou"  | SEGMENT_NA == "Big Cove" | SEGMENT_NA == "Blind Alligator Bayou" | SEGMENT_NA == "Boggy Bayou"| SEGMENT_NA == "Boggy Creek" | SEGMENT_NA == "Burkett Bayou" | SEGMENT_NA == "California Bayou" | SEGMENT_NA == "Callaway Bayou" |  SEGMENT_NA == "Chatterson Bayou" | SEGMENT_NA == "Cooks Bayou" | SEGMENT_NA == "Dolan Bayou" | SEGMENT_NA == "Eagle Nest Bayou" | SEGMENT_NA == "East Bay" | SEGMENT_NA == "Emmons Bayou" | SEGMENT_NA == "Farmdale Bayou" | SEGMENT_NA == "Fred Bayou" | SEGMENT_NA == "Freshwater Bayou" | SEGMENT_NA == "Gulf ICWW between St. Andrew Bay and St. Joseph Bay"| SEGMENT_NA == "Higgs Bayou" | SEGMENT_NA == "Horseshoe Bayou" | SEGMENT_NA == "Laird Bayou" | SEGMENT_NA == "Lake Clair" |SEGMENT_NA == "Lake Van Vac" | SEGMENT_NA == "Lathrop Bayou" | SEGMENT_NA == "Murray Bayou" | SEGMENT_NA == "Parker Bayou" | SEGMENT_NA == "Pearl Bayou" | SEGMENT_NA == "Piclalene Bayou" | SEGMENT_NA == "Pitts Bayou" | SEGMENT_NA == "Pratt Bayou" |SEGMENT_NA == "Richard Bayou" |SEGMENT_NA == "Sandy Creek" |SEGMENT_NA == "SheepheadBayou_EastBay" |SEGMENT_NA == "Smack Bayou" |SEGMENT_NA == "Smith Bayou" | SEGMENT_NA == "Strange Bayou" | SEGMENT_NA == "Turtle Slough" | SEGMENT_NA == "Walker Bayou" | SEGMENT_NA == "Watson Bayou"| SEGMENT_NA == "Wetappo Creek")
# returns 42 obs 

mapview(bayous_eb1)

mapview(eb_station1, col.regions = "red") + mapview(bayous_eb1)

# use short name of East Bay ICWW for SEGMENT_NA
bayous_eb2 <- bayous_eb1 |>
  mutate(
  SEGMENT_NA = SEGMENT_NA |>
    replace_values(
      "Gulf ICWW between St. Andrew Bay and St. Joseph Bay" ~ "East Bay ICWW"
    )
  )


# 3 monitoring points/stations are assigned bay = Parker Bayou, but the 2 collected by FDEP TALLAHASSEE REGIONAL OPERATIONS CENTER, G3TLHR0018 and G3TLHR0043, appear to be in Callaway Bayou. G3TLHR0043 is assigned Parker Bayou @ Mouth While BAY-PARKER-BAYOU-1 appears to be in Parker Bayou monitoring location name. G3TLHR0018, monitoring_location_name of Parker Bayou Continuous Monitoring, but that actual polygon is Baby Bayou, an arm of Callaway Bayou.
odd_point <- filter(eb_station1, bay == "Parker Bayou")

mapview(odd_point)

# note that just have an outline of Parker Bayou polygon as most of that polygon is overlaid by East Bay Polygon
odd_polygon <- filter(bbl_7, SEGMENT_NA == "Parker Bayou")

# Rose Horned fixed polygon slivers from bbl_4. In bbl_5, now have a single polygon for Parker Bayou and a single polygon for Pratt Bayou
# show over a dozen polygon slivers along the margins of Parker Bayou. In Google Maps, looks like Parker Bayou should enclose embayments capture by a line from Parker Point to Donalson Point
mapview(odd_polygon)

# FDEP stations labeled as Parker Bayou excluded by filtering below

# getting 5-number summary from points data
tp_eb2020 <- st_set_geometry(eb_station1, NULL) %>% 
  dplyr::filter(., year == 2020 & sampling_agency_name == "ST. ANDREW BAY RESOURCE MANAGEMENT ASSOCIATION INC.") %>% 
  dplyr::select(., tp, bay) %>% 
  group_by(., bay) %>% 
  summarize(., tp_min = fivenum(tp)[1],
            tp_q1 = fivenum(tp)[2],
            tp_med = fivenum(tp)[3],
            tp_q3 = fivenum(tp)[4],
            tp_max = fivenum(tp)[5],
            tp_lmin = log(fivenum(tp)[1]),
            tp_lq1 = log(fivenum(tp)[2]),
            tp_lmed = log(fivenum(tp)[3]),
            tp_lq3 = log(fivenum(tp)[4]),
            tp_lmax = log(fivenum(tp)[5]),
            tp_count = dplyr::n(),
            tp_mean = mean(tp, na.rm = TRUE),
            tp_na_count = sum(is.na(tp))) %>% 
  print(n = Inf)
# returns 6 obs

# use left_join as have a many polygons (n=42) to few stats (n=6 obs) and want to keep all the polygons whether polygons have stats or not
tp_eb2020v2 <- left_join(bayous_eb2, tp_eb2020, by = join_by(SEGMENT_NA == bay), keep = TRUE)

# use right_join so cut down polygons to n=6 to match stats
tp_eb2020v3 <- right_join(bayous_eb2, tp_eb2020, by = join_by(SEGMENT_NA == bay), keep = TRUE)

# think there is a mismatch between Parker Bayou station coordinates and Parker Bayou polygon as originally had an empty row for Parker Bayou, which led to the investigation above

# need appropriate polygon for Parker Bayou as slivers are not visible

# using tp_eb2020v2 shows how 36 bayous that do not have samples
mmplot(map.data=tp_eb2020v2,
       panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 6,
       ord.by='tp_lmed', rev.ord=TRUE,
       median.row=FALSE)

# using tp_nb2020v3 shows only the 6 polygons with wq stats
mmplot(map.data=tp_eb2020v3,
       panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 3,
       ord.by='tp_lmed', rev.ord=TRUE,
       median.row=FALSE)


# using tp_eb2020v3 shows only the 6 polygons with stats
# TP Refined Plot for East Bay and Bayous
mmplot(map.data=tp_eb2020v3,
       panel.types=c('dot_legend', 'labels', 'labels', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       'tp_count',
                       'tp_na_count',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 3,
       ord.by='tp_lmed',
       rev.ord=TRUE,
       median.row=FALSE,
       plot.pGrp.spacing=1,
       
       plot.height=6, plot.width=9, #for poster use plot.height=6, plot.width=9,
       colors=brewer.pal(4, "Spectral"), #for journal use plot.height=9, plot.width=6.8
              
       panel.att=list(list(1, panel.width=1, point.type=20, point.size=2,
                           point.border=TRUE), 
                      
                      list(2, header='East Bay Segment,\n & Bayous', panel.width=1,#increased panel.width to 1.1 from .8
                           align='left', text.size=.8),
                      
                      list(3, header='Total #\n of Samples',
                           panel.width=.45,
                           align='right',
                           text.size=.8),
                      
                      list(4, header='# of NA\n Samples',
                           panel.width=.4,
                           align='right',
                           text.size=.8),
                      
                      list(5, header='2020 Detectable Total Phoshporous',
                           graph.bgcolor='lightgray',
                           xaxis.ticks=c(-6.0, -5.0, -4.0, -3.0, -2.0),
                           xaxis.labels=c(-6,-5, -4, -3, -2),
                           xaxis.labels.size=1.25,
                           graph.bar.size = 0.5,
                           #xaxis.labels.angle=90,
                           add.line=-4.13, add.line.col="black",
                           xaxis.title='Log(TP, mg/L)', xaxis.title.size=1.25, panel.width=1.75),
                      
                      list(6, header='Light Gray Means Highlighted Above',
                           inactive.border.color=gray(.7), inactive.border.size=2, panel.width=1.75)),
print.file = "output_micromap/mmplot_tp_eb2020v3.jpeg", res=100)

```

## St. Andrew Bay and Bayous

Do I need to do a spatial intesect of SAB polygon to monitoring_location_id == "BAY-WE-STBAY-5" so that station gets assigned to SAB and not West Bay?
Also, with bbl_7, Massalina Bay is assigned to St. Andrew Bay major segment.

Filtered bbl_5 by Major Segment equal St. Andrew Bay, and returned 10 polygons, which included St. Andrew Bay to St. Joseph Bay ICCW.

In time-series plot see decreasing spatial trend in Grand Lagoon 1 station, furthest from SAB, to Grand Lagoon station 5, closest to SAB. Which stations are upstream of Navy Blvd bridge?

Note 2 stations/points are in East Bay, one is a FDEP station, which will be filtered out in 5-number summary by sampling agency name, and the other is a Baywatch station, sta_nr 3.3, which will be filtered out by year in that summary.

```{r sab_stats_spatial_join}}

sab_station1 <- dplyr::filter(np_sf1, bay == "Johnson Bayou" | bay == "Posten Bayou" | bay == "Pretty Bayou" | bay == "Grand Lagoon"| bay == "St. Andrew Bay")
# returns 2786 obs

mapview(sab_station1)
# shows one station by Dupont Bridge and other station off of Bunkers Point near corner of East Beach Drive and Cherry Street, with both being in East Bay

sab_station2 <- dplyr::filter_out(sab_station1, monitoring_location_id == "NWFYI0018" | monitoring_location_id == "BAY-ST-EWBAY-3")
# returns 2753 obs as 2 obs from FDEP NORTHWEST REGIONAL OPERATIONS CENTER taken near DuPont Bridge removed and 31 obs of BAY-ST-EWBAY-3 removed

mapview(sab_station2, col.regions = "red") + bbl_5
# show sta_nr 3.3 from 2006 and NWFY0017 from 2020 in East Bay, and both will be removed in stats by filtering.

distinct(sab_station2, monitoring_location_id) %>% arrange(., monitoring_location_id) %>% print(n=Inf)
# returns 26 observations including 1 NA, 11 320 coded stations,  5 Grand Lagoon stations, 1 Johnson Bayou station, 1 Posten Bayou station, 1 Pretty Bayou station, 4 BAY-ST-EWBAY stations (1,2,6,7), 1 BAY-ST. ANDREW BAY-12 Station, and 1 NWFY10017 station

ggplot(sab_station2 |> filter(year >=2020), aes(x=yrmm_factor, y=log(tp))) + geom_point() + facet_wrap(vars(monitoring_location_id))

ggplot(sab_station2 |> filter(year >=2020), aes(y=bay, x=log(tp))) + geom_boxplot() + facet_wrap(vars(yrmm_factor))

# evaluate samples within bays across all years
sab_station2 %>% 
  group_by(bay) %>% 
  summarize(., count = dplyr::n())

# evaluate samples within bays sampled in 2020
sab_station2 %>% 
  dplyr::filter(., year == 2020) %>% 
  group_by(bay) %>% 
  summarize(., count = dplyr::n())

# evaluate which years have the most samples
sab_station2 %>% 
  dplyr::filter(., year >= 2020) %>% 
  group_by(year) %>% 
  summarize(., count = dplyr::n())


# appears to be 2 Johnson Bayous, one near power plant and other just east of Panama City marina
bayous_sab1 <- dplyr::filter(bbl_5, SEGMENT_NA == "Hog Island Sound" | SEGMENT_NA == "Johnson Bayou"| SEGMENT_NA == "Lake Huntington" | SEGMENT_NA == "Posten Bayou" | SEGMENT_NA == "Pretty Bayou" | SEGMENT_NA == "Alligator Bayou" | SEGMENT_NA == "Grand Lagoon" | SEGMENT_NA == "Crooked Island Sound" |SEGMENT_NA == "St. Andrew Bay")
# returns 9 obs

mapview(bayous_sab1)

mapview(sab_station2, col.regions = "green") + bayous_sab1

# getting 5-number summary from points data
tp_sab2020 <- st_set_geometry(sab_station2, NULL) %>% 
  dplyr::filter(., year == 2020 & sampling_agency_name == "ST. ANDREW BAY RESOURCE MANAGEMENT ASSOCIATION INC.") %>% 
  dplyr::select(., tp, bay) %>% 
  group_by(., bay) %>% 
  summarize(., tp_min = fivenum(tp)[1],
            tp_q1 = fivenum(tp)[2],
            tp_med = fivenum(tp)[3],
            tp_q3 = fivenum(tp)[4],
            tp_max = fivenum(tp)[5],
            tp_lmin = log(fivenum(tp)[1]),
            tp_lq1 = log(fivenum(tp)[2]),
            tp_lmed = log(fivenum(tp)[3]),
            tp_lq3 = log(fivenum(tp)[4]),
            tp_lmax = log(fivenum(tp)[5]),
            tp_count = dplyr::n(),
            tp_mean = mean(tp, na.rm = TRUE),
            tp_na_count = sum(is.na(tp))) %>% 
  print(n = Inf)
# returns 5 obs


# use left_join as have a many polygons (n=9) to few stats (n=5 obs) and want to keep all the polygons whether polygons have stats or not
tp_sab2020v1 <- left_join(bayous_sab1, tp_sab2020, by = join_by(SEGMENT_NA == bay), keep = TRUE)

# use right_join so cut down polygons to n=5 to match stats
tp_sab2020v2 <- right_join(bayous_sab1, tp_sab2020, by = join_by(SEGMENT_NA == bay), keep = TRUE)

# using tp_sab2020v1 shows how 4 bayous that do not have samples
mmplot(map.data=tp_sab2020v1,
       panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 5,
       ord.by='tp_lmed', rev.ord=TRUE,
       median.row=FALSE)

# using tp_sab2020v2 shows only the 5 polygons with wq stats
mmplot(map.data=tp_sab2020v2,
       panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 3,
       ord.by='tp_lmed', rev.ord=TRUE,
       median.row=FALSE)


# using tp_sab2020v2 shows only the 5 polygons with stats
# TP Refined Plot for St. Andrew Bay and Bayous
mmplot(map.data=tp_sab2020v2,
       panel.types=c('dot_legend', 'labels', 'labels', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       'tp_count',
                       'tp_na_count',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 3,
       ord.by='tp_lmed',
       rev.ord=TRUE,
       median.row=FALSE,
       plot.pGrp.spacing=1,
       
       plot.height=6, plot.width=9, #for poster use plot.height=6, plot.width=9,
       colors=brewer.pal(4, "Spectral"), #for journal use plot.height=9, plot.width=6.8
              
       panel.att=list(list(1, panel.width=1, point.type=20, point.size=2,
                           point.border=TRUE), 
                      
                      list(2, header='St. Andrew Bay Segment,\n & Bayous', panel.width=1,#increased panel.width to 1.1 from .8
                           align='left', text.size=.8),
                      
                      list(3, header='Total #\n of Samples',
                           panel.width=.45,
                           align='right',
                           text.size=.8),
                      
                      list(4, header='# of NA\n Samples',
                           panel.width=.4,
                           align='right',
                           text.size=.8),
                      
                      list(5, header='2020 Detectable Total Phoshporous',
                           graph.bgcolor='lightgray',
                           xaxis.ticks=c(-6.0, -5.0, -4.0, -3.0, -2.0),
                           xaxis.labels=c(-6,-5, -4, -3, -2),
                           xaxis.labels.size=1.25,
                           graph.bar.size = 0.5,
                           #xaxis.labels.angle=90,
                           add.line=-3.96, add.line.col="black",
                           xaxis.title='Log(TP, mg/L)', xaxis.title.size=1.25, panel.width=1.75),
                      
                      list(6, header='Light Gray Means Highlighted Above',
                           inactive.border.color=gray(.7), inactive.border.size=2,
                            panel.width=1.75)),
print.file = "output_micromap/mmplot_tp_sab2020v2.jpeg", res=100)



```

## Station and Polygon Issues

Only 2 issues with polygons in bbl_5. First, Massalina Bayou is mistakenly assigned to Major Segment of North Bay and should be corrected to be East Bay. Second, in West Bay Harrrison Bayou polygon should be split in two. The southern bayou is Harrison Bayou, and the northern bayou is Little Johnson Bayou. At least that is how those 2 bayous are labeled in Google Maps and OpenTopoMap.

Comments below pertain to bbl_4, and I believe polygon issues with bbl_4 have been corrected. Specific follow-up needed for some stations, such as coordinates that put a station on land, monitoring location ID BAY-WE-STBAY-3 near Burnt Mill Creek. Also, monitoring location ID BAY-WE-STBAY-5 that has a bay assignment of West Bay but coordinates put it in St. Andrew Bay.

Note that Sheepshead Bayou in North Bay actually has Major_Seg of "East Bay" assigned to it. There is a Sheepshead Bayou in East Bay, so maybe the one in North Bay is labelled incorrectly. Also, note that in North Bay, Fannin Bayou has 3 arms to that bayou.

In East Bay, monitoring location id NWFYI0018 is assigned bay value of St. Andrew Bay, which I believe is incorrect as coordinates put that point by Dupont Bridge. Monitoring location ids G3TLHR0043 and G3TLHR0018 are assigned to bay Parker Bayou, which I believe is wrong. Monitoring location id Z1-LL-14008 LAKE MARTIN is assigned to bay Deer Point Lake. Have a monitoring location id of BAY-LA-BAYOU-3 assigned to bay Laird Bayou, but point falls in Cooks Bayou, which is an arm of Laird Bayou. Need to get entire Parker Bayou polygon as currently only have slivers of that polygon, with East Bay polygon overlaying much of Parker Bayou.

Also, Johnson Bayou has a duplicated entry with one bayou near the power plant (and it Major_Seg is St. Andrew Bay). The other Johnson Bayou polygon is off of West Beach Drive, just west of Panama City marina, and it has monitoring location id BAY-JOHNSON-BAYOU-1 located in it. Google Map shows both bayous with a Johnson Bayou label.

Two stations do not fall in St. Andrew Bay polygon, NWFYI0017 and BAY-ST-EWBAY-3. Both have a bay assignment of St. Andrew Bay. NWFYI0017 was collected by FDEP NORTHWEST REGIONAL OPERATIONS CENTER, and BAY-ST-EWBAY-3 was collected by ST. ANDREW BAY RESOURCE MANAGEMENT ASSOCIATION INC. Those 2 stations do fall in the East Bay and Bayous polygon, bayous_eb1.
