---
title: "ESDA of St. Andrew Bay Watershed Water Quality"
author: "Mike McManus and Ruth Horn"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Exploratory Spatial Data Analysis 
The purpose of this script to examine statistical and spatial patterns in the water qualiy of St. Andrews Bay (SAB) watershed using monitoring data collected by several agencies and organizations.

## Libraries

```{r libraries}
library(tidyverse)
library(sf)
library(mapview)

```

## Log

2025-11-21 Bring in csv files


## Reading in Data
Two csv file, one on nutrients and other on conductivity, provided by Rose Horn, SASJBEP, by email on November 13, 2025. Rose pulled the data from WIN 
```{r data}
np_ds1 <- read_csv("data/WQ_Values_Station_Baysection_NP.csv")

spec(np_ds1)
names(np_ds1)

# See Long and Lat in decimal degrees
View(np_ds1) 
str(np_ds1)
glimpse(np_ds1)
coords <- data.frame(longitude  = np_ds1$DEP_Longitude, latitude = np_ds1$DEP_Latitude)
head(coords)

# See Marc Weber workshop above about Coordinate Reference Systems (CRS)

np_sf1 <- st_as_sf(np_ds1, coords = c("DEP_Longitude", "DEP_Latitude"), crs =4269) %>% st_transform((5070))

class(np_sf1)
# read in shapefile
nnc_1 <- st_read("data_gis/NNC_Bay_Segments.shp")
class(nnc_1)
summary(nnc_1)
# see that espg code is 3087
# asking do the two sf objects have the same CRS
st_crs(np_sf1) == st_crs(nnc_1)

np_sf1 <- np_sf1 %>% st_transform(3087)
st_crs(np_sf1) == st_crs(nnc_1)

mapview(np_sf1) + mapview(nnc_1)

# mapview(hab_sf1, zcol = "Depth", cex = "Depth", alpha.regions = .8, legend = TRUE, popup = popupTable(hab_sf1, zcol = c("Depth", "Site"))) + mapview(bigbend_seagrass)

```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
