---
title: "ESDA of St. Andrew Bay Watershed Water Quality"
author: "Mike McManus and Ruth Horn"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Exploratory Spatial Data Analysis

The purpose of this script to examine statistical and spatial patterns in the water qualiy of St. Andrews Bay (sub) watershed using monitoring data collected by several agencies and organizations.

## Libraries

```{r libraries}
library(tidyverse)
library(sf)
library(mapview)
library(janitor)
library(leaflet)
library(leafpop) # for popups in 
library(leafsync) # sync mapview maps
library(patchwork)
library(nhdplusTools)
library(tmap)
library(forcats)
library(gifski) #for tmap animation
```

## Log

2025-12-12 made tmap animation using monthly loop

2025-12-05 made date a factor for tmap animation

2025-12-02 pulled in Bays_Bayous_Lakes_V1.shp

2025-11-22 Unsure the DEP longitude and latitude are actual coordinates of monitoring stations

2025-11-21 Bring in csv files

## Reading in Data and Formatting Date Variables

Two csv file, one on nutrients and other on conductivity, provided by Rose Horn, SASJBEP, by email on November 13, 2025. Rose pulled the data from WIN. Note activity_start_date comes in as character.

Read in Bays_Bayous_Lakes_V1 shapefile that Rose uploaded to repository.

```{r data}
np_ds1 <- read_csv("data/WQ_Values_Station_Baysection_NP.csv")
np_ds1 <- clean_names(np_ds1)
# spec(np_ds1)
# names(np_ds1)


# See Long and Lat in decimal degrees and date as character
View(np_ds1) 
str(np_ds1)
glimpse(np_ds1)
class(np_ds1$activity_start_date)

#zdates <- distinct(np_ds1, activity_start_date)
#ym_ds1 <- lubridate::mdy(zdates$activity_start_date)
#floor_date(ym_ds1, "month")
#saveRDS(zdates, "zdates")

# with activity start date still character have to assign factors
# have to create activity start date as factor for tmap animation
np_ds1$date_factor <- np_ds1$activity_start_date
np_ds1$yrmm_factor <- np_ds1$activity_start_date

# convert to date
np_ds1$activity_start_date <- lubridate::mdy(np_ds1$activity_start_date)
class(np_ds1$activity_start_date)
np_ds1$yrmm <- floor_date(np_ds1$activity_start_date, "month")
head(np_ds1$yrmm)
class(np_ds1$yrmm)

# have to create activity start date as factor for tmap animation
np_ds1 <- arrange(np_ds1, activity_start_date)
np_ds1$date_factor <- as_factor(np_ds1$date_factor)
head(np_ds1$date_factor)
levels(np_ds1$date_factor)


np_ds1$yrmm_factor <- as_factor(as.character(np_ds1$yrmm))
head(np_ds1$yrmm_factor)
levels(np_ds1$yrmm_factor)

```

## Create sf object

```{r create_sf}
# create sf object
coords <- data.frame(longitude  = np_ds1$dep_longitude, latitude = np_ds1$dep_latitude)
head(coords)

# See Marc Weber workshop above about Coordinate Reference Systems (CRS)

np_sf1 <- st_as_sf(np_ds1, coords = c("dep_longitude", "dep_latitude"), crs =4269) %>% st_transform((5070))

class(np_sf1)
# read in shapefile
nnc_1 <- st_read("data_gis/NNC_Bay_Segments.shp")
class(nnc_1)
summary(nnc_1)
# see that espg code is 3087
# asking do the two sf objects have the same CRS
st_crs(np_sf1) == st_crs(nnc_1)

np_sf1 <- np_sf1 %>% st_transform(3087)
st_crs(np_sf1) == st_crs(nnc_1)

mapview(np_sf1) + mapview(nnc_1)

# mapview(hab_sf1, zcol = "Depth", cex = "Depth", alpha.regions = .8, legend = TRUE, popup = popupTable(hab_sf1, zcol = c("Depth", "Site"))) + mapview(bigbend_seagrass)

bbl_1 <- st_read("data_gis/Bays_Bayous_Lakes_V1.shp")
class(bbl_1)
summary(bbl_1)

bbl_1 <- bbl_1 %>% st_transform(3087)

mapview(bbl_1)

mapview(np_sf1) + mapview(bbl_1)

```

## Data Structure and Summaries

Seven distinct organizations collected monitoring data with Florida Lakewatch being a synonym for St. Andrew Baywatch. Get 890 dates spanning 2013-01-24 to 2023-12-20. Have 213 distinct monitoring location IDs. Have 210 distinct monitoring location names. There are 7894 distinct coordinates/locations.

But, I am wondering if DEP assigns coordinates that are not the actual locations of the monitoring station, but instead are always to same place (the centroid) of the spatial feature. Also, are there any analyte code remarks about questionable values?

```{r data_structure}
names(np_ds1)
distinct(np_ds1, organization_name)

distinct(np_ds1, activity_start_date) %>% arrange(., activity_start_date) %>% 
  print(n = Inf)

np_ds1 %>% 
  group_by(organization_name) %>% 
  summarise(count = n_distinct(monitoring_location_name))

distinct(np_ds1, monitoring_location_id) %>% arrange(., monitoring_location_id) %>% 
  print(n = Inf)

distinct(np_ds1, monitoring_location_name) %>% arrange(., monitoring_location_name) %>% 
  print(n = Inf)

distinct(np_ds1, dep_analyte_name)

# distinct monitoring stations
np_stations <- unique.data.frame(np_sf1) # returns 7894 obs/stations & geometries

np_stations %>% 
  group_by(organization_name) %>% 
  summarise(count = n_distinct(monitoring_location_name))


```

Interesting that total phosphorus has monitoring names that include format of DEP's status survey of zone 1 (Z1) for NWFMD region followed by resource type large lake (LL), but that's not the case for total nitrogen.

```{r summaries}

tn_ds1 <- filter(np_ds1, dep_analyte_name == "Nitrogen- Total")



summarize(tn_ds1, min = fivenum(dep_result_value_number)[1],
              q1 = fivenum(dep_result_value_number)[2],
              median = fivenum(dep_result_value_number)[3],
              q3 = fivenum(dep_result_value_number)[4],
              max = fivenum(dep_result_value_number)[5],
              count = dplyr::n(),
              mean = mean(dep_result_value_number))
  
tn_sum_monitoring_name <- tn_ds1 %>% 
  group_by(monitoring_location_name) %>% 
  summarize(min = fivenum(dep_result_value_number)[1],
              q1 = fivenum(dep_result_value_number)[2],
              median = fivenum(dep_result_value_number)[3],
              q3 = fivenum(dep_result_value_number)[4],
              max = fivenum(dep_result_value_number)[5],
              count = dplyr::n(),
              mean = mean(dep_result_value_number))


tp_ds1 <- filter(np_ds1, dep_analyte_name == "Phosphorus- Total")

summarize(tp_ds1, min = fivenum(dep_result_value_number)[1],
              q1 = fivenum(dep_result_value_number)[2],
              median = fivenum(dep_result_value_number)[3],
              q3 = fivenum(dep_result_value_number)[4],
              max = fivenum(dep_result_value_number)[5],
              count = dplyr::n(),
              mean = mean(dep_result_value_number))
  
tp_sum_monitoring_name <- tp_ds1 %>% 
  group_by(monitoring_location_name) %>% 
  summarize(min = fivenum(dep_result_value_number)[1],
              q1 = fivenum(dep_result_value_number)[2],
              median = fivenum(dep_result_value_number)[3],
              q3 = fivenum(dep_result_value_number)[4],
              max = fivenum(dep_result_value_number)[5],
              count = dplyr::n(),
              mean = mean(dep_result_value_number))

    iris %>%
          group_by(Species) %>%
          summarize(min = fivenum(Petal.Width)[1],
                    low = fivenum(Petal.Width)[2],
                    median = fivenum(Petal.Width)[3],
                    upper = fivenum(Petal.Width)[4],
                    max = fivenum(Petal.Width)[5],
                    count = dplyr::n())

```

## Sub-bay Spatial Intersect

Does spatial intersection result in, for example, of upper Crooked Creek being excluded?

```{r st_intersect}
# four sub_bays
subbay_1 <- dplyr::filter(bbl_1, SEGMENT_NA == "West Bay" | SEGMENT_NA == "North Bay" | SEGMENT_NA == "St. Andrew Bay" | SEGMENT_NA == "East Bay")

mapview(subbay_1)
sub_np1 <- st_intersection(np_sf1, subbay_1)
head(sub_np1)


mapview(sub_np1) + mapview(bbl_1)
mapview(sub_np1) + mapview(subbay_1)

```

## Sub-bay Data Structure

```{r subbay_data_struct}
names(sub_np1)
distinct(sub_np1, organization_name) # 4 organizations

distinct(sub_np1, activity_start_date) %>% arrange(., activity_start_date) %>% 
  print(n = Inf) #242 dates from 2016-07-20 to 2023-12-20

sub_np1 %>% 
  group_by(organization_name) %>% 
  summarise(count = n_distinct(monitoring_location_name))

distinct(sub_np1, monitoring_location_id) %>% arrange(., monitoring_location_id) %>% 
  print(n = Inf)

distinct(sub_np1, monitoring_location_name) %>% arrange(., monitoring_location_name) %>%   print(n = Inf)

distinct(sub_np1, dep_analyte_name)

# distinct monitoring stations
sub_np_stations <- unique.data.frame(sub_np1) # returns 1651 obs/stations & geometries

sub_np_stations %>% 
  group_by(organization_name) %>% 
  summarise(count = n_distinct(monitoring_location_name))


```

## Sub-bay Stations

tn_stations and tp_stations get rid of duplicates. Also try get_dupes function in janitor package.

A lot more tp stations than tn stations. Is that because tp almost always has samples with detectable concentrations whereas tn has fewer samples with detectable concentrations? Are nondetects entered into WIN? Do an inner join to find what observations in tn have a matching key in y. Also, I need to check for duplicate entries.

```{r subbay_summaries}
tn_sub1 <- filter(sub_np1, dep_analyte_name == "Nitrogen- Total") 
# tn data is only from Florida Lakewatch

distinct(tn_sub1, organization_name)

# distinct tn monitoring stations
tn_stations <- unique.data.frame(tn_sub1)

mapview(tn_sub1)

mapview(tn_stations)

tn_sub1$bay_section_1 == tn_sub1$SEGMENT_NA
# obs 287 to 303 returns FALSE as have Robinson Bayou vs North Bay
# note that station appears located in St Andrew Bay as below subbay line

summarize(tn_sub1, min = fivenum(dep_result_value_number)[1],
              q1 = fivenum(dep_result_value_number)[2],
              median = fivenum(dep_result_value_number)[3],
              q3 = fivenum(dep_result_value_number)[4],
              max = fivenum(dep_result_value_number)[5],
              count = dplyr::n(),
              mean = mean(dep_result_value_number))
  
tn_sum_sub <- tn_sub1 %>% 
  group_by(SEGMENT_NA) %>% 
  summarize(min = fivenum(dep_result_value_number)[1],
              q1 = fivenum(dep_result_value_number)[2],
              median = fivenum(dep_result_value_number)[3],
              q3 = fivenum(dep_result_value_number)[4],
              max = fivenum(dep_result_value_number)[5],
              count = dplyr::n(),
              mean = mean(dep_result_value_number))


tn_sub1 %>% 
  group_by(organization_name) %>% 
  summarize(min = fivenum(dep_result_value_number)[1],
              q1 = fivenum(dep_result_value_number)[2],
              median = fivenum(dep_result_value_number)[3],
              q3 = fivenum(dep_result_value_number)[4],
              max = fivenum(dep_result_value_number)[5],
              count = dplyr::n(),
              mean = mean(dep_result_value_number))

tp_sub1 <- filter(sub_np1, dep_analyte_name == "Phosphorus- Total")
# lot more tp obs/stations than tn. is that b/c tn samples had a lot of nondetects whereas tp samples are almost always detectable?

distinct(tp_sub1, organization_name)
# FDEP WATERSHED ASSESSMENT SECTION, FLORIDA LAKEWATCH, FDEP NORTHWEST REGIONAL OPERATIONS CENTER,FDEP TALLAHASSEE REGIONAL OPERATIONS CENTER 

# distinct tp monitoring stations
tp_stations <- unique.data.frame(tp_sub1)

mapview(tp_sub1)

tp_locations <- tp_sub1$bay_section_1 == tp_sub1$SEGMENT_NA
print(as.vector(tp_locations))

View(as.data.frame(tp_locations))
# obs 317-321 FALSE corresponds to Sandy Creek vs East Bay
# obs 478-494 FALSE corresponds to Robinson Bayou vs North Bay


summarize(tp_sub1, min = fivenum(dep_result_value_number)[1],
              q1 = fivenum(dep_result_value_number)[2],
              median = fivenum(dep_result_value_number)[3],
              q3 = fivenum(dep_result_value_number)[4],
              max = fivenum(dep_result_value_number)[5],
              count = dplyr::n(),
              mean = mean(dep_result_value_number))

# min tp is 0.002, which I believe is detection limit
  
tp_sum_sub <- tp_sub1 %>% 
  group_by(SEGMENT_NA) %>% 
  summarize(min = fivenum(dep_result_value_number)[1],
              q1 = fivenum(dep_result_value_number)[2],
              median = fivenum(dep_result_value_number)[3],
              q3 = fivenum(dep_result_value_number)[4],
              max = fivenum(dep_result_value_number)[5],
              count = dplyr::n(),
              mean = mean(dep_result_value_number))

tp_sub1 %>% 
  group_by(organization_name) %>% 
  summarize(min = fivenum(dep_result_value_number)[1],
              q1 = fivenum(dep_result_value_number)[2],
              median = fivenum(dep_result_value_number)[3],
              q3 = fivenum(dep_result_value_number)[4],
              max = fivenum(dep_result_value_number)[5],
              count = dplyr::n(),
              mean = mean(dep_result_value_number))

# 65% of obs for tp are from Florida Watershed Assessment Section
    iris %>%
          group_by(Species) %>%
          summarize(min = fivenum(Petal.Width)[1],
                    low = fivenum(Petal.Width)[2],
                    median = fivenum(Petal.Width)[3],
                    upper = fivenum(Petal.Width)[4],
                    max = fivenum(Petal.Width)[5],
                    count = dplyr::n())

```

I am very suspicious that tn_sub1 is almost exactly twice the number of observations of tn_stations, and likewise for tpsub1 and tp_stations. Viewing tn_sub1 shows first two entries for Bay-Parker-Bayou-1 being collected on same date with same tn concentrations. Can we talk to someone at FDEP knowledgeable about WIN to tell us about any QA/QC to check for exact duplicates? Or does Lakewatch always run samples in duplicate?

For now, I am going to believe and assume that tn_stations and tp_stations accurately reflects the number of observations.

## Boxplots

```{r boxplot_tn}

ggplot(tn_stations, aes(x = SEGMENT_NA, y = dep_result_value_number)) + geom_boxplot(varwidth = TRUE) + labs(x = "Segment Name", y = "TN, (mg/L)", title = "Total Nitrogen in Bay Segments", subtitle = "WIN Data from 2018-12-07 to 2023-08-28, n = 420")


ggplot(tn_stations, aes(x = dep_result_value_number, y = monitoring_location_id)) + geom_boxplot(varwidth = TRUE) + labs(y = "Station ID", x = "TN, (mg/L)", title = "Total Nitrogen in Bay Segments", subtitle = "WIN Data from 2018-12-07 to 2023-08-28, n = 420")

ggplot(tn_stations, aes(x = dep_result_value_number, y = monitoring_location_id)) + geom_boxplot(varwidth = TRUE) + labs(y = "Station ID", x = "TN, (mg/L)", title = "Total Nitrogen in Bay Segments", subtitle = "WIN Data from 2018-12-07 to 2023-08-28, n = 420") +
  facet_wrap(vars(SEGMENT_NA))

```

```{r boxplot_tp}

ggplot(tp_stations, aes(x = SEGMENT_NA, y = log(dep_result_value_number))) + geom_boxplot(varwidth = TRUE) + labs(x = "Segment Name", y = "Log TP, (mg/L)", title = "Total Phosphorus in Bay Segments", subtitle = "WIN Data from 2016-07-20 to 2023-12-20, n = 1231")

ggplot(tp_stations, aes(x = SEGMENT_NA, y = log(dep_result_value_number))) + geom_boxplot(varwidth = TRUE) + labs(x = "Segment Name", y = "Log TP, (mg/L)", title = "Total Phosphorus in Bay Segments", subtitle = "WIN Data from 2016-07-20 to 2023-12-20, n = 1231") + facet_wrap(vars(organization_name))

ggplot(tp_stations, aes(x = log(dep_result_value_number), y = monitoring_location_id)) + geom_boxplot(varwidth = TRUE) + labs(y = "Station ID", x = "Log TP, (mg/L)", title = "Total Phosphorus in Bay Segments", subtitle = "WIN Data from 2016-07-20 to 2023-12-20, n = 1231") + facet_wrap(vars(organization_name))


ggplot(tp_stations, aes(x = log(dep_result_value_number), y = monitoring_location_id)) + geom_boxplot(varwidth = TRUE) + labs(y = "Station ID", x = "Log TP, (mg/L)", title = "Total Phosphorus in Bay Segments", subtitle = "WIN Data from 2016-07-20 to 2023-12-20, n = 1231") + facet_wrap(vars(SEGMENT_NA))


ggplot(tp_stations |> filter(organization_name == "FDEP WATERSHED ASSESSMENT SECTION"), aes(x = log(dep_result_value_number), y = monitoring_location_id)) + geom_boxplot(varwidth = TRUE) + labs(y = "Station ID", x = "Log TP, (mg/L)", title = "Total Phosphorus in Bay Segments", subtitle = "WIN Data from 2016-07-20 to 2023-12-20, n = 1231") + facet_wrap(vars(SEGMENT_NA))

ggplot(tp_stations |> filter(organization_name == "FLORIDA LAKEWATCH"), aes(x = log(dep_result_value_number), y = monitoring_location_id)) + geom_boxplot(varwidth = TRUE) + labs(y = "Station ID", x = "Log TP, (mg/L)", title = "Total Phosphorus in Bay Segments", subtitle = "WIN Data from 2016-07-20 to 2023-12-20, n = 1231") + facet_wrap(vars(SEGMENT_NA))


```

## Map Animation

Make sure use tn_stations and tp_stations so don't have duplicate stations. Also, need to set year-month as factor using the ym function in lubridate rather than date. Have both daily and monthly loops.

```{r animate}

tn_bubble <- tm_shape(subbay_1) +
  tm_fill("#dddddd") + 
tm_shape(tn_stations) +
  tm_bubbles(size =  "dep_result_value_number",
             size.legend = tm_legend("TN (mg/L"))

# daily loop
tn_bubble + tm_animate(frame = "date_factor", fps = 1, play = "loop")
# monthly loop
tn_bubble + tm_animate(frame = "yrmm_factor", fps = 1, play = "loop")

tm_shape(World) +
    tm_fill("#dddddd") +
tm_shape(metro) +
    tm_bubbles(size = paste0("pop", seq(1960,2030, by = 10)),
               size.scale = tm_scale_continuous(values.scale = 2),
               size.legend = tm_legend("Population"),
               fill = "gold",
               size.free = FALSE) +
    tm_animate()

tp_bubble <- tm_shape(subbay_1) +
  tm_fill("#dddddd") + 
tm_shape(tp_stations) +
  tm_bubbles(size =  "dep_result_value_number",
             # size.scale = tm_scale_continuous_log(values.scale = "dep_result_value_number"),
             size.legend = tm_legend("TP (mg/L)"))

#daily loop
tp_bubble + tm_animate(frame = "date_factor", fps = 1, play = "loop")

# monthly loop
tp_bubble + tm_animate(frame = "yrmm_factor", fps = 1, play = "loop")

```

## Time Series

```{r time_series}

ggplot(tn_stations, aes(x = activity_start_date, y = dep_result_value_number)) + 
  geom_point()

ggplot(tn_stations, aes(x = activity_start_date, y = dep_result_value_number)) + 
  geom_point() + 
  labs(x = "Date", y = "TN, (mg/L)", title = "Total Nitrogen in Bay Segments", subtitle = "WIN Data from 2018-12-07 to 2023-08-28, n = 420") + 
  facet_wrap(vars(SEGMENT_NA))


ggplot(tp_stations, aes(x = activity_start_date, y = dep_result_value_number)) + 
  geom_point()

ggplot(tp_stations, aes(x = activity_start_date, y = log(dep_result_value_number))) + 
  geom_point()

ggplot(tp_stations, aes(x = activity_start_date, y = log(dep_result_value_number))) + 
  geom_point() + 
  labs(x = "Date", y = "Log TP, (mg/L)", title = "Total Phosphorus in Bay Segments", subtitle = "WIN Data from 2018-12-07 to 2023-08-28, n = 1231") + 
  facet_wrap(vars(SEGMENT_NA))

ggplot(tn_stations, aes(x = SEGMENT_NA, y = dep_result_value_number)) + geom_boxplot(varwidth = TRUE) + labs(x = "Segment Name", y = "TN, (mg/L)", title = "Total Nitrogen in Bay Segments", subtitle = "WIN Data from 2018-12-07 to 2023-08-28, n = 420")


```

## HUCs

```{r hucs}
huc8 <- get_huc(id = '03140101', type = "huc08")

mapview(huc8)

huc8 <- st_transform(huc8, crs = 3087)
```
