---
title: "Linked Micromap 2019"
author: "Mike McManus"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Libraries

```{r libraries}
library(tidyverse)
library(sf)
library(mapview)
library(janitor)
library(leaflet)
library(leafpop) # for popups in 
library(leafsync) # sync mapview maps
library(leafem) # add labels to map
library(patchwork)
library(nhdplusTools)
library(tmap)
library(forcats)
library(gifski) #for tmap animation
library(plotly)
library(webshot) # save mapview object
library(nngeo)
library(micromap)
```

## Log

2026-05-06 renamed with YYYY in script name

2026-05-04 bring over code chunks from script 02 to make this script 03, self-contained

2026-04-24 linked micromap of logged chla with 2 perceptual groups

2026-04-22 pasted workshop code

## Reading in Data and Formatting Date Variables

One csv file, win_baywatch_sabnut.csv, uploaded by Ryann Rossi, SASJBEP, on December 23, 2025. Ryan pulled the data from WIN and did some data cleaning. Note activity_start_date comes in as character.

Read in Bays_Bayous_Lakes_V4 shapefile that Rose uploaded to repository. Can I merge wq data, np_ds1, with Baywatch station IDs, bw_id1, by coordinates?

```{r data}
np_ds1 <- read_csv("data/win_baywatch_sabnut.csv")
names(np_ds1)
np_ds1 <- clean_names(np_ds1)
# spec(np_ds1)
# names(np_ds1)


# See Long and Lat in decimal degrees and date as character
View(np_ds1) 
str(np_ds1)
glimpse(np_ds1)
class(np_ds1$date)

#zdates <- distinct(np_ds1, activity_start_date)
#ym_ds1 <- lubridate::mdy(zdates$activity_start_date)
#floor_date(ym_ds1, "month")
#saveRDS(zdates, "zdates")

# with activity start date still character have to assign factors
# have to create activity start date as factor for tmap animation
np_ds1$date_factor <- np_ds1$date
np_ds1$yrmm_factor <- np_ds1$date

# convert to date
np_ds1$date <- lubridate::mdy(np_ds1$date)
class(np_ds1$date)
np_ds1$yrmm <- floor_date(np_ds1$date, "month")
head(np_ds1$yrmm)
class(np_ds1$yrmm)

# have to create activity start date as factor for tmap animation
np_ds1 <- arrange(np_ds1, date)
np_ds1$date_factor <- as_factor(np_ds1$date_factor)
head(np_ds1$date_factor)
levels(np_ds1$date_factor)


np_ds1$yrmm_factor <- as_factor(as.character(np_ds1$yrmm))
head(np_ds1$yrmm_factor)
levels(np_ds1$yrmm_factor)

# filter(dg_sf1, (!is.na(log_n2o))

np_ds1 <- filter(np_ds1, (!is.na(long)))
# drops to 10,369 obs and summary(mp_ds2$lat) returns no NA

unique(np_ds1$organization_name)

z1 <- filter(np_ds1, is.na(organization_name))
# return 7,162 obs
summary(z1$sta_nr) # code min 1.1 to max 9.5
# believe having a numeric sta_nr code means this is a Lakewatch/Baywatch station and have long/lat coordinates for these stations

z2 <- filter(np_ds1, organization_name == "FLORIDA LAKEWATCH")
# return 2,380
summary(z2$sta_nr) # returns all NAs
# want to get numeric codes assigned to sta_nr, struggling to get that done


```

## Explore Duplicates and NAs

```{r dupes_na}
# z1 <- get_dupes(np_ds1)
# 
# z2 <- filter(np_ds1, (is.na(tn) & !is.na(tp)))
# 
# z3 <- filter(np_ds1, is.na(tn))
```

## Create sf object

Have NA for some lat and long entries so remove these. Also, Numeric Nutrient Criteria (NNC) polygons omit bayous, creeks, and lakes, but includes intracoastal waterway and St. Joe Bay. But bayous, bays, and lakes does have all those polygons. Although, a Burnt Mill Creek station does not hit that polygon. Not clear to me why Bay-West Bay-3 monitoring location on April 25, 2024 is being plotted so far from Burnt Mill Creek.

```{r create_sf}

# create sf object
coords <- data.frame(longitude  = np_ds1$long, latitude = np_ds1$lat)
head(coords)

# See Marc Weber workshop above about Coordinate Reference Systems (CRS)

np_sf1 <- st_as_sf(np_ds1, coords = c("long", "lat"), crs =4269) %>% st_transform((5070))

class(np_sf1)
# read in shapefile
nnc_1 <- st_read("data_gis/NNC_Bay_Segments.shp")
class(nnc_1)
summary(nnc_1)
distinct(nnc_1, SEGMENT_NA)
distinct(nnc_1, WATERBODY_)

mapview(nnc_1)
# Numeric

# see that espg code is 3087
# asking do the two sf objects have the same CRS
st_crs(np_sf1) == st_crs(nnc_1)

np_sf1 <- np_sf1 %>% st_transform(3087)
st_crs(np_sf1) == st_crs(nnc_1)

mapview(np_sf1) + mapview(nnc_1)

# mapview(hab_sf1, zcol = "Depth", cex = "Depth", alpha.regions = .8, legend = TRUE, popup = popupTable(hab_sf1, zcol = c("Depth", "Site"))) + mapview(bigbend_seagrass)


```

## Read Bays, Bayous and Lakes V4

Rose Horn took 2 polygons for Alligator Bayou, which borders the U.S. Coast Guard Station on St. Andrew Bay, and splits them into separate sampling areas. Also, in version the new update reflects the removal of small edge portions of St. Andrew Bay that were along the edges of East Bay and Grand Lagoon. Looks like I should pick-up Upper and Lower Burnt Mill Creek Stations.

Need to see if I have 8 polygon names for: East Bay, Grand Lagoon, North Bay, St. Andrew Bay, West Bay, Crooked and Burnt Mill Creeks, and Deer Point Lake as I want to use those 6 polygons in linked micromap. Perhaps by using SEGMENT_NA and Major_Seg I can get the polygons I want.

```{r bbl_v4}

bbl_4 <- st_read("data_gis/Bays_Bayous_Lakes_V4/Bays_Bayous_Lakes_V4.shp")
summary(bbl_4)
mapview(bbl_4)

distinct(bbl_4, SEGMENT_NA) %>% dplyr::arrange(., SEGMENT_NA)
# have 190 distinct SEGMENT_NA, but 191 obs suggesting one polygon is sharing a name

# Is Major_Seg a new field from Rose?
distinct(bbl_4, Major_Seg) %>% dplyr::arrange(., Major_Seg)
# returns 8 Major_Seg

bbl_4 <- bbl_4 %>% st_transform(3087)

mapview(np_sf1) + mapview(bbl_4)
# a burnt mill creek station from 2024-04-25 is off polygon

# these 8 polygons might be what I use for a linked micromap
polygons8 <- dplyr::filter(bbl_4, SEGMENT_NA == "West Bay" | SEGMENT_NA == "North Bay" | SEGMENT_NA == "St. Andrew Bay" | SEGMENT_NA == "East Bay" | SEGMENT_NA == "Burnt Mill Creek" | SEGMENT_NA == "Crooked Creek" | SEGMENT_NA == "Grand Lagoon" |SEGMENT_NA == "Deer Point Lake")

mapview(polygons8)

mapview(np_sf1) + mapview(polygons8)

```

## Stations for Linked Micromap

Inconsistent data entry for Lake Marin or just Marin. The "mm\_" naming convention refers to getting the data structured for a linked micromap.

```{r micromap_stations}

mm_station1 <- dplyr::filter(np_sf1, bay != "Lake Powell" & bay != "Powell Creek" & bay != "Lullwater" & bay != "Lake Marin" & bay != "Marin")

names(mm_station1)

mapview(mm_station1)

distinct(mm_station1, bay) %>% dplyr::arrange(., bay) %>% print(n=Inf) 

# np_ds1 %>% 
#   group_by(organization_name) %>% 
#   summarise(count = n_distinct(monitoring_location_name))

mapview(mm_station1) + mapview(polygons8, col.regions = "yellow")

```

## Spatial Intersection of MM Stations to Polygons 8

Spatial intersection so station points get polygon attributes, specifically SEGMENT_NA. For linked micromap the data frame of summary statistics has to have the same ID linking variable as in the polygon data frame. Note this intersection will drop many bayou and creek stations, but I am okay as this is just for a draft linked micromap.

In West Bay, the spatial intersection omits the intracoastal waterway, and upper stations in Crooked and Burnt Mill Creeks. In North Bay, omit many of the bayous, and likewise for East Bay.

```{r mmstations_intersect_polygons}

mm_station2 <- st_intersection(mm_station1, polygons8)

mm_stat1map <- mapview(mm_station1)
mm_stat2map <- mapview(mm_station2, col.regions = "yellow")

sync(mm_stat1map, mm_stat2map, ncol = 1)

mapview(mm_station2) + mapview(polygons8, col.regions = "yellow")

names(mm_station2)

mm_station2 %>% 
  group_by(SEGMENT_NA) %>% 
  summarize(., count = dplyr::n())

mm_station2 %>% 
  group_by(SEGMENT_NA, year) %>% 
  summarize(., count = dplyr::n()) %>% 
  print(n=Inf)

summary(mm_station2$sta_nr)
# returns 1727 NAs so would likely need to do nearest neighbor analysis to get numeric codes assigned to sta_nr

# na's for sta_nr appear in 2016 and later years
# will need to work this out so can get an accurate count of distinct within each
# of the polygons
mm_station2 %>% 
  group_by(SEGMENT_NA, year) %>% 
  summarize(., na_count = sum(is.na(sta_nr))) %>% 
print(n = Inf)

# np_ds1 %>% 
#   group_by(organization_name) %>% 
#   summarise(count = n_distinct(monitoring_location_name))

# dplyr::select(sub_np_stations, tn, SEGMENT_NA) %>% 
# group_by(., SEGMENT_NA) %>% 
# summarize(., min = fivenum(tn)[1],
#               q1 = fivenum(tn)[2],
#               median = fivenum(tn)[3],
#               q3 = fivenum(tn)[4],
#               max = fivenum(tn)[5],
#               count = dplyr::n(),
#               mean = mean(tn, na.rm = TRUE),
#               na_count = sum(is.na(tn))) %>% 
# print(n = Inf)
distinct(mm_station2, SEGMENT_NA) %>% dplyr::arrange(., SEGMENT_NA)

distinct(mm_station2, organization_id)
distinct(mm_station2, organization_name)
distinct(mm_station2, year) %>% dplyr::arrange(., year) %>% print(n=Inf)
distinct(mm_station2, sampling_agency_name)

#2019-2023 covers St. Andrew Baywatch WQ report
dplyr::filter(mm_station2, year == 2019 & sampling_agency_name == "ST. ANDREW BAY RESOURCE MANAGEMENT ASSOCIATION INC.") %>% 
dplyr::select(., tn, SEGMENT_NA) %>% 
group_by(., SEGMENT_NA) %>% 
summarize(., min = fivenum(tn)[1],
              q1 = fivenum(tn)[2],
              median = fivenum(tn)[3],
              q3 = fivenum(tn)[4],
              max = fivenum(tn)[5],
              count = dplyr::n(),
              mean = mean(tn, na.rm = TRUE),
              na_count = sum(is.na(tn))) %>% 
print(n = Inf)


# get years that match St. Andrew Baywatch WQ Report
mm_station3 <- dplyr::filter(mm_station2, sampling_agency_name == "ST. ANDREW BAY RESOURCE MANAGEMENT ASSOCIATION INC.")
# returns 1082 obs

distinct(mm_station3, organization_id)
distinct(mm_station3, organization_name)
distinct(mm_station3, sampling_agency_name)
distinct(mm_station3, year)

# mm_stations3$yr <- as.factor(year)

# note in 2018 facet no East Bay samples for tn, tp and chl
# for chl all missing values in 2021 and some in 2023 and 2024
tm_shape(polygons8) +
  tm_polygons(fill = "white", lwd = 2) +
  tm_title("Detectable and NA Entries of Total Nitrogen (mg/L)")  + 
  tm_shape(mm_station3) +
  tm_symbols(fill = "tn") +
  tm_facets(by = "year")
```

## Statistics for Linked Micromaps

Get 5-number summary statistics to use in making boxplot for linked micromaps. Also, getting stats on log-transformed data.

```{r stats_micromap}
# remove geometry from summary stats
chla_2019 <- st_set_geometry(mm_station2, NULL) %>% 
dplyr::filter(., year == 2019 & sampling_agency_name == "ST. ANDREW BAY RESOURCE MANAGEMENT ASSOCIATION INC.") %>% 
dplyr::select(., chl, SEGMENT_NA) %>% 
group_by(., SEGMENT_NA) %>% 
summarize(., chla_min = fivenum(chl)[1],
              chla_q1 = fivenum(chl)[2],
              chla_med = fivenum(chl)[3],
              chla_q3 = fivenum(chl)[4],
              chla_max = fivenum(chl)[5],
              chla_lmin = log(fivenum(chl)[1]),
              chla_lq1 = log(fivenum(chl)[2]),
              chla_lmed = log(fivenum(chl)[3]),
              chla_lq3 = log(fivenum(chl)[4]),
              chla_lmax = log(fivenum(chl)[5]),
              chla_count = dplyr::n(),
              chla_mean = mean(chl, na.rm = TRUE),
              chla_na_count = sum(is.na(chl))) %>% 
print(n = Inf)

tn_2019 <- st_set_geometry(mm_station2, NULL) %>% 
  dplyr::filter(., year == 2019 & sampling_agency_name == "ST. ANDREW BAY RESOURCE MANAGEMENT ASSOCIATION INC.") %>% 
  dplyr::select(., tn, SEGMENT_NA) %>% 
  group_by(., SEGMENT_NA) %>% 
  summarize(., tn_min = fivenum(tn)[1],
            tn_q1 = fivenum(tn)[2],
            tn_med = fivenum(tn)[3],
            tn_q3 = fivenum(tn)[4],
            tn_max = fivenum(tn)[5],
            tn_lmin = log(fivenum(tn)[1]),
            tn_lq1 = log(fivenum(tn)[2]),
            tn_lmed = log(fivenum(tn)[3]),
            tn_lq3 = log(fivenum(tn)[4]),
            tn_lmax = log(fivenum(tn)[5]),
            tn_count = dplyr::n(),
            tn_mean = mean(tn, na.rm = TRUE),
            tn_na_count = sum(is.na(tn))) %>% 
  print(n = Inf)

tp_2019 <- st_set_geometry(mm_station2, NULL) %>% 
  dplyr::filter(., year == 2019 & sampling_agency_name == "ST. ANDREW BAY RESOURCE MANAGEMENT ASSOCIATION INC.") %>% 
  dplyr::select(., tp, SEGMENT_NA) %>% 
  group_by(., SEGMENT_NA) %>% 
  summarize(., tp_min = fivenum(tp)[1],
            tp_q1 = fivenum(tp)[2],
            tp_med = fivenum(tp)[3],
            tp_q3 = fivenum(tp)[4],
            tp_max = fivenum(tp)[5],
            tp_lmin = log(fivenum(tp)[1]),
            tp_lq1 = log(fivenum(tp)[2]),
            tp_lmed = log(fivenum(tp)[3]),
            tp_lq3 = log(fivenum(tp)[4]),
            tp_lmax = log(fivenum(tp)[5]),
            tp_count = dplyr::n(),
            tp_mean = mean(tp, na.rm = TRUE),
            tp_na_count = sum(is.na(tp))) %>% 
  print(n = Inf)



```

## Edit Polygons and Stats

Edit data, both polygons and stats, to just the areas of interest. Dropping Burnt Mill Creek and Crooked Creek polygons and stats as have too few observations. Also, merge the chlorophyll, tn, and tp statistics to the polygon data as a single sf object.

```{r polygons_stats}
names(polygons8)
View(polygons8)

polygons6 <- filter(polygons8, SEGMENT_NA != "Crooked Creek" & SEGMENT_NA != "Burnt Mill Creek")

#chla
chla_2019v1 <- filter(chla_2019, SEGMENT_NA != "Crooked Creek")
names(chla_2019v1)

# merge polygon to chla stats
# list sf object first, then dataframe of stats to create polygons with stats
chla_2019v2 <- merge(polygons6, chla_2019v1, by = "SEGMENT_NA" )
summary(chla_2019v2)
# note that geometry is polygon

# tn
tn_2019v1 <- filter(tn_2019, SEGMENT_NA != "Crooked Creek")

chlatn_2019v2 <- merge(chla_2019v2, tn_2019v1, by = "SEGMENT_NA" )

# tp
tp_2019v1 <- filter(tp_2019, SEGMENT_NA != "Crooked Creek")

nutr_2019 <- merge(chlatn_2019v2, tp_2019v1, by = "SEGMENT_NA" )

```

## Draft Linked Micromap

Make draft linked micromaps to get code working, and note any statistical and spatial patterns.

Note cannot have a column called "median" in data. Also, for chla data needed log transformation, as did tn and tp. Getting 3 micromaps, one uncolored, one with 6 colors, and one all gray when grouping was 6. With grouping = 3 get 2 micromaps corresponding to the 2 perceptual groups.

For chla, interesting how little variation West Bay shows. For tn, Grand Lagoon has highest median and largest variation, and West Bay has an outlier that almost exceed Grand Lagoon. For tp, Grand Lagoon shows previous pattern similar to tn. Is tp, also showing a salinity graident correlation with high tp in higher salinity waters?

```{r draft_linked_micromap}
#Draft plot
mmplot(map.data=nutr_2019,
       panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       list('chla_lmin', 'chla_lq1', 'chla_lmed', 'chla_lq3', 'chla_lmax'),
                       NA),
       grouping = 3,
       ord.by='chla_lmed', rev.ord=TRUE,
       median.row=FALSE)
       

mmplot(map.data=nutr_2019,
       panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       list('tn_lmin', 'tn_lq1', 'tn_lmed', 'tn_lq3', 'tn_lmax'),
                       NA),
       grouping = 3,
       ord.by='tn_lmed', rev.ord=TRUE,
       median.row=FALSE)


mmplot(map.data=nutr_2019,
       panel.types=c('dot_legend', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 3,
       ord.by='tp_lmed', rev.ord=TRUE,
       median.row=FALSE)

```

## Refine Linked Micromaps

Use more detailed coding arguments to make better looking linked micromaps. Refined TP

```{r refine_micromap_tp}
# TP Refined Plot
mmplot(map.data=nutr_2019,
       panel.types=c('dot_legend', 'labels', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       'tp_count',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 3,
       ord.by='tp_lmed',
       rev.ord=TRUE,
       median.row=FALSE,
       plot.pGrp.spacing=1,
       
       plot.height=6, plot.width=9, #for poster use plot.height=6, plot.width=9,
       colors=brewer.pal(5, "Spectral"), #for journal use plot.height=9, plot.width=6.8
              
       panel.att=list(list(1, panel.width=1, point.type=20, point.size=2,
                           point.border=TRUE), 
                      
                      list(2, header='St.Andrew Bay Segments', panel.width=1.2,#increased panel.width to 1.1 from .8
                           align='left', text.size=.8),
                      
                      list(3, header='Samples',
                           panel.width=.35,
                           align='right',
                           text.size=.8),
                      
                      list(4, header='2019 Detectable Total Phoshporous',
                           graph.bgcolor='lightgray',
                           xaxis.ticks=c(-6.0, -5.0, -4.0, -3.0, -2.0),
                           xaxis.labels=c(-6,-5, -4, -3, -2),
                           xaxis.labels.size=1.25,
                           graph.bar.size = 0.5,
                           #xaxis.labels.angle=90,
                           xaxis.title='Log(TP, mg/L)', xaxis.title.size=1.25, panel.width=2),
                      
                      list(5, header='Light Gray Means Highlighted Above',
                           inactive.border.color=gray(.7), inactive.border.size=2,
                            panel.width=1.5)),
print.file = "output_micromap/mmplot_tp2019.jpeg", res=100)



```

Refined TN

```{r}
# TN Refined Plot
mmplot(map.data=nutr_2019,
       panel.types=c('dot_legend', 'labels', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       'tn_count',
                       list('tn_lmin', 'tn_lq1', 'tn_lmed', 'tn_lq3', 'tn_lmax'),
                       NA),
       grouping = 3,
       ord.by='tn_lmed',
       rev.ord=TRUE,
       median.row=FALSE,
       plot.pGrp.spacing=1,
       
       plot.height=6, plot.width=9, #for poster use plot.height=6, plot.width=9,
       colors=brewer.pal(5, "Spectral"), #for journal use plot.height=9, plot.width=6.8
              
       panel.att=list(list(1, panel.width=1, point.type=20, point.size=2,
                           point.border=TRUE), 
                      
                      list(2, header='St.Andrew Bay Segments', panel.width=1.2,#increased panel.width to 1.1 from .8
                           align='left', text.size=.8),
                      
                      list(3, header='Samples',
                           panel.width=.35,
                           align='right',
                           text.size=.8),
                      
                      list(4, header='2019 Detectable Total Nitrogen',
                           graph.bgcolor='lightgray',
                           xaxis.ticks=c(-2.5, -2.0, -1.5, -1.0, -0.5, -0.0),
                           xaxis.labels=c(-2.5, -2.0, -1.5, -1.0, -0.5, -0.0),
                           xaxis.labels.size=1.25,
                           graph.bar.size = 0.5,
                           #xaxis.labels.angle=90,
                           xaxis.title='Log(TN, mg/L)', xaxis.title.size=1.25, panel.width=2),
                      
                      list(5, header='Light Gray Means Highlighted Above',
                           inactive.border.color=gray(.7), inactive.border.size=2,
                            panel.width=1.5)),
# to run so just appears in Plots pane have 3 parentheses after the last panel.width
# to run and save the image delete 3rd parenthesis and insert comma       
print.file = "output_micromap/mmplot_tn2019.jpeg", res=100)


```

Refined Chlorophyll a Note need to add column "\# of NA Samples".

```{r}
# Chla Refined Plot
mmplot(map.data=nutr_2019,
       panel.types=c('dot_legend', 'labels', 'labels', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       'chla_count',
                       'chla_na_count',
                       list('chla_lmin', 'chla_lq1', 'chla_lmed', 'chla_lq3', 'chla_lmax'),
                       NA),
       grouping = 3,
       ord.by='chla_lmed',
       rev.ord=TRUE,
       median.row=FALSE,
       plot.pGrp.spacing=1,
       
       plot.height=6, plot.width=9, #for poster use plot.height=6, plot.width=9,
       colors=brewer.pal(5, "Spectral"), #for journal use plot.height=9, plot.width=6.8
              
       panel.att=list(list(1, panel.width=1, point.type=20, point.size=2,
                           point.border=TRUE), 
                      
                      list(2, header='St.Andrew \n Bay Segments', panel.width=1,#increased panel.width to 1.1 from .8
                           align='left', text.size=.8),
                      
                      list(3, header='Total #\n of Samples',
                           panel.width=.5,
                           align='right',
                           text.size=.8),
                      
                      list(4, header='# of NA\n Samples',
                           panel.width=.4,
                           align='right',
                           text.size=.8),
                      
                      list(5, header='2019 Detectable\n Chlorophyll a',
                           graph.bgcolor='lightgray',
                           xaxis.ticks=c(0,1.5,3,4.5),
                           xaxis.labels=c(0,1.5,3,4.5),
                           xaxis.labels.size=1.25,
                           graph.bar.size = 0.5,
                           #xaxis.labels.angle=90,
                           xaxis.title='Log(µg/L)', xaxis.title.size=1.25, panel.width=1.75),
                      
                      list(6, header='Light Gray Means\n Highlighted Above',
                           inactive.border.color=gray(.7), inactive.border.size=2,
                            panel.width=1.75)),
# to run so just appears in Plots pane have 3 parentheses after the last panel.width
# to run and save the image replace 3rd parenthesis with comma       
print.file = "output_micromap/mmplot_chla2019.jpeg", res=100)


```

## Add Numeric Nutrient Criteria Lines

Apparently, can only add one line and no more in a panel. So, I added ln of 0.019 mg/L, which in the annual geometric mean for St. Andrew Bay, and the highest AGM among, East (0.016 mg/L), North (0.014 mg/L) and West (0.017 mg/L) Bays.

```{r nnc_micromap_tp}
# TP Refined Plot
mmplot(map.data=nutr_2019,
       panel.types=c('dot_legend', 'labels', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       'tp_count',
                       list('tp_lmin', 'tp_lq1', 'tp_lmed', 'tp_lq3', 'tp_lmax'),
                       NA),
       grouping = 3,
       ord.by='tp_lmed',
       rev.ord=TRUE,
       median.row=FALSE,
       plot.pGrp.spacing=1,
       
       plot.height=6, plot.width=9, #for poster use plot.height=6, plot.width=9,
       colors=brewer.pal(5, "Spectral"), #for journal use plot.height=9, plot.width=6.8
              
       panel.att=list(list(1, panel.width=1, point.type=20, point.size=2,
                           point.border=TRUE), 
                      
                      list(2, header='St.Andrew Bay Segments', panel.width=1.2,#increased panel.width to 1.1 from .8
                           align='left', text.size=.8),
                      
                      list(3, header='Samples',
                           panel.width=.35,
                           align='right',
                           text.size=.8),
                      
                      list(4, header='2019 Detectable Total Phoshporous',
                           graph.bgcolor='lightgray',
                           xaxis.ticks=c(-6.0, -5.0, -4.0, -3.0, -2.0),
                           xaxis.labels=c(-6,-5, -4, -3, -2),
                           xaxis.labels.size=1.25,
                           graph.bar.size = 0.5,
                           #add.line=-4.27, add.line.col="black", add.line.typ="dashed",
                           add.line=-3.96, add.line.col="black", add.line.typ="solid",
                           xaxis.title='Log(TP, mg/L)', xaxis.title.size=1.25, panel.width=2),
                      
                      list(5, header='Light Gray Means Highlighted Above',
                           inactive.border.color=gray(.7), inactive.border.size=2,
                            panel.width=1.5)),
print.file = "output_micromap/mmplot_tp2019v2.jpeg", res=100)



```

Added line for West Bay annual geometric mean threshold, which is 0.35mg/L, and on natura log scale is -1.050. West Bay AGM is maximum among the Bay segments.

```{r nnc_tn_micromap}
# TN Refined Plot
mmplot(map.data=nutr_2019,
       panel.types=c('dot_legend', 'labels', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       'tn_count',
                       list('tn_lmin', 'tn_lq1', 'tn_lmed', 'tn_lq3', 'tn_lmax'),
                       NA),
       grouping = 3,
       ord.by='tn_lmed',
       rev.ord=TRUE,
       median.row=FALSE,
       plot.pGrp.spacing=1,
       
       plot.height=6, plot.width=9, #for poster use plot.height=6, plot.width=9,
       colors=brewer.pal(5, "Spectral"), #for journal use plot.height=9, plot.width=6.8
              
       panel.att=list(list(1, panel.width=1, point.type=20, point.size=2,
                           point.border=TRUE), 
                      
                      list(2, header='St.Andrew Bay Segments', panel.width=1.2,#increased panel.width to 1.1 from .8
                           align='left', text.size=.8),
                      
                      list(3, header='Samples',
                           panel.width=.35,
                           align='right',
                           text.size=.8),
                      
                      list(4, header='2019 Detectable Total Nitrogen',
                           graph.bgcolor='lightgray',
                           xaxis.ticks=c(-2.5, -2.0, -1.5, -1.0, -0.5, -0.0),
                           xaxis.labels=c(-2.5, -2.0, -1.5, -1.0, -0.5, -0.0),
                           xaxis.labels.size=1.25,
                           graph.bar.size = 0.5,
                           add.line=-1.050, add.line.col="black", add.line.typ="solid",
                           #xaxis.labels.angle=90,
                           xaxis.title='Log(TN, mg/L)', xaxis.title.size=1.25, panel.width=2),
                      
                      list(5, header='Light Gray Means Highlighted Above',
                           inactive.border.color=gray(.7), inactive.border.size=2,
                            panel.width=1.5)),
# to run so just appears in Plots pane have 3 parentheses after the last panel.width
# to run and save the image delete 3rd parenthesis and insert comma       
print.file = "output_micromap/mmplot_tn2019v2.jpeg", res=100)


```

Added East Bay annual geometric line of 3.9 micrograms/L, which is maximum AGM among Bay segments, and is 1.4 with natural log transformation.

```{r nnc_chla_micromap}
# Chla Refined Plot
mmplot(map.data=nutr_2019,
       panel.types=c('dot_legend', 'labels', 'labels', 'labels', 'box_summary', 'map'),
       panel.data=list(NA,
                       'SEGMENT_NA',
                       'chla_count',
                       'chla_na_count',
                       list('chla_lmin', 'chla_lq1', 'chla_lmed', 'chla_lq3', 'chla_lmax'),
                       NA),
       grouping = 3,
       ord.by='chla_lmed',
       rev.ord=TRUE,
       median.row=FALSE,
       plot.pGrp.spacing=1,
       
       plot.height=6, plot.width=9, #for poster use plot.height=6, plot.width=9,
       colors=brewer.pal(5, "Spectral"), #for journal use plot.height=9, plot.width=6.8
              
       panel.att=list(list(1, panel.width=1, point.type=20, point.size=2,
                           point.border=TRUE), 
                      
                      list(2, header='St.Andrew \n Bay Segments', panel.width=1,#increased panel.width to 1.1 from .8
                           align='left', text.size=.8),
                      
                      list(3, header='Total #\n of Samples',
                           panel.width=.5,
                           align='right',
                           text.size=.8),
                      
                      list(4, header='# of NA\n Samples',
                           panel.width=.4,
                           align='right',
                           text.size=.8),
                      
                      list(5, header='2019 Detectable\n Chlorophyll a',
                           graph.bgcolor='lightgray',
                           xaxis.ticks=c(0,1.5,3,4.5),
                           xaxis.labels=c(0,1.5,3,4.5),
                           xaxis.labels.size=1.25,
                           graph.bar.size = 0.5,
                           add.line=1.4, add.line.col="black", add.line.typ="solid",
                           #xaxis.labels.angle=90,
                           xaxis.title='Log(µg/L)', xaxis.title.size=1.25, panel.width=1.75),
                      
                      list(6, header='Light Gray Means\n Highlighted Above',
                           inactive.border.color=gray(.7), inactive.border.size=2,
                            panel.width=1.75)),
# to run so just appears in Plots pane have 3 parentheses after the last panel.width
# to run and save the image replace 3rd parenthesis with comma       
print.file = "output_micromap/mmplot_chla2019v2.jpeg", res=100)


```

